### 2024-07-11 src/main.tex
intital skeleton created, related work section

\documentclass[12pt]{article}
\usepackage{graphicx}
\usepackage[utf8]{inputenc}

\begin{document}

\begin{titlepage}

\newcommand{\HRule}{\rule{\linewidth}{0.5mm}}
\center

%----------------------------------------------------------------------------------------
%	HEADING SECTIONS
%----------------------------------------------------------------------------------------

\textsc{\LARGE The University of Iowa}\\[1.5cm]
\includegraphics[scale=.1]{figures/iowa}\\[1cm]
\textsc{\Large Report for Qualifying Exam}\\[0.5cm]
%\textsc{\large Ph.D.}\\[0.5cm]
\title {Report for Qualifying Exam}
\HRule \\[0.4cm]
{ \huge \bfseries Understanding the flow of Temporal Context Shift in Data Voids}\\[0.4cm]
\HRule \\[1.5cm]

\emph{Author:} \\
Manisha \textsc{Keim} \\
manisha-keim@uiowa.edu \\[0.5cm]

\emph{Advisor:}\\
Rishab \textsc{Nithyanand}\\
rishab-nithyanand@uiowa.edu \\[0.5cm]

Department of Computer Science

{\today}\\[2cm]

\end{titlepage}

%%%%%%%%%%%%%%%%%%%%%%%%%%% Table of Contents
\tableofcontents
\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%% Main document
\section{Abstract} \label{sec:abstract}
    In the digital age, online information seekers frequently encounter search results laden with misinformation, biased narratives, and conspiracy theories.
    This exposure can potentially lead users to accept and propagate false information.
    However, this phenomenon is not universal across all search queries.
    Certain searches, known as \textit{data voids} are particularly susceptible to these issues.
    \\
    \textit{Data voids} occur when searching for terms that yield limited, non-existent, or highly problematic relevant information.
    Unlike common searches that produce abundant data, queries falling into data voids may return no results or present irrelevant and often inaccurate information.
    Data voids can be categorized into two main types: newly coined terms with no established information base, and existing terms whose meanings have evolved over time.
    This report focuses on the latter, examining data voids where context drift occurs.
    The following papers has been selected to explore this phenomenon.
    \begin{itemize}
        \item Statistically Significant Detection of Linguistic Change. ACM WWW 2014 \cite{kulkarni2014statistically}
        \item Diachronic Word Embeddings Reveal Statistical Laws of Semantic Change. ACL 2016 \cite{hamilton2018diachronic}
        \item Specious Sites: Tracking the Spread and Sway of Spurious News Stories at Scale. IEEE S\&P 2024 \cite{hanley2024specious}
    \end{itemize}

\section{Introduction} \label{sec:introduction}
\subsection{What is a Data Void?}
\subsection{Example of a data void?(Migrant Caravan)}
\subsection{Lifecycle of data void (from long tail to spike)}
\subsection{Types of Data Void}
\subsection{Technical Challenge}
analyzing data void involves context drift over historical time. (How context drift happens with data void.)

\newpage
\section{Related Work} \label{sec:relatedwork}
The evolution of word meanings over time has been a subject of significant interest.
Researchers have employed a diverse array of methodologies, including word embeddings, neural language models, and diachronic analysis, to track and examine semantic shifts across historical periods.
Much of this research has utilized the Google Books Ngram corpus, a vast collection spanning from 1900 to 2009, encompassing over 500 billion books in seven languages.
This dataset provides n-grams with corresponding yearly occurrences and frequencies.

Diverse categories of semantic shifts exist, such as semantic, syntactic, or as Hamilton et al. (2016) noted, can be driven by cultural or linguistic factors.
Certain studies concentrate on semantic changes, while others explore the broader evolution that words undergo, with some works focus on transformations in word usage and meaning.

\vspace{5mm}
Gulordava and Baroni (2011) introduce a novel method for finding semantic changes in words over time.
They focus on detecting shifts between the 1960s and 1990s using the Google Books Ngram corpus.
Their approach relies on comparing the similarity of a word's surrounding words in these two periods.
They define context by considering the words that appear within two positions (a window size of 2) before and after the target word, essentially analyzing bigram statistics.
%Example: \'parent\' and \'sleep\'

Kim et al. (2014) propose a distinct approach that leverages the power of neural language models.
By analyzing how the model represents words at different points in time, they can capture how word meaning evolves.
Unlike previous methods that focus on individual words or word frequency, this technique allows them to pinpoint not only words with shifting meanings but also the specific years when those changes occurred.
Their approach goes beyond simple co-occurrence by using a window size of four and comparing the cosine similarity between the same word's vector representations across different time periods.
This enables them to detect even subtle changes in how words are used.
Additionally, by plotting the time series of a word's distance to its neighboring words in the model's vector space, they can visualize the exact period during which the semantic shift occurred.
They identified words like {\lq{cell}\rq} and {\lq{gay}\rq} that have undergone significant semantic shifts over the analyzed period.

Kulkarni et al. (2015) introduced method for significant detection of linguistic change, which laid out the foundation that utilized word embeddings for detecting semantic shifts.
Building on the concept of distributed representations introduced by Hinton, they map symbolic data (words) into a continuous vector space where words with similar meanings are positioned close together.
Their approach incorporates three methods to analyze these embeddings: frequency-based, syntactic, and distributional.

Furthering the research on word embeddings, Liao and Cheng (2016) compared the use of word embeddings with a clustering algorithm called DBSCAN to analyze semantic change.
Their work investigated which method was more effective for identifying these semantic shifts.

Hamilton et al. (2016) address the challenge of distinguishing between cultural shifts and linguistic drift, both of which can contribute to semantic change.
They propose two distinct computational measures based on word embeddings.
One measure captures global semantic shifts, while the other, captures localized changes that reflect cultural influences.
By analyzing these two measures together, researchers can know the causes behind semantic change.

Expanding on this concept, their subsequent work titled "\textit{Diachronic Word Embeddings Reveal Statistical Laws of Semantic Change}" (2016) investigates recurring patterns in semantic change across different languages.
They identify two key principles: \textit{the law of conformity}, which suggests that a word's frequency is inversely proportional to the rate of its semantic change.
In other words, words that are used more frequently tend to experience slower shifts in meaning.
\textit{The law of innovation}, on the other hand, proposes that words with multiple meanings are more likely to undergo semantic transformations over time.

Furthermore, Kutuzov et al. (2018) conducted a survey on diachronic word embeddings and semantic shifts, consolidating the existing academic research in this domain.
Their work provides a comprehensive overview of the methodologies and findings related to tracking semantic changes over time using computational techniques.

Giulianelli (2020) explores the computational analysis of lexical semantic change using contextualized word representations.
Contextualized word representations, such as those derived from models like BERT or ELMo takes into account the context in which words are used.
This enables a more detailed analysis of how semantic changes occur based on specific word contexts.

The impact of significant events like the COVID-19 pandemic on language and semantic change has also been a subject of study.
Laurino et al. (2023) explores tracking fast semantic changes through a large-scale word association task, aiming to understand how the collective mental lexicon evolves in response to such global events.
Their research highlights the dynamic nature of language and how it incorporates new senses.
For instance, words like "quarantine," "mask," and "social distancing" took on new and prominent meanings in everyday conversation.

 \vspace{5mm}
%Among the reviewed literature, three studies are particularly pertinent to our analysis of data voids.
%In case of data voids, we want to find shifts from it's previous usage to new usage.
Several studies from our literature review provide valuable insights for addressing data voids in our analysis.
\begin{itemize}
    \item The first being the work by Kulkarni et al. which introduces us to time series construction.
Their distributional methods focuses on finding subtle semantic shits to determine the context where a word in used.
This concept aligns perfectly with our goal of uncovering shifts in word usage when encountering data voids.
    \item Second paper by Hamilton et al. introduces us to word emebeddings alignment.
Since we want to compare word vectors from different time periods, vectors should be aligned in same coordinate axes.
After aligning the embeddings for individual time periods, we can use the aligned word vectors to compute the semantic displacement that a word has undergone during a certain time-period.
    \item Third one by Laurino et al. investigates the impact of the COVID-19 pandemic on word meaning.
One key finding from their work is that words directly related to the pandemic exhibited a greater difference in semantic similarity between pre-pandemic and pandemic time periods.
This suggests that these words underwent a more significant and rapid semantic shift compared to control words not associated with the pandemic.
They also employ semantic similarity analysis to quantify the shifts in meaning for pandemic-related words and provides evidence that the COVID-19 pandemic acted as a catalyst for rapid semantic change.

\end{itemize}

\newpage
\section{Research Papers} \label{sec:paper}
\subsection{Statistically Significant Detection of Linguistic Change}\label{subsec:statistically-significant-detection-of-linguistic-change}

\subsubsection{About}
This research aims to quantify linguistic shifts in word meaning and usage across time, focusing on semantic or contextual changes.
The study utilizes diverse datasets, including Twitter posts, Amazon product reviews, and Google Book Ngrams, to construct time series for individual words.
Three distinct methods are employed for statistical modeling:
\begin{itemize}
    \item frequency analysis to capture sudden changes in word usage,
    \item syntactical analysis examining each word's part-of-speech tag distribution, and
    \item distributional analysis investigating contextual cues through word co-occurrence statistics.
\end{itemize}
To determine the statistical significance of word changes over time, the authors are developing a change point detection algorithm.
Additionally, the research incorporates time series data representing the percentage of search queries according to Google Trends, providing further insight into evolving language patterns and word usage shifts.

\subsubsection{Research Questions}
\begin{enumerate}
\item What methods can quantify the statistical relevance of observed changes in a word's usage across different time periods?
\item Given that a word's usage has changed, how can the precise moment or period of this shift be determined?
\end{enumerate}

\subsubsection{Methodology}

\subsubsection{Results}

\subsection{Diachronic Word Embeddings Reveal Statistical Laws of Semantic Change}\label{subsec:diachronic-word-embeddings-reveal-statistical-laws-of-semantic-change}
\subsubsection{About}
\subsubsection{Research Questions}
\subsubsection{Methodology}
\subsubsection{Results}

\subsection{Specious Sites: Tracking the Spread and Sway of Spurious News Stories at Scale}\label{subsec:specious-sites:-tracking-the-spread-and-sway-of-spurious-news-stories-at-scale}
\subsubsection{About}
\subsubsection{Research Questions}
\subsubsection{Methodology}
\subsubsection{Results}

\bibliographystyle{plain}
\bibliography{main}

\end{document}

### 2024-07-11 src/main.tex
changed to ACM format

@@
-\documentclass[12pt]{article}
-\usepackage{graphicx}
+\documentclass[manuscript, screen, acmsmall]{acmart}
+\usepackage[noend]{algorithm2e}
+\usepackage{graphicx,balance,multirow,xspace,subcaption,longtable}
+\usepackage{tabularx,multirow,booktabs,blindtext}
+\usepackage{makecell,amsmath,cleveref,rotating}
+\usepackage[font={small}, textfont=md]{caption}
 \usepackage[utf8]{inputenc}
+\usepackage[draft,inline,nomargin,index]{fixme}
+\usepackage{marginnote}
+\usepackage[bb=boondox,bbscaled=.95,cal=boondoxo]{mathalfa}
+\usepackage{adjustbox}
+\usepackage{colortbl}
+
+\settopmatter{printacmref=false}
+\setcopyright{none}
+
+\AtBeginDocument{%
+  \providecommand\BibTeX{{%
+    Bib\TeX}}}
+
+\PassOptionsToPackage{hyphens}{url}
+\usepackage{hyperref} \urlstyle{tt}
+\newcolumntype{P}[1]{>{\arraybackslash}p{#1}}
+%\newcolumntype{X}[1]{>{\centering\arraybackslash}p{#1}}
+
+\crefformat{section}{\S#2#1#3}
+\crefformat{subsection}{\S#2#1#3}
+\crefformat{subsubsection}{\S#2#1#3}
+
+\hypersetup{%
+  pdftitle={Report for Qualifying Exam},
+  pdfauthor={},
+  pdfkeywords={},
+  bookmarksnumbered,
+  pdfstartview={FitH},
+  colorlinks,
+  citecolor=black,
+  filecolor=black,
+  linkcolor=black,
+  urlcolor=linkColor,
+  breaklinks=true,
+  hypertexnames=false
+}
+
+\clubpenalty=10000
+\widowpenalty=10000
+
+\newcommand{\red}[1]{\textcolor{red}{\bf #1}}
+\newcommand{\blue}[1]{\textcolor{blue}{\bf #1}}
+
+\newenvironment{packed_itemize}{
+\begin{itemize}
+  \setlength{\itemsep}{2pt}
+  \setlength{\parskip}{0pt}
+  \setlength{\parsep}{0pt}
+  \setlength{\topsep}{2pt}
+}{\end{itemize}}
+
+\newenvironment{packed_enumerate}{
+\begin{enumerate}
+  \setlength{\itemsep}{2pt}
+  \setlength{\parskip}{0pt}
+  \setlength{\parsep}{0pt}
+  \setlength{\topsep}{2pt}
+}{\end{enumerate}}
+
+% \renewcommand{\footnotesize}{\scriptsize}
+\newcommand{\etc}{etc.}
+\newcommand{\eg}{e.g.,\ }
+\newcommand{\etal}{et al.\xspace}
+\newcommand{\ie}{i.e.,\ }
+\newcommand{\re}{r.e.\ }
+\newcommand{\aka}{a.k.a.\ }
+\newcommand{\cf}{\emph{cf.\ }}
+\newcommand{\note}[1]{\textcolor{blue}{#1}}
+\newcommand{\superem}[1]{\underline{\emph{#1}}}
+\newcommand{\para}[1]{\vspace{.05in}\noindent\textbf{#1}}
+\newcommand{\paraem}[1]{\vspace{.05in}\noindent\emph{#1}}
+\newcommand{\yt}{YouTube\xspace}
+\newcommand{\X}{$\mathbb{X}$\xspace}
+\newcommand\mycommfont[1]{\footnotesize\ttfamily\textcolor{blue}{#1}}
 
@@
 
-\begin{titlepage}
+\renewcommand\footnotetextcopyrightpermission[1]{} % removes footnote with conference information in first column
+\pagestyle{plain} % removes running headers
 
-\newcommand{\HRule}{\rule{\linewidth}{0.5mm}}
-\center
+\renewcommand{\sectionautorefname}{\S}
+\renewcommand{\subsectionautorefname}{\S}
+\renewcommand{\subsubsectionautorefname}{\S}
 
-%----------------------------------------------------------------------------------------
-%	HEADING SECTIONS
-%----------------------------------------------------------------------------------------
-
-\textsc{\LARGE The University of Iowa}\\[1.5cm]
-\includegraphics[scale=.1]{figures/iowa}\\[1cm]
-\textsc{\Large Report for Qualifying Exam}\\[0.5cm]
-%\textsc{\large Ph.D.}\\[0.5cm]
-\title {Report for Qualifying Exam}
-\HRule \\[0.4cm]
-{ \huge \bfseries Understanding the flow of Temporal Context Shift in Data Voids}\\[0.4cm]
-\HRule \\[1.5cm]
-
-\emph{Author:} \\
-Manisha \textsc{Keim} \\
-manisha-keim@uiowa.edu \\[0.5cm]
-
-\emph{Advisor:}\\
-Rishab \textsc{Nithyanand}\\
-rishab-nithyanand@uiowa.edu \\[0.5cm]
-
-Department of Computer Science
-
-{\today}\\[2cm]
-
-\end{titlepage}
-
+\input{sections/titlepage}
 %%%%%%%%%%%%%%%%%%%%%%%%%%% Table of Contents
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%% Main document
-\section{Abstract} \label{sec:abstract}
-    In the digital age, online information seekers frequently encounter search results laden with misinformation, biased narratives, and conspiracy theories.
-    This exposure can potentially lead users to accept and propagate false information.
-    However, this phenomenon is not universal across all search queries.
-    Certain searches, known as \textit{data voids} are particularly susceptible to these issues.
-    \\
-    \textit{Data voids} occur when searching for terms that yield limited, non-existent, or highly problematic relevant information.
-    Unlike common searches that produce abundant data, queries falling into data voids may return no results or present irrelevant and often inaccurate information.
-    Data voids can be categorized into two main types: newly coined terms with no established information base, and existing terms whose meanings have evolved over time.
-    This report focuses on the latter, examining data voids where context drift occurs.
-    The following papers has been selected to explore this phenomenon.
-    \begin{itemize}
-        \item Statistically Significant Detection of Linguistic Change. ACM WWW 2014 \cite{kulkarni2014statistically}
-        \item Diachronic Word Embeddings Reveal Statistical Laws of Semantic Change. ACL 2016 \cite{hamilton2018diachronic}
-        \item Specious Sites: Tracking the Spread and Sway of Spurious News Stories at Scale. IEEE S\&P 2024 \cite{hanley2024specious}
-    \end{itemize}
-
-\section{Introduction} \label{sec:introduction}
-\subsection{What is a Data Void?}
-\subsection{Example of a data void?(Migrant Caravan)}
-\subsection{Lifecycle of data void (from long tail to spike)}
-\subsection{Types of Data Void}
-\subsection{Technical Challenge}
-analyzing data void involves context drift over historical time. (How context drift happens with data void.)
+%\maketitle
+%\sloppy
+\input{sections/abstract}
+\input{sections/introduction}
+\newpage
+\input{sections/relatedwork}
+\newpage
+\input{sections/papers}
+\balance
 
 \newpage
-\section{Related Work} \label{sec:relatedwork}
-The evolution of word meanings over time has been a subject of significant interest.
-Researchers have employed a diverse array of methodologies, including word embeddings, neural language models, and diachronic analysis, to track and examine semantic shifts across historical periods.
-Much of this research has utilized the Google Books Ngram corpus, a vast collection spanning from 1900 to 2009, encompassing over 500 billion books in seven languages.
-This dataset provides n-grams with corresponding yearly occurrences and frequencies.
-
-Diverse categories of semantic shifts exist, such as semantic, syntactic, or as Hamilton et al. (2016) noted, can be driven by cultural or linguistic factors.
-Certain studies concentrate on semantic changes, while others explore the broader evolution that words undergo, with some works focus on transformations in word usage and meaning.
-
-\vspace{5mm}
-Gulordava and Baroni (2011) introduce a novel method for finding semantic changes in words over time.
-They focus on detecting shifts between the 1960s and 1990s using the Google Books Ngram corpus.
-Their approach relies on comparing the similarity of a word's surrounding words in these two periods.
-They define context by considering the words that appear within two positions (a window size of 2) before and after the target word, essentially analyzing bigram statistics.
-%Example: \'parent\' and \'sleep\'
-
-Kim et al. (2014) propose a distinct approach that leverages the power of neural language models.
-By analyzing how the model represents words at different points in time, they can capture how word meaning evolves.
-Unlike previous methods that focus on individual words or word frequency, this technique allows them to pinpoint not only words with shifting meanings but also the specific years when those changes occurred.
-Their approach goes beyond simple co-occurrence by using a window size of four and comparing the cosine similarity between the same word's vector representations across different time periods.
-This enables them to detect even subtle changes in how words are used.
-Additionally, by plotting the time series of a word's distance to its neighboring words in the model's vector space, they can visualize the exact period during which the semantic shift occurred.
-They identified words like {\lq{cell}\rq} and {\lq{gay}\rq} that have undergone significant semantic shifts over the analyzed period.
-
-Kulkarni et al. (2015) introduced method for significant detection of linguistic change, which laid out the foundation that utilized word embeddings for detecting semantic shifts.
-Building on the concept of distributed representations introduced by Hinton, they map symbolic data (words) into a continuous vector space where words with similar meanings are positioned close together.
-Their approach incorporates three methods to analyze these embeddings: frequency-based, syntactic, and distributional.
-
-Furthering the research on word embeddings, Liao and Cheng (2016) compared the use of word embeddings with a clustering algorithm called DBSCAN to analyze semantic change.
-Their work investigated which method was more effective for identifying these semantic shifts.
-
-Hamilton et al. (2016) address the challenge of distinguishing between cultural shifts and linguistic drift, both of which can contribute to semantic change.
-They propose two distinct computational measures based on word embeddings.
-One measure captures global semantic shifts, while the other, captures localized changes that reflect cultural influences.
-By analyzing these two measures together, researchers can know the causes behind semantic change.
-
-Expanding on this concept, their subsequent work titled "\textit{Diachronic Word Embeddings Reveal Statistical Laws of Semantic Change}" (2016) investigates recurring patterns in semantic change across different languages.
-They identify two key principles: \textit{the law of conformity}, which suggests that a word's frequency is inversely proportional to the rate of its semantic change.
-In other words, words that are used more frequently tend to experience slower shifts in meaning.
-\textit{The law of innovation}, on the other hand, proposes that words with multiple meanings are more likely to undergo semantic transformations over time.
-
-Furthermore, Kutuzov et al. (2018) conducted a survey on diachronic word embeddings and semantic shifts, consolidating the existing academic research in this domain.
-Their work provides a comprehensive overview of the methodologies and findings related to tracking semantic changes over time using computational techniques.
-
-Giulianelli (2020) explores the computational analysis of lexical semantic change using contextualized word representations.
-Contextualized word representations, such as those derived from models like BERT or ELMo takes into account the context in which words are used.
-This enables a more detailed analysis of how semantic changes occur based on specific word contexts.
-
-The impact of significant events like the COVID-19 pandemic on language and semantic change has also been a subject of study.
-Laurino et al.
-(2023) explores tracking fast semantic changes through a large-scale word association task, aiming to understand how the collective mental lexicon evolves in response to such global events.
-Their research highlights the dynamic nature of language and how it incorporates new senses.
-For instance, words like "quarantine," "mask," and "social distancing" took on new and prominent meanings in everyday conversation.
-
- \vspace{5mm}
-%Among the reviewed literature, three studies are particularly pertinent to our analysis of data voids.
-%In case of data voids, we want to find shifts from it's previous usage to new usage.
-Several studies from our literature review provide valuable insights for addressing data voids in our analysis.
-\begin{itemize}
-    \item The first being the work by Kulkarni et al. which introduces us to time series construction.
-Their distributional methods focuses on finding subtle semantic shits to determine the context where a word in used.
-This concept aligns perfectly with our goal of uncovering shifts in word usage when encountering data voids.
-    \item Second paper by Hamilton et al. introduces us to word emebeddings alignment.
-Since we want to compare word vectors from different time periods, vectors should be aligned in same coordinate axes.
-After aligning the embeddings for individual time periods, we can use the aligned word vectors to compute the semantic displacement that a word has undergone during a certain time-period.
-    \item Third one by Laurino et al. investigates the impact of the COVID-19 pandemic on word meaning.
-One key finding from their work is that words directly related to the pandemic exhibited a greater difference in semantic similarity between pre-pandemic and pandemic time periods.
-This suggests that these words underwent a more significant and rapid semantic shift compared to control words not associated with the pandemic.
-They also employ semantic similarity analysis to quantify the shifts in meaning for pandemic-related words and provides evidence that the COVID-19 pandemic acted as a catalyst for rapid semantic change.
-
-\end{itemize}
-
-\newpage
-\section{Research Papers} \label{sec:paper}
-\subsection{Statistically Significant Detection of Linguistic Change}\label{subsec:statistically-significant-detection-of-linguistic-change}
-
-\subsubsection{About}
-This research aims to quantify linguistic shifts in word meaning and usage across time, focusing on semantic or contextual changes.
-The study utilizes diverse datasets, including Twitter posts, Amazon product reviews, and Google Book Ngrams, to construct time series for individual words.
-Three distinct methods are employed for statistical modeling:
-\begin{itemize}
-    \item frequency analysis to capture sudden changes in word usage,
-    \item syntactical analysis examining each word's part-of-speech tag distribution, and
-    \item distributional analysis investigating contextual cues through word co-occurrence statistics.
-\end{itemize}
-To determine the statistical significance of word changes over time, the authors are developing a change point detection algorithm.
-Additionally, the research incorporates time series data representing the percentage of search queries according to Google Trends, providing further insight into evolving language patterns and word usage shifts.
-
-\subsubsection{Research Questions}
-\begin{enumerate}
-\item What methods can quantify the statistical relevance of observed changes in a word's usage across different time periods?
-\item Given that a word's usage has changed, how can the precise moment or period of this shift be determined?
-\end{enumerate}
-
-\subsubsection{Methodology}
-
-\subsubsection{Results}
-
-\subsection{Diachronic Word Embeddings Reveal Statistical Laws of Semantic Change}\label{subsec:diachronic-word-embeddings-reveal-statistical-laws-of-semantic-change}
-\subsubsection{About}
-\subsubsection{Research Questions}
-\subsubsection{Methodology}
-\subsubsection{Results}
-
-\subsection{Specious Sites: Tracking the Spread and Sway of Spurious News Stories at Scale}\label{subsec:specious-sites:-tracking-the-spread-and-sway-of-spurious-news-stories-at-scale}
-\subsubsection{About}
-\subsubsection{Research Questions}
-\subsubsection{Methodology}
-\subsubsection{Results}
-
-\bibliographystyle{plain}
+\bibliographystyle{ACM-Reference-Format}
 \bibliography{main}

### 2024-07-11 src/sections/abstract.tex
changed to ACM format

%! Author = mkeim
%! Date = 7/11/24

% Document
\section{Abstract} \label{sec:abstract}
    In the digital age, online information seekers frequently encounter search results laden with misinformation, biased narratives, and conspiracy theories.
    This exposure can potentially lead users to accept and propagate false information.
    However, this phenomenon is not universal across all search queries.
    Certain searches, known as \textit{data voids} are particularly susceptible to these issues.
    \\
    \textit{Data voids} occur when searching for terms that yield limited, non-existent, or highly problematic relevant information.
    Unlike common searches that produce abundant data, queries falling into data voids may return no results or present irrelevant and often inaccurate information.
    Data voids can be categorized into two main types: newly coined terms with no established information base, and existing terms whose meanings have evolved over time.
    This report focuses on the latter, examining data voids where context drift occurs.
    The following papers has been selected to explore this phenomenon.
    \begin{itemize}
        \item Statistically Significant Detection of Linguistic Change. ACM WWW 2014 \cite{kulkarni2014statisticallysignificantdetectionlinguistic}
        \item Diachronic Word Embeddings Reveal Statistical Laws of Semantic Change. ACL 2016 \cite{hamilton-etal-2016-diachronic}
        \item The Pandemic in Words: Tracking Fast Semantic Changes via a Large-Scale Word Association Task. \cite{10.1162/opmi_a_00081}
    \end{itemize}

### 2024-07-11 src/sections/introduction.tex
changed to ACM format

%! Author = mkeim
%! Date = 7/11/24

\section{Introduction} \label{sec:introduction}
\subsection{What is a Data Void?}
\subsection{Example of a data void?(Migrant Caravan)}
\subsection{Lifecycle of data void (from long tail to spike)}
\subsection{Types of Data Void}
\subsection{Technical Challenge}
analyzing data void involves context drift over historical time. (How context drift happens with data void.)

### 2024-07-11 src/sections/papers.tex
changed to ACM format

%! Author = mkeim
%! Date = 7/11/24

\section{Research Papers} \label{sec:paper}
\subsection{Statistically Significant Detection of Linguistic Change}\label{subsec:statistically-significant-detection-of-linguistic-change}

\subsubsection{About}
This research aims to quantify linguistic shifts in word meaning and usage across time, focusing on semantic or contextual changes.
The study utilizes diverse datasets, including Twitter posts, Amazon product reviews, and Google Book Ngrams, to construct time series for individual words.
Three distinct methods are employed for statistical modeling:
\begin{itemize}
    \item frequency analysis to capture sudden changes in word usage,
    \item syntactical analysis examining each word's part-of-speech tag distribution, and
    \item distributional analysis investigating contextual cues through word co-occurrence statistics.
\end{itemize}
To determine the statistical significance of word changes over time, the authors are developing a change point detection algorithm.
Additionally, the research incorporates time series data representing the percentage of search queries according to Google Trends, providing further insight into evolving language patterns and word usage shifts.

\subsubsection{Research Questions}
\begin{enumerate}
\item What methods can quantify the statistical relevance of observed changes in a word's usage across different time periods?
\item Given that a word's usage has changed, how can the precise moment or period of this shift be determined?
\end{enumerate}

\subsubsection{Methodology}

\subsubsection{Results}

\subsection{Diachronic Word Embeddings Reveal Statistical Laws of Semantic Change}\label{subsec:diachronic-word-embeddings-reveal-statistical-laws-of-semantic-change}
\subsubsection{About}
\subsubsection{Research Questions}
\subsubsection{Methodology}
\subsubsection{Results}

\subsection{Specious Sites: Tracking the Spread and Sway of Spurious News Stories at Scale}\label{subsec:specious-sites:-tracking-the-spread-and-sway-of-spurious-news-stories-at-scale}
\subsubsection{About}
\subsubsection{Research Questions}
\subsubsection{Methodology}
\subsubsection{Results}

### 2024-07-11 src/sections/relatedwork.tex
changed to ACM format

%! Author = mkeim
%! Date = 7/11/24

\section{Related Work} \label{sec:relatedwork}
The evolution of word meanings over time has been a subject of significant interest.
Researchers have employed a diverse array of methodologies, including word embeddings, neural language models, and diachronic analysis, to track and examine semantic shifts across historical periods.
Much of this research has utilized the Google Books Ngram corpus, a vast collection spanning from 1900 to 2009, encompassing over 500 billion books in seven languages.
This dataset provides n-grams with corresponding yearly occurrences and frequencies.

Diverse categories of semantic shifts exist, such as semantic, syntactic, or as Hamilton et al. (2016) noted, can be driven by cultural or linguistic factors.
Certain studies concentrate on semantic changes, while others explore the broader evolution that words undergo, with some works focus on transformations in word usage and meaning.

\vspace{5mm}
Gulordava and Baroni (2011) introduce a novel method for finding semantic changes in words over time.
They focus on detecting shifts between the 1960s and 1990s using the Google Books Ngram corpus.
Their approach relies on comparing the similarity of a word's surrounding words in these two periods.
They define context by considering the words that appear within two positions (a window size of 2) before and after the target word, essentially analyzing bigram statistics.
%Example: \'parent\' and \'sleep\'

Kim et al. (2014) propose a distinct approach that leverages the power of neural language models.
By analyzing how the model represents words at different points in time, they can capture how word meaning evolves.
Unlike previous methods that focus on individual words or word frequency, this technique allows them to pinpoint not only words with shifting meanings but also the specific years when those changes occurred.
Their approach goes beyond simple co-occurrence by using a window size of four and comparing the cosine similarity between the same word's vector representations across different time periods.
This enables them to detect even subtle changes in how words are used.
Additionally, by plotting the time series of a word's distance to its neighboring words in the model's vector space, they can visualize the exact period during which the semantic shift occurred.
They identified words like {\lq{cell}\rq} and {\lq{gay}\rq} that have undergone significant semantic shifts over the analyzed period.

Kulkarni et al. (2015) introduced method for significant detection of linguistic change, which laid out the foundation that utilized word embeddings for detecting semantic shifts.
Building on the concept of distributed representations introduced by Hinton, they map symbolic data (words) into a continuous vector space where words with similar meanings are positioned close together.
Their approach incorporates three methods to analyze these embeddings: frequency-based, syntactic, and distributional.

Furthering the research on word embeddings, Liao and Cheng (2016) compared the use of word embeddings with a clustering algorithm called DBSCAN to analyze semantic change.
Their work investigated which method was more effective for identifying these semantic shifts.

Hamilton et al. (2016) address the challenge of distinguishing between cultural shifts and linguistic drift, both of which can contribute to semantic change.
They propose two distinct computational measures based on word embeddings.
One measure captures global semantic shifts, while the other, captures localized changes that reflect cultural influences.
By analyzing these two measures together, researchers can know the causes behind semantic change.

Expanding on this concept, their subsequent work titled "\textit{Diachronic Word Embeddings Reveal Statistical Laws of Semantic Change}" (2016) investigates recurring patterns in semantic change across different languages.
They identify two key principles: \textit{the law of conformity}, which suggests that a word's frequency is inversely proportional to the rate of its semantic change.
In other words, words that are used more frequently tend to experience slower shifts in meaning.
\textit{The law of innovation}, on the other hand, proposes that words with multiple meanings are more likely to undergo semantic transformations over time.

Furthermore, Kutuzov et al. (2018) conducted a survey on diachronic word embeddings and semantic shifts, consolidating the existing academic research in this domain.
Their work provides a comprehensive overview of the methodologies and findings related to tracking semantic changes over time using computational techniques.

Giulianelli (2020) explores the computational analysis of lexical semantic change using contextualized word representations.
Contextualized word representations, such as those derived from models like BERT or ELMo takes into account the context in which words are used.
This enables a more detailed analysis of how semantic changes occur based on specific word contexts.

The impact of significant events like the COVID-19 pandemic on language and semantic change has also been a subject of study.
Laurino et al. (2023) explores tracking fast semantic changes through a large-scale word association task, aiming to understand how the collective mental lexicon evolves in response to such global events.
Their research highlights the dynamic nature of language and how it incorporates new senses.
For instance, words like "quarantine," "mask," and "social distancing" took on new and prominent meanings in everyday conversation.

 \vspace{5mm}
%Among the reviewed literature, three studies are particularly pertinent to our analysis of data voids.
%In case of data voids, we want to find shifts from it's previous usage to new usage.
Several studies from our literature review provide valuable insights for addressing data voids in our analysis.
\begin{itemize}
    \item The first being the work by Kulkarni et al. which introduces us to time series construction.
Their distributional methods focuses on finding subtle semantic shits to determine the context where a word in used.
This concept aligns perfectly with our goal of uncovering shifts in word usage when encountering data voids.
    \item Second paper by Hamilton et al. introduces us to word emebeddings alignment.
Since we want to compare word vectors from different time periods, vectors should be aligned in same coordinate axes.
After aligning the embeddings for individual time periods, we can use the aligned word vectors to compute the semantic displacement that a word has undergone during a certain time-period.
    \item Third one by Laurino et al. investigates the impact of the COVID-19 pandemic on word meaning.
One key finding from their work is that words directly related to the pandemic exhibited a greater difference in semantic similarity between pre-pandemic and pandemic time periods.
This suggests that these words underwent a more significant and rapid semantic shift compared to control words not associated with the pandemic.
They also employ semantic similarity analysis to quantify the shifts in meaning for pandemic-related words and provides evidence that the COVID-19 pandemic acted as a catalyst for rapid semantic change.

\end{itemize}

### 2024-07-17 src/sections/relatedwork.tex
completed related work section w/o table (due review)

@@
 \section{Related Work} \label{sec:relatedwork}
+\para{Tracking how word meaning change over time.}
 The evolution of word meanings over time has been a subject of significant interest.
-Researchers have employed a diverse array of methodologies, including word embeddings, neural language models, and diachronic analysis, to track and examine semantic shifts across historical periods.
-Much of this research has utilized the Google Books Ngram corpus, a vast collection spanning from 1900 to 2009, encompassing over 500 billion books in seven languages.
+Researchers employed a diverse array of methodologies, including \emph{word embeddings} ~\cite{kulkarni2014statisticallysignificantdetectionlinguistic},
+\emph{neural language models} ~\cite{kim-etal-2014-temporal}, and \emph{diachronic analysis} ~\cite{hamilton-etal-2016-diachronic, kutuzov-etal-2018-diachronic},
+to track and examine semantic shifts across historical periods.
+Much of this research utilized the \emph{Google Books Ngram corpus}
+~\cite{gulordava-baroni-2011-distributional, kim-etal-2014-temporal, kulkarni2014statisticallysignificantdetectionlinguistic, 10.1007/978-3-319-50496-4_18, hamilton-etal-2016-cultural, hamilton-etal-2016-diachronic, kutuzov-etal-2018-diachronic},
+a vast collection spanning from 1900 to 2009,
+encompassing over 500 billion books in seven languages.
 This dataset provides n-grams with corresponding yearly occurrences and frequencies.
 
-Diverse categories of semantic shifts exist, such as semantic, syntactic, or as Hamilton et al. (2016) noted, can be driven by cultural or linguistic factors.
-Certain studies concentrate on semantic changes, while others explore the broader evolution that words undergo, with some works focus on transformations in word usage and meaning.
+\para{Various type of linguistic changes.}
+Words have the ability to undergo shifts in meaning over the course of time as a result of various linguistic mechanisms,
+including but not limited to \emph{semantic drift} ~\cite{gulordava-baroni-2011-distributional, kulkarni2014statisticallysignificantdetectionlinguistic},
+\emph{syntactic alterations} ~\cite{kulkarni2014statisticallysignificantdetectionlinguistic, hamilton-etal-2016-cultural},
+\emph{broadening} as opposed to \emph{narrowing} of meaning, and the processes of \emph{amelioration} versus \emph{pejoration}.
+These transformations, as highlighted by Hamilton et al.\ (2016), could be driven by \emph{cultural} or \emph{linguistic} factors.
+Certain studies concentrate on semantic changes, while others explored the broader evolution that words underwent ~\cite{kim-etal-2014-temporal}, with some works focus on transformations in word usage and meaning ~\cite{gulordava-baroni-2011-distributional, kulkarni2014statisticallysignificantdetectionlinguistic, hamilton-etal-2016-cultural}.
+The causes of linguistic changes and the methodologies employed to identify them are outlined in \Cref{tab:related-work}.
 
-\vspace{5mm}
-Gulordava and Baroni (2011) introduce a novel method for finding semantic changes in words over time.
+\para{Semantic Change.}
+Semantic change refers to any change in the meaning(s) of a word over time or acquiring a new sense ~\cite{gulordava-baroni-2011-distributional, 10.1162/opmi_a_00081}.
+Kutuzov et al.\ (2018) conducted a survey on semantic shifts, consolidating the existing academic research in this domain.
+Their work provides a comprehensive overview of the methodologies and findings related to tracking semantic changes over time using computational techniques.
+\vspace{0mm}
+\begin{itemize}
+    \item \textbf{Co-occurrence Counts:} Gulordava and Baroni (2011) quantify how frequently words appear within the same context.
         They focus on detecting shifts between the 1960s and 1990s using the Google Books Ngram corpus.
         Their approach relies on comparing the similarity of a word's surrounding words in these two periods.
-They define context by considering the words that appear within two positions (a window size of 2) before and after the target word, essentially analyzing bigram statistics.
+        They define context by considering the words that appear within two positions (a window size of 2) before and after the target word.
         %Example: \'parent\' and \'sleep\'
-
-Kim et al. (2014) propose a distinct approach that leverages the power of neural language models.
-By analyzing how the model represents words at different points in time, they can capture how word meaning evolves.
-Unlike previous methods that focus on individual words or word frequency, this technique allows them to pinpoint not only words with shifting meanings but also the specific years when those changes occurred.
-Their approach goes beyond simple co-occurrence by using a window size of four and comparing the cosine similarity between the same word's vector representations across different time periods.
-This enables them to detect even subtle changes in how words are used.
-Additionally, by plotting the time series of a word's distance to its neighboring words in the model's vector space, they can visualize the exact period during which the semantic shift occurred.
-They identified words like {\lq{cell}\rq} and {\lq{gay}\rq} that have undergone significant semantic shifts over the analyzed period.
-
-Kulkarni et al. (2015) introduced method for significant detection of linguistic change, which laid out the foundation that utilized word embeddings for detecting semantic shifts.
-Building on the concept of distributed representations introduced by Hinton, they map symbolic data (words) into a continuous vector space where words with similar meanings are positioned close together.
-Their approach incorporates three methods to analyze these embeddings: frequency-based, syntactic, and distributional.
-
-Furthering the research on word embeddings, Liao and Cheng (2016) compared the use of word embeddings with a clustering algorithm called DBSCAN to analyze semantic change.
+    \item \para{Neural Language Models:} Kim et al.\ (2014) suggest a new method using \emph{neural language models} to track word meaning evolution.
+        They analyze word vectors over time to capture changes.
+        Their method compares word vectors using a window size of four from 1900--2009 to detect subtle changes in word usage.
+        Additionally, by plotting the time series of a word’s distance to its neighboring words in the model’s vector space, they can visualize the exact period during which the semantic shift occurred.
+        Words like `cell' and `gay' show significant changes over time.
+    \item \para{Word Embeddings:} Kulkarni et al.\ (2015) introduced method for detecting linguistic change with statistical significance that utilized \emph{word embeddings}.
+        Building on the concept of distributed representations proposed by Hinton, they map symbolic data (words) into a continuous vector space where words with similar meanings are positioned close together.
+        Their approach incorporates three methods to analyze these embeddings: \emph{frequency-based, syntactic, and distributional}, which are later discussed in  detail in next section.(\Cref{sec:paper})
+    \item \para{Clustering:} Liao and Cheng (2016) compared the use of word embeddings with a clustering algorithm called DBSCAN to analyze semantic change.
         Their work investigated which method was more effective for identifying these semantic shifts.
-
-Hamilton et al. (2016) address the challenge of distinguishing between cultural shifts and linguistic drift, both of which can contribute to semantic change.
+    \item \para{Cultural vs. Linguistic:} Hamilton et al.\ (2016) address the challenge of distinguishing between cultural shifts and linguistic drift, both of which can contribute to semantic change.
         They propose two distinct computational measures based on word embeddings.
         One measure captures global semantic shifts, while the other, captures localized changes that reflect cultural influences.
-By analyzing these two measures together, researchers can know the causes behind semantic change.
+        Cultural changes like new technologies are related to local neighbourhood change.
+        Examples: gay, virus, cell
+        These words gained new meanings due to uses in community-specific vernacular (gay) or technological advances (virus, cell).
+    \item \para{Diachronic Word Embeddings: } Expanding on the above concepts, their subsequent work investigates recurring patterns in semantic change across different languages.
+        They propose two key principles: \emph{the law of conformity}, which suggests that a word's frequency is inversely proportional to the rate of its semantic change.
+        In other words, words that are used more frequently tend to experience slower shifts in meaning.
+        \emph{The law of innovation}, on the other hand, proposes that words with multiple meanings are more likely to undergo semantic transformations over time.
+    \item \para{BERT:} Giulianelli (2020) explores the computational analysis of lexical semantic change using contextualized word representations.
+        Contextualized word representations, such as those derived from models like BERT takes into account the context in which words are used.
+        This enables a more detailed analysis of how semantic changes occur based on specific word contexts.
+    \item \para{Semantic Similarity:} The impact of significant events like the COVID-19 pandemic on language and semantic change has also been a subject of study.
+        Laurino et al.\ (2023) explores tracking fast semantic changes through a large-scale word association task, aiming to understand how the collective mental lexicon evolves in response to such global events.
+        Their research highlights the dynamic nature of language and how it incorporates new senses.
+        For instance, words like `quarantine', `mask', and `social distancing' took on new and prominent meanings in everyday conversation.
+\end{itemize}
 
-Expanding on this concept, their subsequent work titled "\textit{Diachronic Word Embeddings Reveal Statistical Laws of Semantic Change}" (2016) investigates recurring patterns in semantic change across different languages.
-They identify two key principles: \textit{the law of conformity}, which suggests that a word's frequency is inversely proportional to the rate of its semantic change.
-In other words, words that are used more frequently tend to experience slower shifts in meaning.
-\textit{The law of innovation}, on the other hand, proposes that words with multiple meanings are more likely to undergo semantic transformations over time.
-
-Furthermore, Kutuzov et al. (2018) conducted a survey on diachronic word embeddings and semantic shifts, consolidating the existing academic research in this domain.
-Their work provides a comprehensive overview of the methodologies and findings related to tracking semantic changes over time using computational techniques.
-
-Giulianelli (2020) explores the computational analysis of lexical semantic change using contextualized word representations.
-Contextualized word representations, such as those derived from models like BERT or ELMo takes into account the context in which words are used.
-This enables a more detailed analysis of how semantic changes occur based on specific word contexts.
-
-The impact of significant events like the COVID-19 pandemic on language and semantic change has also been a subject of study.
-Laurino et al.
-(2023) explores tracking fast semantic changes through a large-scale word association task, aiming to understand how the collective mental lexicon evolves in response to such global events.
-Their research highlights the dynamic nature of language and how it incorporates new senses.
-For instance, words like "quarantine," "mask," and "social distancing" took on new and prominent meanings in everyday conversation.
+\para{Syntactic Change.}
+The syntactic functionality of a word can evolve by transitioning into a new part-of-speech (POS) category.
+Nouns, due to their inherent flexibility in meaning, exhibit a greater tendency to undergo these changes driven by cultural shifts.
+While are more likely to participate in gradual semantic changes that follow established linguistic patterns.
+\vspace{0mm}
+\begin{itemize}
+    \item \textbf{Acquiring a new POS:}
+    Kulkarni et al.\ assigned part-of-speech (POS) tags to a large collection of text (corpus) and then calculated how likely different words were to appear in certain grammatical contexts (probability distributions).
+    They used the word `apple' as an example, which transitioned from being used as a common noun (e.g., a fruit) to a proper noun (referring to the Apple company) after the company's rise in the 1980s.
+    \item \para{Nouns vs. Verbs:}
+     Hamilton et al.\ (2016) highlight how cultural changes, often influenced by new technologies, are closely tied to transformations within local neighborhoods, particularly sensitive to shifts in nouns.
+     On the other hand, linguistic changes, exemplified by subjectification, are more associated with global measures and are particularly responsive to variations in verbs.
+     The evolution of words like `actually', `must', and `promise' demonstrate these changes.
+     For instance, `must' has transitioned from expressing obligation to indicating necessity, showcasing a common pattern seen in modal verbs.
+     Similarly, `promise' shows how performative speech acts undergo significant pragmatic and subjectification-related changes over time.
+%    Cultural changes like new technologies are related to local neighbourhood change, which is more sensitive to changes in nouns.
+%    Examples: gay, virus, cell
+%    Linguistic changes like subjectification is related to global measure, which is more sensitive to change in verbs.
+%    Examples: actually, must, promise
+%    actually is a case of subjectification
+%    must shifted from a deontic/obligation usage (“you must do X”) to a epistemic one (“X must be the case”), exemplifying a regular pattern of change common to many modal verb
+%    promise represents the class of shifting “performative speech acts” that undergo rich changes due to their pragmatic uses and subjectification
+    \item \para{Acquiring a new sense:}
+    Words are complex entities with multiple senses that can evolve over time while their form remains constant, as discussed by (Laurino et al., 2023).
+    Their research studies the overall mental lexicon of words, focusing on quantifying the drift of word concepts.
+    The study highlights how most words possess a range of meanings that can be added, removed, or modified across different temporal contexts.
+\end{itemize}
 
@@
 \begin{itemize}
-    \item The first being the work by Kulkarni et al. which introduces us to time series construction.
+    \item The first being the work by Kulkarni et al.\ which introduces us to time series construction.
 Their distributional methods focuses on finding subtle semantic shits to determine the context where a word in used.
 This concept aligns perfectly with our goal of uncovering shifts in word usage when encountering data voids.
-    \item Second paper by Hamilton et al. introduces us to word emebeddings alignment.
+    \item Second paper by Hamilton et al.\ introduces us to word emebeddings alignment.
 Since we want to compare word vectors from different time periods, vectors should be aligned in same coordinate axes.
 After aligning the embeddings for individual time periods, we can use the aligned word vectors to compute the semantic displacement that a word has undergone during a certain time-period.
-    \item Third one by Laurino et al. investigates the impact of the COVID-19 pandemic on word meaning.
+    \item Third one by Laurino et al.\ investigates the impact of the COVID-19 pandemic on word meaning.
 One key finding from their work is that words directly related to the pandemic exhibited a greater difference in semantic similarity between pre-pandemic and pandemic time periods.

### 2024-07-22 src/sections/paper_hamilton.tex
changes in related work, table pending

%! Author = mkeim
%! Date = 7/18/24

\section{RP2.} \label{sec:paper_hamilton}
\subsection{Diachronic Word Embeddings Reveal Statistical Laws of Semantic Change}

1.	Creating Word Embeddings:
•	They create word embeddings for different time periods using historical text corpora.
2.	Using SVD, PPMI, and SGNS:
•	SVD (Singular Value Decomposition): They apply SVD to the Positive Pointwise Mutual Information (PPMI) matrices to reduce the dimensionality of the word vectors.
•	PPMI (Positive Pointwise Mutual Information): They use PPMI to measure the association between words and their contexts. PPMI values are computed for each time period to capture word-context relationships.
•	SGNS (Skip-Gram with Negative Sampling): They also use the SGNS model to create word embeddings, leveraging its effectiveness in capturing semantic relationships.
3.	Aligning Word Embeddings:
•	They align word embeddings across different time periods to ensure that the same word in different periods is represented in a comparable way. This alignment allows for tracking changes in word meanings over time.
4.	Analyzing Semantic Change:
•	By comparing word embeddings from different time periods, they analyze the semantic shifts. They use cosine similarity to measure changes in word meanings and identify words that have undergone significant semantic shifts.

•	PPMI: They computed PPMI matrices for each time period, capturing the association between words and their contexts.
•	SVD: They applied SVD to the PPMI matrices to reduce the dimensionality of the vectors, making them more manageable and interpretable.
•	SGNS: They used the SGNS model to directly create word embeddings, leveraging its ability to capture semantic relationships in a low-dimensional space.

### 2024-07-22 src/sections/paper_kulkarni.tex
changes in related work, table pending

%! Author = mkeim
%! Date = 7/18/24

\section{RP1: Statistically Significant Detection of Linguistic Change} \label{sec:paper_kulkarni}
\subsection{Introduction} \label{subsec:introduction}
This research aims to quantify linguistic shifts in word meaning and usage across time, focusing on semantic or contextual changes.
To understand data voids, we are sure that a change is happening, but we need to understand the change that is happening over time.
Also after knowing that a change has occurred, we would like to know when did that change really occur so that we can co-relate that the data void occured when a political event happened or how the COVID outbreak event,
which could help us identify that the term were introduced due to that event, which can be a future direction we can look into.
This paper is exciting because the ideas inspired me to pursue this project and get future directions as well.

Their approach focuses on three methods to construct time series for a word.
The first one is based on frequency, which captures sudden changes in word usage.
In our work, this corresponds to strategic new terms data voids or breaking news data void,
where breaking news data void shows that there's comes a sudden change in a word activity whenever's there an event happen.
For example, the term "filmyourhospital" was introduced in during 2020 when coronavirus took place, which corresponds to a sudden spike.
And this term was never used before 2020, making it a new term that introduced to support the conspiracy theory regarding encouraged people to visit local hospitals to take pictures and videos of empty hospitals to help “prove” that the COVID-19 pandemic is an elaborate hoax.

Their second method is Syntactic method where they analyze each word's part of speech tag.

And last method is distributional method which captures overall usage of word using word co-occurence.
This type of method can help us identify data voids which have changed their meaning over time, such as Fragmented Concepts or Outdated Terms DV.

They apply these techniques on three different domains: books, tweets and reviews.

The study utilizes diverse datasets, including Twitter posts, Amazon product reviews, and Google Book Ngrams, to construct time series for individual words.
Three distinct methods are employed for statistical modeling.

\begin{itemize}
    \item \para{RQ1.} \emph{What methods can quantify the statistical relevance of observed changes in a word's usage across different time periods?}
    \item \para{RQ2.} \emph{Given that a word's usage has changed, how can the precise moment or period of this shift be determined?}
\end{itemize}

### 2024-07-22 src/sections/relatedwork.tex
changes in related work, table pending

@@
 \section{Related Work} \label{sec:relatedwork}
-\para{Tracking how word meaning change over time.}
+\para{Tracking how word meanings change over time.}
 The evolution of word meanings over time has been a subject of significant interest.
@@
 
-\para{Various type of linguistic changes.}
+\para{Words can change meaning over time through several linguistic processes.}
 Words have the ability to undergo shifts in meaning over the course of time as a result of various linguistic mechanisms,
-including but not limited to \emph{semantic drift} ~\cite{gulordava-baroni-2011-distributional, kulkarni2014statisticallysignificantdetectionlinguistic},
-\emph{syntactic alterations} ~\cite{kulkarni2014statisticallysignificantdetectionlinguistic, hamilton-etal-2016-cultural},
-\emph{broadening} as opposed to \emph{narrowing} of meaning, and the processes of \emph{amelioration} versus \emph{pejoration}.
-These transformations, as highlighted by Hamilton et al.\ (2016), could be driven by \emph{cultural} or \emph{linguistic} factors.
-Certain studies concentrate on semantic changes, while others explored the broader evolution that words underwent ~\cite{kim-etal-2014-temporal}, with some works focus on transformations in word usage and meaning ~\cite{gulordava-baroni-2011-distributional, kulkarni2014statisticallysignificantdetectionlinguistic, hamilton-etal-2016-cultural}.
-The causes of linguistic changes and the methodologies employed to identify them are outlined in \Cref{tab:related-work}.
+including but not limited to:
+\begin{itemize}
+    \item \emph{semantic drift} ~\cite{gulordava-baroni-2011-distributional, kim-etal-2014-temporal, kulkarni2014statisticallysignificantdetectionlinguistic}
+    \begin{itemize}
+        \item Refers to the evolution of word \underline{\emph{meanings}} over time.
+        \item Example: The word `gay' originally meant `happy' or `carefree', but now it predominantly means `homosexual'.
+    \end{itemize}
+    \item \emph{syntactic alterations} ~\cite{kulkarni2014statisticallysignificantdetectionlinguistic, hamilton-etal-2016-cultural, giulianelli-etal-2020-analysing}
+    \begin{packed_itemize}
+        \item Syntax focuses on the \underline{\emph{structure}} of language.
+        This type of change pertains to modifications in the arrangement of words in sentences.
+        \item Example:  The word `apple', which transitioned from being used as a `common noun' (e.g., a fruit) to a `proper noun' (referring to the Apple company) after the company’s rise in the 1980s.
+    \end{packed_itemize}
+    \item \emph{broadening} ~\cite{kulkarni2014statisticallysignificantdetectionlinguistic, hamilton-etal-2016-cultural, giulianelli-etal-2020-analysing}
+        \begin{itemize}
+            \item A word’s meaning becomes more general than its earlier meaning, also known as \underline{\emph{generalization.}}
+            \item Example: `Holiday' originally meant a religious festival, but now it can refer to any day of celebration or time off.
+        \end{itemize}
+    \item \emph{narrowing} ~\cite{kulkarni2014statisticallysignificantdetectionlinguistic, giulianelli-etal-2020-analysing}
+    \begin{itemize}
+            \item A word’s meaning becomes more specific than its earlier meaning, also known as \underline{\emph{specialization}}.
+            \item Example: `Literally' used to mean `figuratively' or `symbolically'.
+            Now, it is used to emphasize the truthfulness of a statement.
+        \end{itemize}
+    \item \emph{amelioration}
+     \begin{itemize}
+            \item A word takes on a more \underline{\emph{positive}} meaning over time.
+            \item Example: The word `thrifty' once meant `cheap' but now suggests responsible use of resources.
+        \end{itemize}
+    \item \emph{pejoration} ~\cite{10.1162/opmi_a_00081, gulordava-baroni-2011-distributional}
+        \begin{itemize}
+            \item A word takes on a more \underline{\emph{negative}} meaning over time.
+            \item Example: The word `awful' meant `full of awe' which transitioned to `terrible' or `appalling'.
+        \end{itemize}
 
-\para{Semantic Change.}
+\end{itemize}
+
+Hamilton et al.\ (2016) highlighted that \emph{cultural} or \emph{linguistic} factors could have driven these transformations.
+Certain studies concentrated on semantic changes, some worked on syntactic changes, ~\cite{kulkarni2014statisticallysignificantdetectionlinguistic, hamilton-etal-2016-cultural, 10.1162/opmi_a_00081}
+and others explored the broader evolution that words underwent ~\cite{gulordava-baroni-2011-distributional, kim-etal-2014-temporal, kulkarni2014statisticallysignificantdetectionlinguistic, hamilton-etal-2016-diachronic}.
+\Cref{tab:related-work} outlines various types of linguistic changes and the approaches employed to identify them.
+
+\para{Building on the extensive research on \emph{semantic} change.}
 Semantic change refers to any change in the meaning(s) of a word over time or acquiring a new sense ~\cite{gulordava-baroni-2011-distributional, 10.1162/opmi_a_00081}.
@@
 Their work provides a comprehensive overview of the methodologies and findings related to tracking semantic changes over time using computational techniques.
-\vspace{0mm}
+
 \begin{itemize}
-    \item \textbf{Co-occurrence Counts:} Gulordava and Baroni (2011) quantify how frequently words appear within the same context.
-        They focus on detecting shifts between the 1960s and 1990s using the Google Books Ngram corpus.
-        Their approach relies on comparing the similarity of a word's surrounding words in these two periods.
-        They define context by considering the words that appear within two positions (a window size of 2) before and after the target word.
-        %Example: \'parent\' and \'sleep\'
-    \item \para{Neural Language Models:} Kim et al.\ (2014) suggest a new method using \emph{neural language models} to track word meaning evolution.
-        They analyze word vectors over time to capture changes.
-        Their method compares word vectors using a window size of four from 1900--2009 to detect subtle changes in word usage.
-        Additionally, by plotting the time series of a word’s distance to its neighboring words in the model’s vector space, they can visualize the exact period during which the semantic shift occurred.
-        Words like `cell' and `gay' show significant changes over time.
-    \item \para{Word Embeddings:} Kulkarni et al.\ (2015) introduced method for detecting linguistic change with statistical significance that utilized \emph{word embeddings}.
-        Building on the concept of distributed representations proposed by Hinton, they map symbolic data (words) into a continuous vector space where words with similar meanings are positioned close together.
-        Their approach incorporates three methods to analyze these embeddings: \emph{frequency-based, syntactic, and distributional}, which are later discussed in  detail in next section.(\Cref{sec:paper})
-    \item \para{Clustering:} Liao and Cheng (2016) compared the use of word embeddings with a clustering algorithm called DBSCAN to analyze semantic change.
-        Their work investigated which method was more effective for identifying these semantic shifts.
-    \item \para{Cultural vs. Linguistic:} Hamilton et al.\ (2016) address the challenge of distinguishing between cultural shifts and linguistic drift, both of which can contribute to semantic change.
-        They propose two distinct computational measures based on word embeddings.
-        One measure captures global semantic shifts, while the other, captures localized changes that reflect cultural influences.
-        Cultural changes like new technologies are related to local neighbourhood change.
-        Examples: gay, virus, cell
-        These words gained new meanings due to uses in community-specific vernacular (gay) or technological advances (virus, cell).
-    \item \para{Diachronic Word Embeddings: } Expanding on the above concepts, their subsequent work investigates recurring patterns in semantic change across different languages.
-        They propose two key principles: \emph{the law of conformity}, which suggests that a word's frequency is inversely proportional to the rate of its semantic change.
-        In other words, words that are used more frequently tend to experience slower shifts in meaning.
-        \emph{The law of innovation}, on the other hand, proposes that words with multiple meanings are more likely to undergo semantic transformations over time.
-    \item \para{BERT:} Giulianelli (2020) explores the computational analysis of lexical semantic change using contextualized word representations.
-        Contextualized word representations, such as those derived from models like BERT takes into account the context in which words are used.
-        This enables a more detailed analysis of how semantic changes occur based on specific word contexts.
+    \item \textbf{Investigating Semantic Shifts Through Word Contexts.}
+    Gulordava and Baroni (2011) detected semantic change by focusing on words used in the 1960s and 1990s.
+        They compared the similarity of the surrounding words (words that co-occur with the target word) in these two time periods.
+    \begin{itemize}
+        \item To assess similarity, the researchers calculated the \emph{Local Mutual Information (LMI)} score between the central word and its surrounding words.
+        A low LMI score between the target word and its surrounding words across the two time periods indicated a potential semantic shift.
+        \item The study demonstrated that their distributional similarity models were effective in capturing \emph{cultural shifts} in word meaning.
+        For example, they found that the word `sleep' acquired more negative connotations related to sleep disorders when comparing its contexts in the 1960s to those in the 1990s.
+    \end{itemize}
+
+    \item \para{Tracking Meaning Evolution Through Neural Nets.}
+    Kim et al.\ (2014) focused on analyzing how word meanings evolved between 1900 and 2009.
+    \begin{itemize}
+        \item They developed the first method that employed \emph{prediction-based model} to trace semantic shifts.
+        This involved training a model on data from a specific year $y_i$ and then using the resulting word vectors as the starting point for training the model on the next year's data $y_i+1$.
+        \item Their method analyzed \emph{global shifts} in a word's vector semantics.
+        Additionally, by plotting the time series of a word's distance to its neighboring words in the model's vector space, they visualized the period during which the semantic shift occurred.
+        \item They demonstrated this for the word `cell' compared to its early neighbors, `closet' and `dungeon,' and the more recent neighbors, `phone' and `cordless.'
+        For `cell,' the identified period of change (1985-2009), which interestingly coincides with the introduction and widespread adoption of cell phones by the public.
+    \end{itemize}
+
+    \item \para{Pinpointing Significant Shifts Statistically.}
+    Kulkarni et al.\ (2015) proposes a novel computational approach to identify and quantify the semantic and usage changes in words across various media (new products, movies and books).
+    \begin{itemize}
+        \item Building on the concept of distributed representations proposed by Hinton ~\cite{hinton1986learning}, they map words into a continuous vector space where words with similar meanings are positioned close together.
+        \item The approach hinges on constructing \emph{property time series} for each word.
+        They propose three methods for constructing these time series (\Cref{sec:paper_hamilton}).:
+        \begin{itemize}
+            \item \textbf{Frequency:}
+            This method analyzes changes in a word's overall frequency of use, assuming a sudden shift in frequency might indicate a semantic shift.
+            \item \textbf{Syntactic:}
+            This method examines the distribution of a word's part-of-speech tags (e.g., noun, verb) across different time periods, aiming to capture changes in how the word functions grammatically.
+            \item \textbf{Distributional:}
+            This approach leverages word embeddings which are created for each year, and then alignment is done to represent them in joint embedding space,
+            and it's utilized to construct distributional time series for a word's displacement.
+        \end{itemize}
+        \item Finally, they employed statistically sound \emph{change point detection algorithm} to identify significant moments in these time series, pinpointing the periods where word meaning or usage likely underwent a shift.
+        Their results indicated that computational methods for the detection of semantic shifts can be robustly applied to time spans less than a decade.
+    \end{itemize}
+
+    \item \para{Word Semantic Modelling of Polysemant.}
+        Liao and Cheng (2016) explored the semantic changes of words.
+        Their approach built on the understanding that word meaning is closely tied to its context.
+        When a word's meaning changes, the surrounding words used with it \emph{(context words)} are likely to change as well.
+        They focused on \emph{polysemous} words (words with multiple meanings) and aimed to detect when new meanings emerged.
+    \begin{itemize}
+        \item They used the \emph{skip-gram architecture with negative sampling} ~\cite{10.5555/2999792.2999959} to obtain word embeddings.
+        This technique helped in capturing the contextual meaning of words by representing them in a continuous vector space.
+        \item They employed \emph{DBSCAN} to group word embeddings.
+        DBSCAN helped in identifying clusters of similar word contexts and distinguishing them from noise, which could indicate semantic changes.
+        \item To find similar words, they used a nearest neighbor search method called \emph{Random Project Forest}.
+        This method helped in identifying words that are contextually similar to a given word.
+        \item Finally, they compared the stability of \emph{similar words} with the stability of their \emph{context words}.
+    \end{itemize}
+
+    \item \para{Distinguishing Cultural Shifts from Linguistic Drift.}
+    Hamilton et al.\ (2016) addressed the challenge of distinguishing between \emph{cultural shifts} and \emph{linguistic drift}, both of which can contribute to semantic change.
+    \begin{itemize}
+        \item They proposed two distinct measures based on distributional semantics to distinguish between these two types of semantic change:
+        \begin{itemize}
+            \item \para{Local Neighborhood Measure:} This measure focused on the closest neighbors (most similar words) in a word's embedding.
+            A drastic shift in these nearest neighbors suggests a significant change in core meaning, potentially driven by a \emph{cultural shift}
+            (e.g., `gay' changing from carefree to referring to homosexuality).
+            \item \para{Global Measure:} This measure considered the overall distribution of a word's surrounding words in a larger context window.
+            Gradual changes in this broader distribution are more likely to reflect \emph{linguistic drift}, the natural evolution of language due to regular processes
+            (e.g., `promise' expanding from a declaration to also suggesting a likelihood).
+        \end{itemize}
+        \item Prior research often treated semantic change as a single phenomenon.
+        Hamilton et al.\ offered a novel approach by distinguishing between cultural shifts and linguistic drift using their two measures.
+    \end{itemize}
+
+    \item \para{Diachronic Analysis on Historical Data: }
+    The primary goal of Hamilton et al.\ (2016) is to track semantic changes and understand how the meanings of words shift in different historical contexts.
+    \begin{itemize}
+        \item They created word embeddings for different time periods using both the PPMI matrix with SVD and the SGNS model.
+        These embeddings were generated from historical text corpora to capture the contextual usage of words in each period.
+        \item After creating the embeddings, they aligned the embeddings.
+        Once the embeddings are aligned, they calculate the \emph{semantic displacement} of a word.
+        This essentially measured how much a word's vector representation has moved in the embedding space between two time periods.
+        \item The study demonstrates that their method effectively identified semantic change in words and uncovered two statistical `laws' of semantic change:
+        \begin{itemize}
+            \item \para{Law of Conformity}, which suggests that the rate of semantic change is inversely proportional to a word's frequency.
+            High-frequency words tend to change meaning more slowly.
+            \item \para{Law of Innovation}, on the other hand, proposes that words with multiple meanings are more likely to undergo semantic transformations over time.
+        \end{itemize}
+        \item They leveraged the rich information within word embeddings to quantify the degree of semantic change (semantic displacement) and identified potential patterns.
+    \end{itemize}
+
+    \item \para{Semantic Shifts with Contextual Embeddings:}
+        Giulianelli et al.\ (2020) explored the phenomenon of lexical semantic change using an unsupervised approach using \emph{BERT}.
+        \begin{itemize}
+            \item They utilized the \emph{BERT} model to generate contextual word embeddings, which captured the meaning of a word based on its surrounding context.
+            \item The extracted word usage vectors are then clustered into different \emph{usage types} using the \emph{k-means clustering algorithm}.
+            This helped to identify distinct senses or meanings of the words as they appear in various contexts.
+            \item They analyzed these clusters for a specific word across different time periods.
+            The study proposed three metrics to quantify semantic change:
+            \begin{itemize}
+                \item \textbf{Entropy Difference (ED)}: Measures the change in uncertainty (entropy) of a word’s usage distribution over time.
+                \item \textbf{Jensen-Shannon Divergence (JSD)}: Compares the similarity of word usage distributions across time intervals.
+                \item \textbf{Average Pairwise Distance (APD)}: Computes the average distance between word usage vectors from different periods, indicating shifts in word meaning.
+            \end{itemize}
+            \item The qualitative analysis indicated that the approach could capture various linguistic phenomena, including both synchronic (current usage) and diachronic (historical changes) aspects.
+        \end{itemize}
+
     \item \para{Semantic Similarity:} The impact of significant events like the COVID-19 pandemic on language and semantic change has also been a subject of study.
@@
 
-\para{Syntactic Change.}
-The syntactic functionality of a word can evolve by transitioning into a new part-of-speech (POS) category.
+% --------------------------------- SYNTACTIC ---------------------------------
+\para{Linguistic change studies have shifted to include \emph{syntax} alongside semantics.}
+The field of linguistic change has traditionally focused on \emph{semantics}.
+However, recent studies have begun to explore the role of \emph{syntax} in language evolution as well.
+The \emph{syntactic} functionality of a word can evolve by transitioning into a new part-of-speech (POS) category.
 Nouns, due to their inherent flexibility in meaning, exhibit a greater tendency to undergo these changes driven by cultural shifts.
-While are more likely to participate in gradual semantic changes that follow established linguistic patterns.
+While verbs are more likely to participate in gradual semantic changes that follow established linguistic patterns.
 \vspace{0mm}
 \begin{itemize}
-    \item \textbf{Acquiring a new POS:}
-    Kulkarni et al.\ assigned part-of-speech (POS) tags to a large collection of text (corpus) and then calculated how likely different words were to appear in certain grammatical contexts (probability distributions).
-    They used the word `apple' as an example, which transitioned from being used as a common noun (e.g., a fruit) to a proper noun (referring to the Apple company) after the company's rise in the 1980s.
-    \item \para{Nouns vs. Verbs:}
-     Hamilton et al.\ (2016) highlight how cultural changes, often influenced by new technologies, are closely tied to transformations within local neighborhoods, particularly sensitive to shifts in nouns.
-     On the other hand, linguistic changes, exemplified by subjectification, are more associated with global measures and are particularly responsive to variations in verbs.
-     The evolution of words like `actually', `must', and `promise' demonstrate these changes.
+    \item \para{Acquiring a new POS.}
+    Kulkarni et al.\ assigned part-of-speech (POS) tags to a large collection of text.
+    They calculated the likelihood of a word appearing in specific grammatical contexts over time.
+    \begin{itemize}
+        \item To quantify temporal change, they compared the probability distributions of POS tags for a particular word across different time periods.
+        This essentially measured the divergence between these distributions.
+        \item An example they cited was the word `windows'.
+        Its POS tag shifted from a common noun (referring to doors and windows of a house) to a proper noun (`Microsoft Windows').
+        This highlights how a word's grammatical function can change alongside its meaning.
+    \end{itemize}
+
+    \item \para{Capturing Differences Between Nouns and Verbs.}
+     Hamilton et al.\ (2016) highlight how \emph{cultural changes}, often influenced by new technologies, are closely tied to transformations within \emph{local neighborhoods}, particularly sensitive to shifts in nouns.
+     On the other hand, \emph{linguistic changes} are more associated with \emph{global measures} and are particularly responsive to variations in verbs.
+     \begin{itemize}
+         \item To validate this hypothesis, the authors employed a statistical technique called a \emph{linear mixed model} where word type (noun or verb) is a fixed effect,
+         amount of change measured by each metric (local or global) treated as the dependent variable.
+         By analyzing the model's results, they could assess whether there's a significant difference in the way nouns and verbs exhibit change.
+         \item The evolution of words like `actually', `must', and `promise' demonstrate these changes.
          For instance, `must' has transitioned from expressing obligation to indicating necessity, showcasing a common pattern seen in modal verbs.
-     Similarly, `promise' shows how performative speech acts undergo significant pragmatic and subjectification-related changes over time.
+     \end{itemize}
+
 %    Cultural changes like new technologies are related to local neighbourhood change, which is more sensitive to changes in nouns.
@@
 %    promise represents the class of shifting “performative speech acts” that undergo rich changes due to their pragmatic uses and subjectification
-    \item \para{Acquiring a new sense:}
-    Words are complex entities with multiple senses that can evolve over time while their form remains constant, as discussed by (Laurino et al., 2023).
-    Their research studies the overall mental lexicon of words, focusing on quantifying the drift of word concepts.
-    The study highlights how most words possess a range of meanings that can be added, removed, or modified across different temporal contexts.
+
+%    \item \para{Acquiring a new sense:}
+%    Words are complex entities with multiple senses that can evolve over time while their form remains constant, as discussed by (Laurino et al., 2023).
+%    Their research studies the overall mental lexicon of words, focusing on quantifying the drift of word concepts.
+%    The study highlights how most words possess a range of meanings that can be added, removed, or modified across different temporal contexts.
 \end{itemize}
@@
 They also employ semantic similarity analysis to quantify the shifts in meaning for pandemic-related words and provides evidence that the COVID-19 pandemic acted as a catalyst for rapid semantic change.
+\end{itemize}
 
-\end{itemize}
+\begin{table}[t]
+\centering
+\begin{tabular}{lllll}
+\toprule
+\textbf{Causes}    & \textbf{Methods}        & \textbf{Technical Method} & \textbf{Time Period} & \textbf{Example Papers} \\ \hline
+\textbf{Semantic}  & \textbf{Frequency}      & Log ratio                 & 1960 \& 1990         & {[}1{]}                 \\ \hline
+                   &                         & Word Frequencies          & 1900 - 2005          & {[}5{]}                 \\ \cline{3-5}
+                   &                         & Cosine Distance           &                      &                         \\ \cline{2-5}
+                   & \textbf{Distributional} & Co-occurence              & 1960 \& 1990         & {[}1{]}                 \\ \cline{2-5}
+                   &                         & NLM                       &                      & {[}4{]}                 \\ \cline{3-5}
+                   &                         & Cosine Similarity         & 1900 - 2005          & {[}5{]}                 \\ \cline{3-5}
+                   &                         & Aligning Embeddings       &                      & {[}3{]}                 \\ \cline{3-5}
+                   &                         & DBSCAN                    &                      & {[}8{]}                 \\ \cline{3-5}
+                   &                         & BERT                      &                      & {[}8{]}                 \\ \hline
+\textbf{Syntactic} & \textbf{POS Tags}       & JS Divergence             & 1900-2000            & {[}3{]}                 \\ \hline
+                   &                         & Cosine Distance           & 1800 - 1990          & {[}2{]}                 \\ \cline{3-5}
+\textbf{}          & \textbf{}               & KL Divergence             & 2020 \& 2023         & {[}7{]}                 \\
+\bottomrule
+\end{tabular}
+\caption{Various types of linguistic change and method used to detect them.}
+\label{tab:related-work}
+\end{table}

### 2024-07-22 src/sections/relatedwork.tex
added related work section, pending:table citation, caption

@@
 
+\begin{table}[]
+\centering
+\small
+\begin{tabular}{@{}llllll@{}}
+\toprule
+\multicolumn{1}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}Linguistic \\ Change\end{tabular}}} & \multicolumn{1}{c}{\textbf{Approach}}                                              & \multicolumn{1}{c}{\textbf{TCD}} & \multicolumn{3}{c}{\textbf{Example}}                                                                                                                                                                 \\ \cmidrule(l){4-6}
+\multicolumn{1}{c}{\textbf{}}                                                             & \multicolumn{1}{c}{\textbf{}}                                              & \multicolumn{1}{c}{\textbf{}}    & \textbf{Word}         & \textbf{Former Usage}                                                                & \textbf{New Usage}                                                                    \\ \midrule
+\multicolumn{1}{l}{\textbf{Semantic}}                                                     & \multicolumn{1}{l}{\textbf{FREQUENCY}}                                             & \multicolumn{1}{l}{}            & \multicolumn{1}{l}{} & \multicolumn{1}{l}{}                                                                & \multicolumn{1}{l}{}                                                                 \\ \midrule
+\textbf{}                                                                                 & Log Ratio                                                                  & 1960 \& 1990                     & disk                  & -                                                                                    & -                                                                                     \\ \cmidrule(l){2-6}
+\textbf{}                                                                                 & \begin{tabular}[c]{@{}l@{}}Word \\ Frequency\end{tabular}                  & 1900--2005                      & bitch                 & female dog                                                                           & slang                                                                                 \\ \cmidrule(l){2-6}
+\textbf{}                                                                                 & \begin{tabular}[c]{@{}l@{}}Law of \\ Conformity\end{tabular}               & 1800--2000                      & -                     & -                                                                                    & -                                                                                     \\ \cmidrule(l){2-6}
+                                                                                          & \textbf{DISTRIBUTIONAL}                                                             &                                  &                       &                                                                                      &                                                                              \\ \cmidrule(l){2-6}
+\textbf{}                                                                                 & \begin{tabular}[c]{@{}l@{}}Local Mutual \\ Information\end{tabular}        & 1960 \& 1990                     & sleep                 & deep sleep                                                                           & sleep disorder                                                                        \\ \cmidrule(l){2-6}
+\textbf{}                                                                                 & \begin{tabular}[c]{@{}l@{}}Continuous \\ Word \\ Embeddings\end{tabular}   & 1900--2009                      & cell                  & closet, dungeon                                                                      & phone, cordless                                                                       \\ \cmidrule(l){2-6}
+\textbf{}                                                                                 & \begin{tabular}[c]{@{}l@{}}Change Point \\ Algorithm\end{tabular}          & 1900--2005                      & gay                   & cheerful, dapper                                                                     & \begin{tabular}[c]{@{}l@{}}lesbian, \\ homosexual\end{tabular}                        \\ \cmidrule(l){2-6}
+\textbf{}                                                                                 & Clustering (DBSCAN)                                                        & 1900--2000                      & mouse                 & mice, rat                                                                            & cursor, pointer                                                                       \\ \cmidrule(l){2-6}
+\textbf{}                                                                                 & \begin{tabular}[c]{@{}l@{}}Cultural and \\ linguistic Drift\end{tabular}   & 1800--2000                      & virus                 & \begin{tabular}[c]{@{}l@{}}infected with\\  the virus\end{tabular}                   & \begin{tabular}[c]{@{}l@{}}spreading\\  computer virus\end{tabular}                   \\ \cmidrule(l){2-6}
+\textbf{}                                                                                 & \begin{tabular}[c]{@{}l@{}}Diachronic \\ Word\\ Embeddings\end{tabular}    & 1800--1999                      & awful                 & full of awe                                                                          & terrible or appalling                                                                 \\ \cmidrule(l){2-6}
+\textbf{}                                                                                 & \begin{tabular}[c]{@{}l@{}}Contextual \\ Word \\ Embeddings\end{tabular}                                                                      & 1910 – 2009                      & tenure                & \begin{tabular}[c]{@{}l@{}}short term leases,\\ insecurity of tenure\end{tabular}    & \begin{tabular}[c]{@{}l@{}}tenure of office, \\ employment \\ and tenure\end{tabular} \\ \cmidrule(l){2-6}
+\textbf{}                                                                                 & \begin{tabular}[c]{@{}l@{}}Semantic \\ Similarity \\ Analysis\end{tabular} & 2014--2022                      & immunity              & \begin{tabular}[c]{@{}l@{}}politics (‘legislator’,\\ ‘representative’).\end{tabular} & \begin{tabular}[c]{@{}l@{}}health-related \\ (‘prevention’, ‘AIDS’)\end{tabular}      \\ \cmidrule(l){2-6}
+                                                                                          & \textbf{PART OF SPEECH}                                                                        &                                  &                       &                                                                                      &                                                                   \\ \midrule
+\textbf{Syntactic}                                                                        & POS Tags                                                                   & 1900--2000                        & windows               & \begin{tabular}[c]{@{}l@{}}doors and \\ windows of a house\end{tabular}              & Microsoft Windows                                                                     \\ \midrule
+\textbf{}                                                                                 & \begin{tabular}[c]{@{}l@{}}Mixed-model \\ Regressions\end{tabular}         & 1800--2000                      & actually              & originally, nominally                                                                & presumed, believe\\
+
+\bottomrule
+\end{tabular}
+%\bottomrule
+\caption{Various types of linguistic change and method used to detect them.}
+\label{tab:related-work}
+\end{table}
+
+% ------------------------------------------------------------------ SEMANTIC ------------------------------------------------------------------
 \begin{itemize}
@@
 
-% --------------------------------- SYNTACTIC ---------------------------------
+% ------------------------------------------------------------------ SYNTACTIC ------------------------------------------------------------------
 \para{Linguistic change studies have shifted to include \emph{syntax} alongside semantics.}
@@
      \end{itemize}
-
-%    Cultural changes like new technologies are related to local neighbourhood change, which is more sensitive to changes in nouns.
-%    Examples: gay, virus, cell
-%    Linguistic changes like subjectification is related to global measure, which is more sensitive to change in verbs.
-%    Examples: actually, must, promise
-%    actually is a case of subjectification
-%    must shifted from a deontic/obligation usage (“you must do X”) to a epistemic one (“X must be the case”), exemplifying a regular pattern of change common to many modal verb
-%    promise represents the class of shifting “performative speech acts” that undergo rich changes due to their pragmatic uses and subjectification
-
-%    \item \para{Acquiring a new sense:}
-%    Words are complex entities with multiple senses that can evolve over time while their form remains constant, as discussed by (Laurino et al., 2023).
-%    Their research studies the overall mental lexicon of words, focusing on quantifying the drift of word concepts.
-%    The study highlights how most words possess a range of meanings that can be added, removed, or modified across different temporal contexts.
 \end{itemize}
 
-\vspace{5mm}
-%Among the reviewed literature, three studies are particularly pertinent to our analysis of data voids.
-%In case of data voids, we want to find shifts from it's previous usage to new usage.
+% ------------------------------------------------------------------ THREE PAPERS ------------------------------------------------------------------
 Several studies from our literature review provide valuable insights for addressing data voids in our analysis.
@@
 \end{itemize}
-
-\begin{table}[t]
-\centering
-\begin{tabular}{lllll}
-\toprule
-\textbf{Causes}    & \textbf{Methods}        & \textbf{Technical Method} & \textbf{Time Period} & \textbf{Example Papers} \\ \hline
-\textbf{Semantic}  & \textbf{Frequency}      & Log ratio                 & 1960 \& 1990         & {[}1{]}                 \\ \hline
-                   &                         & Word Frequencies          & 1900 - 2005          & {[}5{]}                 \\ \cline{3-5}
-                   &                         & Cosine Distance           &                      &                         \\ \cline{2-5}
-                   & \textbf{Distributional} & Co-occurence              & 1960 \& 1990         & {[}1{]}                 \\ \cline{2-5}
-                   &                         & NLM                       &                      & {[}4{]}                 \\ \cline{3-5}
-                   &                         & Cosine Similarity         & 1900 - 2005          & {[}5{]}                 \\ \cline{3-5}
-                   &                         & Aligning Embeddings       &                      & {[}3{]}                 \\ \cline{3-5}
-                   &                         & DBSCAN                    &                      & {[}8{]}                 \\ \cline{3-5}
-                   &                         & BERT                      &                      & {[}8{]}                 \\ \hline
-\textbf{Syntactic} & \textbf{POS Tags}       & JS Divergence             & 1900-2000            & {[}3{]}                 \\ \hline
-                   &                         & Cosine Distance           & 1800 - 1990          & {[}2{]}                 \\ \cline{3-5}
-\textbf{}          & \textbf{}               & KL Divergence             & 2020 \& 2023         & {[}7{]}                 \\
-\bottomrule
-\end{tabular}
-\caption{Various types of linguistic change and method used to detect them.}
-\label{tab:related-work}
-\end{table}

### 2024-07-22 src/sections/relatedwork.tex
modified the table, added citations

@@
 \multicolumn{1}{c}{\textbf{}}                                                             & \multicolumn{1}{c}{\textbf{}}                                              & \multicolumn{1}{c}{\textbf{}}    & \textbf{Word}         & \textbf{Former Usage}                                                                & \textbf{New Usage}                                                                    \\ \midrule
-\multicolumn{1}{l}{\textbf{Semantic}}                                                     & \multicolumn{1}{l}{\textbf{FREQUENCY}}                                             & \multicolumn{1}{l}{}            & \multicolumn{1}{l}{} & \multicolumn{1}{l}{}                                                                & \multicolumn{1}{l}{}                                                                 \\ \midrule
+\multicolumn{1}{l}{\textbf{Semantic}}                                                     & \multicolumn{5}{c}{\textbf{FREQUENCY \cite{gulordava-baroni-2011-distributional, kulkarni2014statisticallysignificantdetectionlinguistic, hamilton-etal-2016-diachronic}}}                                             &            &  &                                                                &                                                                 \\ \midrule
 \textbf{}                                                                                 & Log Ratio                                                                  & 1960 \& 1990                     & disk                  & -                                                                                    & -                                                                                     \\ \cmidrule(l){2-6}
 \textbf{}                                                                                 & \begin{tabular}[c]{@{}l@{}}Word \\ Frequency\end{tabular}                  & 1900--2005                      & bitch                 & female dog                                                                           & slang                                                                                 \\ \cmidrule(l){2-6}
-\textbf{}                                                                                 & \begin{tabular}[c]{@{}l@{}}Law of \\ Conformity\end{tabular}               & 1800--2000                      & -                     & -                                                                                    & -                                                                                     \\ \cmidrule(l){2-6}
-                                                                                          & \textbf{DISTRIBUTIONAL}                                                             &                                  &                       &                                                                                      &                                                                              \\ \cmidrule(l){2-6}
+
+                                                                                          & \multicolumn{5}{c}{\textbf{DISTRIBUTIONAL \cite{gulordava-baroni-2011-distributional, kim-etal-2014-temporal, kulkarni2014statisticallysignificantdetectionlinguistic, 10.1007/978-3-319-50496-4_18, hamilton-etal-2016-diachronic, hamilton-etal-2016-cultural, 10.1162/opmi_a_00081, giulianelli-etal-2020-analysing}}}                                                              &                                  &                       &                                                                                      &                                                                              \\ \cmidrule(l){2-6}
 \textbf{}                                                                                 & \begin{tabular}[c]{@{}l@{}}Local Mutual \\ Information\end{tabular}        & 1960 \& 1990                     & sleep                 & deep sleep                                                                           & sleep disorder                                                                        \\ \cmidrule(l){2-6}
@@
 \textbf{}                                                                                 & \begin{tabular}[c]{@{}l@{}}Semantic \\ Similarity \\ Analysis\end{tabular} & 2014--2022                      & immunity              & \begin{tabular}[c]{@{}l@{}}politics (‘legislator’,\\ ‘representative’).\end{tabular} & \begin{tabular}[c]{@{}l@{}}health-related \\ (‘prevention’, ‘AIDS’)\end{tabular}      \\ \cmidrule(l){2-6}
-                                                                                          & \textbf{PART OF SPEECH}                                                                        &                                  &                       &                                                                                      &                                                                   \\ \midrule
+                                                                                          & \multicolumn{5}{c}{\textbf{PART OF SPEECH \cite{kulkarni2014statisticallysignificantdetectionlinguistic, hamilton-etal-2016-cultural}}}                                                                       &                                  &                       &                                                                                      &                                                                   \\ \midrule
 \textbf{Syntactic}                                                                        & POS Tags                                                                   & 1900--2000                        & windows               & \begin{tabular}[c]{@{}l@{}}doors and \\ windows of a house\end{tabular}              & Microsoft Windows                                                                     \\ \midrule

### 2024-07-26 src/sections/introduction.tex
worked on RP1 section

@@
 \section{Introduction} \label{sec:introduction}
-\subsection{What is a Data Void?}
+In the digital age, online information seekers frequently encounter \emph{search results} laden with \emph{misinformation}, \emph{biased narratives}, and \emph{conspiracy theories}.
+This exposure can potentially lead users to accept and propagate false information.
+However, this phenomenon is not universal across all search queries.
+Certain searches, known as \textit{data voids} are particularly susceptible to these issues.
+\\
+\textit{Data voids} occur when searching for terms that yield limited, non-existent, or highly problematic relevant information.
+Unlike common searches that produce abundant data, queries falling into data voids may return no results or present irrelevant and often inaccurate information.
+Data voids can be mainly categorized into two main types:
+\begin{itemize}
+    \item newly coined terms with no established information base, and
+    \item existing terms whose meanings have evolved over time.
+\end{itemize}
+This report focuses on the latter, examining data voids where context drift occurs.
+The following papers has been selected to explore this phenomenon.
+\begin{itemize}
+    \item Statistically Significant Detection of Linguistic Change./ ACM WWW 2014 ~\cite{kulkarni2014statisticallysignificantdetectionlinguistic}
+    \item Diachronic Word Embeddings Reveal Statistical Laws of Semantic Change./ ACL 2016 ~\cite{hamilton-etal-2016-diachronic}
+    \item The Pandemic in Words: Tracking Fast Semantic Changes via a Large-Scale Word Association Task. \cite{10.1162/opmi_a_00081}
+\end{itemize}
+
+\subsection{What is a Data Void?}\label{subsec:what-is-a-data-void?}
+Data voids, as first introduced by Golebiewski and Boyd \cite{unknown},
+are identified while searching terms for which relevant information is either limited, non-existent, or deeply problematic.
+\\
+For example, unlike a typical web search for `basketball' that yields a vast amount of data,
+data voids present a problematic scenario such as searches on these terms may return no results or irrelevant, often inaccurate information.
+This situation could prompt users to pursue information, click on misleading search outcomes, and be exposed to false or misleading content,
+ultimately shaping their perceptions and beliefs with misinformation.
+
+Data Voids often originate in fringe communities.
+When vulnerable users search for information using this term, they may encounter results that are misrepresentations of the topic, as these fringe communities have already shaped the narrative.
+Mainstream media, particularly those with far-left or far-right biases, may pick up and amplify these narratives, contributing to the misinformation cycle.
+
 \subsection{Example of a data void?(Migrant Caravan)}
+Let's understand data void through an example of a term - \emph{migrant caravan}
+The term ‘migrant caravan’ originally described a group of migrants traveling together for safety and mutual support.
+However, its usage, particularly in political discussions, has taken on a more negative tone.
+Today, it is commonly associated with large groups of migrants, often from Central America, journeying through Mexico to reach the United States.
+
 \subsection{Lifecycle of data void (from long tail to spike)}

### 2024-07-26 src/sections/paper_kulkarni.tex
worked on RP1 section

@@
 \section{RP1: Statistically Significant Detection of Linguistic Change} \label{sec:paper_kulkarni}
-\subsection{Introduction} \label{subsec:introduction}
-This research aims to quantify linguistic shifts in word meaning and usage across time, focusing on semantic or contextual changes.
-To understand data voids, we are sure that a change is happening, but we need to understand the change that is happening over time.
-Also after knowing that a change has occurred, we would like to know when did that change really occur so that we can co-relate that the data void occured when a political event happened or how the COVID outbreak event,
-which could help us identify that the term were introduced due to that event, which can be a future direction we can look into.
-This paper is exciting because the ideas inspired me to pursue this project and get future directions as well.
+\subsection{Introduction} \label{subsec:kulkarni_introduction}
 
-Their approach focuses on three methods to construct time series for a word.
-The first one is based on frequency, which captures sudden changes in word usage.
-In our work, this corresponds to strategic new terms data voids or breaking news data void,
-where breaking news data void shows that there's comes a sudden change in a word activity whenever's there an event happen.
-For example, the term "filmyourhospital" was introduced in during 2020 when coronavirus took place, which corresponds to a sudden spike.
-And this term was never used before 2020, making it a new term that introduced to support the conspiracy theory regarding encouraged people to visit local hospitals to take pictures and videos of empty hospitals to help “prove” that the COVID-19 pandemic is an elaborate hoax.
+Kulkarni et al.\ (2015) explores the problem of detecting changes in word usage patterns over time.
+They introduce methods to identify words that have undergone significant shifts in \emph{meaning}, usage \emph{frequency}, or \emph{context}.
+This is achieved by analyzing historical corpora, which contain vast amounts of text data spanning multiple years or even centuries.
+The study utilizes diverse datasets, including Twitter posts, Amazon product reviews, and Google Book Ngrams, to construct time series for individual words.\\
+The research conducted by Kulkarni et al.\ exhibits a compelling alignment with our own research objectives,
+specifically focusing on the identification of data voids characterized by shifting semantic meanings over time.
+Their exploration into these linguistic shifts has revealed interesting insights, such as the evolution of the word `gay,'
+which transitioned from being cheerful or dapper to signifying homosexual or lesbian aspect around the mid-20th century.
+To achieve understanding of this phenomenon, they intend to study the following research questions.
+\begin{itemize}
+    \item \para{RQ1.} \emph{What methods can quantify the statistical relevance of observed changes in a word's usage across different time periods?}\\
+    Kulkarni et al.’s study uses three technical methods to understand how words change over time: \emph{frequency} analysis, \emph{syntactic} analysis, and \emph{distributional} analysis.
+    The frequency method tracks how often words are used, revealing spikes in usage when significant events occur, like how \emph{breaking news} data voids can create a surge in specific keywords.
+    The syntactic method looks at changes in the grammatical roles of words, while the distributional method examines word co-occurrence patterns to see how word meanings shift.
+    Their approach not only uncovers the dynamic nature of language but also highlights “data voids,” -- areas where terms have \emph{evolved}, become \emph{outdated}, or \emph{fragmented} in meaning.
 
-Their second method is Syntactic method where they analyze each word's part of speech tag.
+    \item \para{RQ2.} \emph{Given that a word's usage has changed, how can the precise moment or period of this shift be determined?}\\
+    Understanding data voids involves not only recognizing that a change is occurring but also analyzing the specifics of how that change unfolds over time.
+    It’s crucial to pinpoint not just the fact that a shift has happened, but also the exact moment when it took place.
+    Kulkarni et al.\ tackle this challenge by implementing a \emph{change point detection algorithm} based on the \emph{Mean Shift} model.
+    This method determines whether a word has experienced a significant shift and, if so, identifies the precise point at which this change occurred.
+\end{itemize}
 
-And last method is distributional method which captures overall usage of word using word co-occurence.
-This type of method can help us identify data voids which have changed their meaning over time, such as Fragmented Concepts or Outdated Terms DV.
+Next two sections, \Cref{subsec:kulkarni-rq1} and \Cref{subsec:kulkarni-rq2} focuses on the two research questions that this paper is studying, the approach they took to study and their results.
 
-They apply these techniques on three different domains: books, tweets and reviews.
+\subsection{What methods can quantify the statistical relevance of observed changes in a word's usage across different time periods?} \label{subsec:kulkarni-rq1}
 
-The study utilizes diverse datasets, including Twitter posts, Amazon product reviews, and Google Book Ngrams, to construct time series for individual words.
-Three distinct methods are employed for statistical modeling.
+\para{Overview.}
+The challenge is to identify how the meaning of words changes over time.
+To address this, they examine a corpus containing data from various domains like books, tweets, and reviews, known as the Temporal Corpus $(C)$ which spans over a time period $(S)$.
+This corpus is divided into smaller segments called snapshots $C_t$, each of length $P$.
+From these snapshots, a common vocabulary $V$ is created having words which are common across all snapshots.
+To analyze changes in these words, a time series $T(w)$ is constructed for each word $w\in V$.
+Kulkarni et al.\ propose several methods to create this time series, aimed at understanding the evolution of words over time.
+
+\begin{figure}[!h]
+    \centering
+    \includegraphics[scale=0.5]{figures/kulkarni1}
+    \caption{My caption}
+    \label{fig:label}
+\end{figure}
+
+\para{Construction of Time Series.}
+In this paper, three methods are explored for statistically modeling the evolution of words over time:
+frequency analysis, part-of-speech tagging, and word co-occurrence to create time series for each word.
+
+\textbf{(1) Frequency-based method}: The simplest and most direct method for detecting sudden changes in word usage is by analyzing frequency trends.
+\begin{wrapfigure}{R}{0.25\textwidth}
+    \centering
+    \includegraphics[width=0.25\textwidth]{figures/frequency-gay}
+    \caption{My caption}
+    \label{fig:example-gay}
+\end{wrapfigure}
+
+This approach examines how often a word is used over a given period, providing insights into shifts in its popularity or relevance.
+Changes in frequency can indicate whether a word is gaining new meanings or losing old ones, reflecting broader cultural or societal trends.\\
+For example, the word “gay” in \Cref{fig:example-gay} shows a noticeable spike in usage during the 1980s, signaling a shift in its meaning or cultural relevance.
+Tools such as the Google Books Ngram Viewer and Google Trends are used for this purpose, as they provide large datasets and visual representations of word usage across different timeframes.
+They calculate the change in probability of a word appearing over time.
+\begin{displaymath}
+$T_t(w) = \log \frac{#(w \in C_t)}{|C_t|}$ \label{eq:equation}
+\end{displaymath}
+
+\textbf{(2) Syntactic method:}
+Frequency-based metrics, while simple to compute, are susceptible to errors caused by imbalances in the corpus's domain and genre distribution.
+Fluctuations in word usage due to temporal events or the popularity of specific entities can obscure genuine shifts in word meaning.
+Moreover, a word's grammatical role can evolve over time, such as acquiring a different part of speech category.
+To study this, the corpus is annotated with part-of-speech (POS) tags, and the probability distribution of these tags is calculated for each word across different time snapshots.
+This approach allows researchers to observe how a word’s syntactic role changes over time.
+
+\textbf{(3) Distributional method}:
+For detecting subtle semantic changes, which are not changed either due to frequency or through change in part of speech,
+this method was developed to understand in which context a word is used in and based on that understand the semantic changes.
+Distributional methods focusses on creating a semantic space that maps words to continuos vector space, where each word is represented by a vector.
+Once they created a temporal word embeddings for each word in each time snapshot, then they track the changes of the represtantations across the embedding space.
 
 \begin{itemize}
-    \item \para{RQ1.} \emph{What methods can quantify the statistical relevance of observed changes in a word's usage across different time periods?}
-    \item \para{RQ2.} \emph{Given that a word's usage has changed, how can the precise moment or period of this shift be determined?}
+    \item The researchers aimed to \emph{learn word embeddings} by training neural language models on a corpus at different time snapshots.
+        They initialized word vector representations randomly and optimized the model parameters using stochastic gradient descent.
+        During training, the objective was to maximize the probability of context words appearing around a target word.
+        After training, word embeddings were normalized by their L2 norm to ensure consistent representation.
+    \item The \emph{alignment} process involved aligning word embeddings from different time snapshots into a unified coordinate system to characterize changes between them.
+        Assumptions were made to aid the alignment process, including the equivalence of spaces under linear transformation and the preservation of local structure of most words over time.
+        When the alignment model failed to align a word correctly, it indicated a potential linguistic shift, highlighting the importance of accurate alignment for tracking semantic changes over time.
+    \item By aligning embedding spaces across various time snapshots into a joint embedding space,
+        the distributional method constructs a \emph{distributional time series} that captures the semantic evolution of words over time.
+    \item The paper shows the evolution of the word `tape' over time.
+        Initially, the word `tape' referred to an `adhesive tape' but underwent a semantic shift to also mean a `cassette tape' during the early 1970s.
 \end{itemize}
+
+\subsection{Given that a word's usage has changed, how can the precise moment or period of this shift be determined?} \label{subsec:kulkarni-rq2}
+After constructing time series data for each word in the corpus—whether using frequency-based, syntactic, or distributional methods—the next step is to determine if the word has experienced a significant change.
+If a significant shift is detected, the algorithm then identifies and returns the estimated change point (ECP), which marks the time when this change occurred.
+\begin{itemize}
+    \item Since language exhibits a stochastic drift.
+        In this context, `stochastic' refers to the randomness or probabilistic nature of the training process,
+        where the models are trained on the same dataset but may produce different results each time due to this randomness.
+        To resolve this, time series was normalized for each word by transforming the time series into a Z-score series.
+    \item Then the algorithm utilizes a Mean Shift model to detect changes in the mean of the time series.
+    \item Change points can be thus identified by detecting significant shifts in the mean.
+\end{itemize}
+
+\subsection{Results} \label{subsec:kulkarni-results}
+
+\para{Time Series Analysis \& Historical Analysis.} \\
+
+\begin{table}[]
+\begin{tabular}{@{}lllll@{}}
+\toprule
+\textbf{Method}         & \textbf{Examples}                              &  &  &  \\ \midrule
+\textbf{Frequency}      & transmistted, bitch, sex, her                  &  &  &  \\
+\textbf{Syntactic}      & apple, hug, sink, click, handle, windows, bush &  &  &  \\
+\textbf{Distributional} & diet, tape, plastic                            &  &  &  \\ \bottomrule
+\end{tabular}\label{tab:time-series-examples}
+\end{table}
+
+Frequency \& Distributional - transmistted, bitch, sex, her
+The sharp increase of the word her in frequency around the 1960’s could be attributed to the concurrent rise and popularity of the feminist movement.
+Sudden temporary popularity of specific social and political events could lead the Frequency method to produce many false positives.
+
+Syntactic - apple, hug, sink, click, handle, windows, bush
+The word apple was detected uniquely by the Syntactic method as its most frequent part of speech tag changed significantly from “Noun” to “Proper Noun”
+The Syntactic method is indicative of having low false positive rate, but suffers from a high false negative rate, given that only two words in the table were detected.
+
+Distributional - diet, tape, plastic
+The popularity of books on dieting started with the best selling book Dr. Atkins’ Diet Revolution by Robert C. Atkins in 1972.
+This changed the use of the word diet to mean a life-style of food consumption behavior and not only the food consumed by an individual or group.
+
+\para{Cross Domain Analysis.}
+Amazon Reviews and Twitter (content that spans a much shorter time scale as compared to Google Books Ngram Corpus)
+What is the effect of applying distributional method on Amazon Review, Twitter dataset
+
+\begin{table}[]
+\begin{tabular}{@{}lllll@{}}
+\toprule
+\textbf{Dataset Source} & \textbf{Examples}                     &  &  &  \\ \midrule
+\textbf{Amazon Reviews} & streaming, ray, combo, rays, twilight &  &  &  \\
+\textbf{Twitter Tweets} & candy, myster, rally, sandy           &  &  &  \\ \bottomrule
+\end{tabular}\label{tab:sources-examples}
+\end{table}
+
+Words - streaming, ray, combo, rays, twilight
+
+Twitter - candy, myster, rally, sandy
+The word sandy acquired a new sense after the incident of Hurricane Sandy hitting the East Coast of USA, which was also mentioned in the data voids example by danah boyd.
+
+These examples show that their method can detect the introduction of new products, movies, and books.
+
+\subsection{Takeaways} \label{subsec:kulkarni-takeaways}
+The paper by Kulkarni et al.\ (2015) contributes a novel and rigorous approach to detecting linguistic changes using statistical methods and word embeddings.
+Their contribution in using large historical corpora and word embeddings to capture the dynamic nature of language.
+A statistical framework that ensures the detected changes are significant, avoiding false positives due to noise or minor fluctuations.
+
+Despite the fast pace of change of the web content, our method is able to detect the introduction of new products, movies and books.
+This could help semantically aware web applications to better understand user intentions and requests.
+
+%Also, after knowing that a change has occurred, we would like to know when did that change really occur so that we can co-relate that the data void occured when a political event happened or how the COVID outbreak event,
+%which could help us identify that the term were introduced due to that event, which can be a future direction we can look into.
+
+%For example, the term "filmyourhospital" was introduced in during 2020 when coronavirus took place, which corresponds to a sudden spike.
+%And this term was never used before 2020, making it a new term that introduced to support the conspiracy theory regarding encouraged people to visit local hospitals to take pictures and videos of empty hospitals to help “prove” that the COVID-19 pandemic is an elaborate hoax.
+%Correlate with historical events.

### 2024-07-26 src/sections/relatedwork.tex
worked on RP1 section

@@
 %\bottomrule
-\caption{Various types of linguistic change and method used to detect them.}
+\caption{\textbf{Comparative Overview of Linguistic Change Detection Approaches.}
+This table provides a comprehensive summary of various approaches and contributions to the study of linguistic changes,
+    categorizing the type of changes (semantic or syntactic), and detailing the methodologies used, including frequency and distributional approaches.
+    It highlights each study’s methods, such as word embeddings, statistical analysis, and clustering, along with the time periods of change detection and examples of changing words.
+    The table provides an overview of the advancements and differences in analyzing language evolution,
+    capturing the breadth of approaches from traditional frequency counts to modern embedding techniques.}
 \label{tab:related-work}

### 2024-07-30 src/sections/paper_hamilton.tex
worked on RP1 section
added context for hamilton paper

@@
 
-\section{RP2.} \label{sec:paper_hamilton}
-\subsection{Diachronic Word Embeddings Reveal Statistical Laws of Semantic Change}
+\section{RP2. Diachronic Word Embeddings Reveal Statistical Laws of Semantic Change} \label{sec:paper_hamilton}
+\subsection{Introduction}\label{subsec:hamilton_introduction}
+Hamilton et al.\ (2016) explores the dynamic nature of language by analyzing how word meanings evolve over time.
+Using \emph{diachronic} (historical) word embeddings, a technique that maps words into vector spaces based on historical corpora, the study uncovers patterns and laws governing semantic shifts.
+Their approach provides a quantitative framework to examine how words change meaning, influenced by cultural, and linguistic factors.
 
-1.	Creating Word Embeddings:
-•	They create word embeddings for different time periods using historical text corpora.
-2.	Using SVD, PPMI, and SGNS:
-•	SVD (Singular Value Decomposition): They apply SVD to the Positive Pointwise Mutual Information (PPMI) matrices to reduce the dimensionality of the word vectors.
-•	PPMI (Positive Pointwise Mutual Information): They use PPMI to measure the association between words and their contexts.
-PPMI values are computed for each time period to capture word-context relationships.
-•	SGNS (Skip-Gram with Negative Sampling): They also use the SGNS model to create word embeddings, leveraging its effectiveness in capturing semantic relationships.
-3.	Aligning Word Embeddings:
-•	They align word embeddings across different time periods to ensure that the same word in different periods is represented in a comparable way.
-This alignment allows for tracking changes in word meanings over time.
-4.	Analyzing Semantic Change:
-•	By comparing word embeddings from different time periods, they analyze the semantic shifts.
-They use cosine similarity to measure changes in word meanings and identify words that have undergone significant semantic shifts.
+The pace at which words undergo semantic change differs, with some words changing meaning more frequently compared to others.
+For example, the word `cat' has remained relatively stable in its meaning, while `cast' has evolved to have multiple meanings ~\cite{hamilton-etal-2016-diachronic}.
 
-•	PPMI: They computed PPMI matrices for each time period, capturing the association between words and their contexts.
-•	SVD: They applied SVD to the PPMI matrices to reduce the dimensionality of the vectors, making them more manageable and interpretable.
-•	SGNS: They used the SGNS model to directly create word embeddings, leveraging its ability to capture semantic relationships in a low-dimensional space.
+There are several hypotheses about the patterns in semantic change, including the increasing subjectification of meaning or the grammaticalization.
+This paper, however, focuses on following two specific questions about semantic change:
+\begin{itemize}
+    \item \para{RQ1.} \emph{What role does word frequency play in the evolution of word meanings?}\\
+    Frequency is a significant factor in linguistic changes, with high-frequency words often changing faster, while low-frequency words tend to be more resistant to change.
+    The connection between word frequency and semantic change is a key unanswered question in the field of linguistics.
+    The authors introduce the \textbf{law of conformity} to address this gap, demonstrating that frequent words change more slowly, and it clarifies the role of frequency in semantic change.
+
+    \item \para{RQ2.} \emph{How does polysemy relate to semantic change?}\\
+    Another unresolved question in linguistics is the relationship between semantic change and polysemy.
+    Polysemous words, which have multiple meanings, appear in a variety of contexts.
+    It remains unclear whether the diverse contextual use of these words makes them more or less prone to undergoing semantic change.
+    The authors propose the \textbf{law of innovation}, which demonstrates that polysemous words are more likely to experience faster semantic changes.
+    If a word has multiple meanings, it is more likely to lead to semantic change, especially in the case of rare senses.
+\end{itemize}
+
+The paper aims to develop a robust methodology for quantifying semantic change using word embeddings and comparing different approaches to analyzing semantic change.
+This methodology is then applied in a large-scale cross-linguistic analysis spanning 200 years and four languages (English, German, French, and Chinese) to propose the above two statistical laws relating frequency and polysemy to semantic change.
+
+\subsection{Constructing Word Embeddings.}\label{subsec:constructing-word-embeddings.}
+The authors employ three methods to construct word embeddings for different time periods.
+Initially, they create embeddings for each distinct period and then align these embeddings over time to maintain consistency.
+To quantify semantic change, they use various metrics.
+Specifically, they utilize Singular Value Decomposition (SVD), Positive Pointwise Mutual Information (PPMI), and Skip-Gram with Negative Sampling (SGNS).
+These distributional techniques represent each word by a vector that encapsulates information about the word’s co-occurrence statistics.
+
+\begin{enumerate}
+    \item
+\para{PPMI: }
+PPMI is a statistical measure used to quantify the association between a word and its context within a corpus.
+In the paper, it is used to create word embeddings by capturing the strength of association between words and their co-occurring contexts over time.
+
+\begin{itemize}
+    \item \para{Co-occurrence Matrix Construction.}
+    First, a co-occurrence matrix is built, where each entry represents the frequency with which a word (target) and its context (typically a window of neighboring words) appear together in the corpus.
+    \item \para{Joint Probability Calculation.}
+    The joint probability of a target word $w_i$ and a context word $c_j$ is calculated by dividing the co-occurrence frequency of $w_i$ and $c_j$ by the total number of word pairs in the corpus.
+    \item \para{Marginal Probability Calculation.}
+    The marginal probabilities of the target word $w_i$ and the context word $c_j$ are calculated based on the total occurrences of each word in the corpus.
+    \item \para{Calculating PPMI.}
+    The Positive Pointwise Mutual Information (PPMI) between the target word $w_i$ and the context word $c_j$ is calculated using the formula:
+
+    \begin{equation}
+    \mathbf{M}^{\text{PPMI}}_{i,j} = \max \left\{ \log \left( \frac{\hat{p}(w_i, c_j)}{\hat{p}(w_i) \hat{p}(c_j)} \right) - \alpha, 0 \right\},
+    \label{eq:ppmi}
+    \end{equation}
+
+    where:
+    \begin{itemize}
+    \item $\hat{p}(w_i, c_j)$ is the estimated \emph{joint probability} of the target word $w_i$ and the context word $c_j$ occurring together.
+    \item $\hat{p}(w_i)$ and $\hat{p}(c_j)$ are the estimated \emph{marginal probabilities} of the target word $w_i$ and the context word $c_j$, respectively.
+    \item $\alpha > 0$ is a positive constant used to smooth the values and prevent extreme positive values in the PPMI calculation.
+    \end{itemize}
+    This formula refines the PPMI measure by incorporating $\alpha$ to adjust the log probability ratio.
+    The inclusion of $\alpha$ helps manage the sparseness of the data and the impact of low-frequency events, making the measure more robust and stable.
+\end{itemize}
+
+    \item
+\para{SVD: }
+SVD is a mathematical technique that factorizes a matrix into three other matrices, revealing important features of the original matrix.
+In the context of word embeddings, SVD helps to reduce the dimensionality of the data while preserving the essential structure and patterns in word co-occurrences.
+SVD decomposes the co-occurrence matrix into three matrices in the following manner:
+
+\begin{equation}
+\mathbf{w}_i^{\text{SVD}} = (\mathbf{U} \Sigma^\gamma)_i,
+\label{eq:svd}
+\end{equation}
+
+where:
+\begin{itemize}
+    \item $\mathbf{w}_i^{\text{SVD}}$ is the vector representation of the word $w_i$ in the SVD-based embedding space.
+    \item $\mathbf{U}$ is the matrix of left singular vectors obtained from the decomposition of the co-occurrence matrix.
+    \item $\Sigma$ is the diagonal matrix of singular values.
+    \item $\gamma$ is a parameter that scales the singular values, allowing for tuning the influence of the different components.
+\end{itemize}
+
+In this formula, the word vector $\mathbf{w}_i^{\text{SVD}}$ is derived by scaling the singular values in $\Sigma$ by $\gamma$ and then multiplying by the corresponding vectors in $\mathbf{U}$.
+This approach provides a flexible way to adjust the contribution of the components to the final word embeddings.
+
+    \item
+\para{SGNS: }
+SGNS is a technique used to learn high-quality word vectors by training on large corpora.
+The primary goal of SGNS is to learn word embeddings such that words that appear in similar contexts have similar vector representations.
+\\
+For a given target word  w  and a context word  c , the model tries to maximize the probability  P(c|w)
+
+\begin{equation}
+\hat{p}(c_i \mid w_i) \propto \exp(\mathbf{w}_i^{\text{SGNS}} \cdot \mathbf{c}_j^{\text{SGNS}}),
+\label{eq:sgns}
+\end{equation}
+
+where:
+
+\begin{itemize}
+    \item $\hat{p}(c_i \mid w_i)$ is the estimated probability of a context word $c_i$ given a target word $w_i$.
+    \item $\mathbf{w}_i^{\text{SGNS}}$ is the vector representation of the target word $w_i$ in the SGNS embedding space.
+    \item $\mathbf{c}_j^{\text{SGNS}}$ is the vector representation of the context word $c_j$ in the SGNS embedding space.
+\end{itemize}
+
+\\
+The probability $\hat{p}(c_i \mid w_i)$ that a word $c_i$ is a context word of $w_i$ is proportional to the exponential of the dot product of their corresponding vector representations.
+The dot product $\mathbf{w}_i^{\text{SGNS}} \cdot \mathbf{c}_j^{\text{SGNS}}$ measures the similarity between the target word and the context word in the vector space.
+A higher dot product indicates that the words are more likely to co-occur, leading to a higher estimated probability.\\
+
+The model uses this probability estimation to distinguish true word-context pairs from randomly sampled "negative" pairs.
+The training objective is to maximize the probability of observed (target, context) pairs and minimize the probability of randomly sampled pairs, thereby learning embeddings that reflect the semantic relationships between words.
+
+\end{enumerate}
+
+\subsection{Aligning Embeddings}\label{subsec:aligning-embeddings}
+Since the embeddings are trained independently for each time period, their vector spaces may not be directly comparable.
+The paper addresses this by aligning the embeddings from different time periods.
+They aligned the emebddings across time periods using orthogonal Procrustes, which transforms the embeddings to minimize the distances between corresponding words in different periods.
+This step ensures that similar meanings are represented similarly across time.
+
+\subsection{Measuring Semantic Change}\label{subsec:measuring-semantic-change}
+Once the embeddings are aligned, following methods are performed to quantify how much the meaning of word changes over time.
+\begin{itemize}
+    \item \para{Pair-wise similarity time-series}
+        This method measures changes in the similarity between pairs of words over different time periods using cosine similarity.
+        \begin{equation}
+            s^{(t)}(w_i, w_j) = \text{cos-sim}(w_i^{(t)}, w_j^{(t)})
+            \label{eq:equation2}
+        \end{equation}
+        The Spearman correlation $(\rho)$ is employed to measure the relationship between the similarity scores of word pairs and time.
+        This non-parametric method assesses whether the similarity series shows a significant increase or decrease over time, which is crucial for understanding semantic shifts.
+
+    \item \para{Measuring semantic displacement}
+        This method quantifies how much a word's meaning has changed by calculating the displacement of its vector representation across different time points.
+        They use the aligned word vectors to compute the \emph{semantic displacement} that a word has undergone during a certain time-period.
+        It is measured by calculating the distance between the vector representations of a word in different time periods.
+        For a given word $w$, if $\mathbf{w}^{(t1)}$ and $\mathbf{w}^{(t2)}$ are its embeddings in two time periods $t1$ and $t2$, respectively, then the semantic displacement $\Delta$ is computed as:
+        \begin{equation}
+            \Delta(w) = \text{cos-dist}(\mathbf{w}^{(t1)}, \mathbf{w}^{(t2)}).\label{eq:equation3}
+        \end{equation}
+\end{itemize}
+
+\subsection{Evaluation of various approaches}
+They compared the different approaches discussed in \Cref{subsec:methods-for-quantifying-semantic-change} on \emph{synchronic accuracy} (similarity within time-period) and \emph{diachronic validity} (semantic changes over time).
+\begin{itemize}
+    \item \para{Synchronic Accuracy}
+    Synchronic accuracy refers to the accuracy with which word embeddings capture the semantic relationships and meanings of words at a \emph{specific point in time}.
+    It assesses how well the models represent the semantic relationships among words in a given time period.
+    SVD performed best here.
+    \item \para{Diachronic Validity}
+    Diachronic validity evaluates how effectively the methods can detect and quantify changes in \emph{meaning over time}.
+    It involves testing the methods against historical data to see if they can accurately identify known shifts in word meanings.
+    SGNS perfomed best on both of the tasks:
+    \begin{itemize}
+        \item \textbf{Detecting Known Shifts.}
+        The goal in this task is for the methods to correctly capture whether pairs of words moved closer or further apart in semantic space during a pre-determined time-period.
+
+        \item \textbf{Discovering Shifts from data.}
+        Tested whether the methods discover reasonable shifts by examining the top-10 words that changed the most from the 1900s to the 1990s according to the semantic displacement metric (\Cref{subsec:methods-for-quantifying-semantic-change}).
+    \end{itemize}
+
+\end{itemize}
+
+\subsection{Statistical Laws of Semantic Change}
+The authors establish two statistical laws of semantic change by analyzing a large dataset across multiple languages and time periods.
+They demonstrate these laws by involving the comparison of different word embedding models (PPMI, SVD, SGNS) against known historical semantic shifts and novel benchmarks.
+
+How diachronic embeddings can be used to reveal statistical laws that relate frequency and polysemy to semantic change.
+
+\begin{itemize}
+    \item \para{Law of conformity:} \emph{Frequently used words change at slower rates.}\\
+    Words with higher frequencies tend to experience slower rates of semantic change.
+    This is statistically supported by the observation that high-frequency words have smaller semantic displacements over time.
+    \item \para{Law of innovation:} \emph{Polysemous words change at faster rates.}\\
+    Polysemous words tend to change more rapidly.
+    The study finds that these words show larger semantic displacements, indicating more significant changes in meaning over time.
+\end{itemize}
+
+\subsection{Takeaways}
+This paper explores how different types of word embedding models, such as PPMI, SVD, and SGNS, can be used to study diachronic shifts in word meanings.
+It demonstrates the use of metrics like Spearman correlation to quantify semantic drift and employs Procrustes alignment techniques to align word embeddings across different time periods.
+By analyzing these aligned embeddings, the authors identify two key statistical laws of semantic change.
+These findings provide a quantitative framework for understanding how and why languages evolve over time.

### 2024-07-31 src/sections/introduction.tex
added SVG, fixed table

@@
 \begin{itemize}
-    \item Statistically Significant Detection of Linguistic Change./ ACM WWW 2014 ~\cite{kulkarni2014statisticallysignificantdetectionlinguistic}
-    \item Diachronic Word Embeddings Reveal Statistical Laws of Semantic Change./ ACL 2016 ~\cite{hamilton-etal-2016-diachronic}
-    \item The Pandemic in Words: Tracking Fast Semantic Changes via a Large-Scale Word Association Task. \cite{10.1162/opmi_a_00081}
+    \item Statistically Significant Detection of Linguistic Change.
+    ACM WWW 2014 ~\cite{kulkarni2014statisticallysignificantdetectionlinguistic}
+    \item Diachronic Word Embeddings Reveal Statistical Laws of Semantic Change.
+    ACL 2016 ~\cite{hamilton-etal-2016-diachronic}
 \end{itemize}

### 2024-07-31 src/sections/relatedwork.tex
added SVG, fixed table

--- 2024-07-22 src/sections/relatedwork.tex
@@
 \toprule
-\multicolumn{1}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}Linguistic \\ Change\end{tabular}}} & \multicolumn{1}{c}{\textbf{Approach}}                                              & \multicolumn{1}{c}{\textbf{TCD}} & \multicolumn{3}{c}{\textbf{Example}}                                                                                                                                                                 \\ \cmidrule(l){4-6}
-\multicolumn{1}{c}{\textbf{}}                                                             & \multicolumn{1}{c}{\textbf{}}                                              & \multicolumn{1}{c}{\textbf{}}    & \textbf{Word}         & \textbf{Former Usage}                                                                & \textbf{New Usage}                                                                    \\ \midrule
-\multicolumn{1}{l}{\textbf{Semantic}}                                                     & \multicolumn{1}{l}{\textbf{FREQUENCY}}                                             & \multicolumn{1}{l}{}            & \multicolumn{1}{l}{} & \multicolumn{1}{l}{}                                                                & \multicolumn{1}{l}{}                                                                 \\ \midrule
-\textbf{}                                                                                 & Log Ratio                                                                  & 1960 \& 1990                     & disk                  & -                                                                                    & -                                                                                     \\ \cmidrule(l){2-6}
-\textbf{}                                                                                 & \begin{tabular}[c]{@{}l@{}}Word \\ Frequency\end{tabular}                  & 1900--2005                      & bitch                 & female dog                                                                           & slang                                                                                 \\ \cmidrule(l){2-6}
-\textbf{}                                                                                 & \begin{tabular}[c]{@{}l@{}}Law of \\ Conformity\end{tabular}               & 1800--2000                      & -                     & -                                                                                    & -                                                                                     \\ \cmidrule(l){2-6}
-                                                                                          & \textbf{DISTRIBUTIONAL}                                                             &                                  &                       &                                                                                      &                                                                              \\ \cmidrule(l){2-6}
+\multicolumn{1}{c}{\textbf{Linguistic}}              & \multicolumn{1}{c}{\textbf{Approach}}                                            & \multicolumn{1}{c}{\textbf{TCD}}      & \multicolumn{3}{c}{\textbf{Example}}                                                                                                                                                                              \\ \cmidrule(l){4-6}
+\multicolumn{1}{c}{\textbf{Change}}                  & \multicolumn{1}{c}{\textbf{}}                                                    & \multicolumn{1}{c}{\textbf{}}         & \textbf{Word}             & \textbf{Former Usage}                                                                         & \textbf{New Usage}                                                                    \\ \midrule
+\rowcolor{lightgray}                                 & \multicolumn{5}{c}{\textbf{FREQUENCY \cite{gulordava-baroni-2011-distributional, kulkarni2014statisticallysignificantdetectionlinguistic, hamilton-etal-2016-diachronic}}}                                                                                                                                                                   \\
+\textbf{\textbf{Semantic}}                           & Log Ratio                                                                        & 1960 \& 1990                          & disk                      & -                                                                                             & -                                                                                     \\ \cmidrule(l){2-6}
+\textbf{}                                            & Word Frequency                                                                   & 1900--2005                            & bitch                     & female dog                                                                                    & slang                                                                                 \\ \cmidrule(l){2-6}
+\rowcolor{lightgray}                                 & \multicolumn{5}{c}{\textbf{DISTRIBUTIONAL \cite{gulordava-baroni-2011-distributional, kim-etal-2014-temporal, kulkarni2014statisticallysignificantdetectionlinguistic, 10.1007/978-3-319-50496-4_18, hamilton-etal-2016-diachronic, hamilton-etal-2016-cultural, 10.1162/opmi_a_00081, giulianelli-etal-2020-analysing}}}                                                                                                                                         \\ \cmidrule(l){2-6}
 \textbf{}                                            & \begin{tabular}[c]{@{}l@{}}Local Mutual \\ Information\end{tabular}              & 1960 \& 1990                          & sleep                     & deep sleep                                                                                    & sleep disorder                                                                        \\ \cmidrule(l){2-6}
-\textbf{}                                                                                 & \begin{tabular}[c]{@{}l@{}}Continuous \\ Word \\ Embeddings\end{tabular}   & 1900--2009                      & cell                  & closet, dungeon                                                                      & phone, cordless                                                                       \\ \cmidrule(l){2-6}
+\textbf{}                                            & \begin{tabular}[c]{@{}l@{}}Continuous Word \\ Embeddings\end{tabular}            & 1900--2009                            & cell                      & closet, dungeon                                                                               & phone, cordless                                                                       \\ \cmidrule(l){2-6}
 \textbf{}                                            & \begin{tabular}[c]{@{}l@{}}Change Point \\ Algorithm\end{tabular}                & 1900--2005                            & gay                       & cheerful, dapper                                                                              & \begin{tabular}[c]{@{}l@{}}lesbian, \\ homosexual\end{tabular}                        \\ \cmidrule(l){2-6}
 \textbf{}                                            & Clustering (DBSCAN)                                                              & 1900--2000                            & mouse                     & mice, rat                                                                                     & cursor, pointer                                                                       \\ \cmidrule(l){2-6}
-\textbf{}                                                                                 & \begin{tabular}[c]{@{}l@{}}Cultural and \\ linguistic Drift\end{tabular}   & 1800--2000                      & virus                 & \begin{tabular}[c]{@{}l@{}}infected with\\  the virus\end{tabular}                   & \begin{tabular}[c]{@{}l@{}}spreading\\  computer virus\end{tabular}                   \\ \cmidrule(l){2-6}
-\textbf{}                                                                                 & \begin{tabular}[c]{@{}l@{}}Diachronic \\ Word\\ Embeddings\end{tabular}    & 1800--1999                      & awful                 & full of awe                                                                          & terrible or appalling                                                                 \\ \cmidrule(l){2-6}
-\textbf{}                                                                                 & \begin{tabular}[c]{@{}l@{}}Contextual \\ Word \\ Embeddings\end{tabular}                                                                      & 1910 – 2009                      & tenure                & \begin{tabular}[c]{@{}l@{}}short term leases,\\ insecurity of tenure\end{tabular}    & \begin{tabular}[c]{@{}l@{}}tenure of office, \\ employment \\ and tenure\end{tabular} \\ \cmidrule(l){2-6}
-\textbf{}                                                                                 & \begin{tabular}[c]{@{}l@{}}Semantic \\ Similarity \\ Analysis\end{tabular} & 2014--2022                      & immunity              & \begin{tabular}[c]{@{}l@{}}politics (‘legislator’,\\ ‘representative’).\end{tabular} & \begin{tabular}[c]{@{}l@{}}health-related \\ (‘prevention’, ‘AIDS’)\end{tabular}      \\ \cmidrule(l){2-6}
-                                                                                          & \textbf{PART OF SPEECH}                                                                        &                                  &                       &                                                                                      &                                                                   \\ \midrule
+\textbf{}                                            & \begin{tabular}[c]{@{}l@{}}Cultural and \\ Linguistic Drift\end{tabular}         & 1800--2000                            & virus                     & \begin{tabular}[c]{@{}l@{}}infected with virus\end{tabular}                                   & \begin{tabular}[c]{@{}l@{}}spreading\\  computer virus\end{tabular}                   \\ \cmidrule(l){2-6}
+\textbf{}                                            & \begin{tabular}[c]{@{}l@{}}Diachronic Word\\ Embeddings\end{tabular}             & 1800--1999                            & awful                     & full of awe                                                                                   & terrible or appalling                                                                 \\ \cmidrule(l){2-6}
+\textbf{}                                            & \begin{tabular}[c]{@{}l@{}}Contextual Word \\ Embeddings\end{tabular}            & 1910 – 2009                           & tenure                    & \begin{tabular}[c]{@{}l@{}}short term leases,\\ insecurity of tenure\end{tabular}             & \begin{tabular}[c]{@{}l@{}}tenure of office, \\ employment \end{tabular}              \\ \cmidrule(l){2-6}
+\textbf{}                                            & \begin{tabular}[c]{@{}l@{}}Semantic Similarity \\ Analysis\end{tabular}          & 2014--2022                            & immunity                  & \begin{tabular}[c]{@{}l@{}}politics (‘legislator’,\\ ‘representative’).\end{tabular}          & \begin{tabular}[c]{@{}l@{}}health-related \\ (‘prevention’, ‘AIDS’)\end{tabular}      \\ \cmidrule(l){2-6}
+\rowcolor{lightgray}                                 & \multicolumn{5}{c}{\textbf{PART OF SPEECH \cite{kulkarni2014statisticallysignificantdetectionlinguistic, hamilton-etal-2016-cultural}}}                                                                                                                                                                                                      \\
 \textbf{Syntactic}                                   & POS Tags                                                                         & 1900--2000                            & windows                   & \begin{tabular}[c]{@{}l@{}}doors and \\ windows of a house\end{tabular}                       & Microsoft Windows                                                                     \\ \midrule
 \textbf{}                                            & \begin{tabular}[c]{@{}l@{}}Mixed-model \\ Regressions\end{tabular}               & 1800--2000                            & actually                  & originally, nominally                                                                         & presumed, believe\\
-
 \bottomrule
 \end{tabular}
-%\bottomrule
-\caption{Various types of linguistic change and method used to detect them.}
+
+\caption{\textbf{Comparative Overview of Linguistic Change Detection Approaches.}
+This table provides a comprehensive summary of various approaches and contributions to the study of linguistic changes,
+    categorizing the type of changes (semantic or syntactic), and the methodologies used, including frequency and distributional approaches.
+    It highlights each study’s methods, such as word embeddings, statistical analysis, and clustering, along with the time periods of change detection (TCD) and examples of words.}
 \label{tab:related-work}
 \end{table}
-
+\vspace{0mm}
 % ------------------------------------------------------------------ SEMANTIC ------------------------------------------------------------------
@@
     \item \para{Semantic Similarity:} The impact of significant events like the COVID-19 pandemic on language and semantic change has also been a subject of study.
-        Laurino et al.\ (2023) explores tracking fast semantic changes through a large-scale word association task, aiming to understand how the collective mental lexicon evolves in response to such global events.
+    \begin{itemize}
+        \item Laurino et al.\ (2023) explores tracking fast semantic changes through a large-scale word association task, aiming to understand how the collective mental lexicon evolves in response to such global events.
             Their research highlights the dynamic nature of language and how it incorporates new senses.
+        \item One key finding from their work is that words directly related to the pandemic exhibited a greater difference in semantic similarity between pre-pandemic and pandemic time periods.
+            This suggests that these words underwent a more significant and rapid semantic shift compared to control words not associated with the pandemic.
+            They also employ semantic similarity analysis to quantify the shifts in meaning for pandemic-related words and provides evidence that the COVID-19 pandemic acted as a catalyst for rapid semantic change.
             For instance, words like `quarantine', `mask', and `social distancing' took on new and prominent meanings in everyday conversation.
+    \end{itemize}
 \end{itemize}
@@
 After aligning the embeddings for individual time periods, we can use the aligned word vectors to compute the semantic displacement that a word has undergone during a certain time-period.
-    \item Third one by Laurino et al.\ investigates the impact of the COVID-19 pandemic on word meaning.
-One key finding from their work is that words directly related to the pandemic exhibited a greater difference in semantic similarity between pre-pandemic and pandemic time periods.
-This suggests that these words underwent a more significant and rapid semantic shift compared to control words not associated with the pandemic.
-They also employ semantic similarity analysis to quantify the shifts in meaning for pandemic-related words and provides evidence that the COVID-19 pandemic acted as a catalyst for rapid semantic change.
+%    \item Third one by Laurino et al.\ investigates the impact of the COVID-19 pandemic on word meaning.
+%One key finding from their work is that words directly related to the pandemic exhibited a greater difference in semantic similarity between pre-pandemic and pandemic time periods.
+%This suggests that these words underwent a more significant and rapid semantic shift compared to control words not associated with the pandemic.
+%They also employ semantic similarity analysis to quantify the shifts in meaning for pandemic-related words and provides evidence that the COVID-19 pandemic acted as a catalyst for rapid semantic change.
 \end{itemize}

### 2024-08-02 src/sections/paper_kulkarni.tex
added SVG, fixed table
completed section 2 (DONE :) )

@@
 which transitioned from being cheerful or dapper to signifying homosexual or lesbian aspect around the mid-20th century.
-To achieve understanding of this phenomenon, they intend to study the following research questions.
+To achieve understanding of this phenomenon, they studied the following research questions.
 \begin{itemize}
@@
     \centering
-    \includegraphics[scale=0.5]{figures/kulkarni1}
+    \includegraphics[scale=0.8]{figures/kulkarni_drawio}
     \caption{My caption}
@@
 They calculate the change in probability of a word appearing over time.
-\begin{displaymath}
-$T_t(w) = \log \frac{#(w \in C_t)}{|C_t|}$ \label{eq:equation}
-\end{displaymath}
+\begin{equation}
+\mathbf{T_t(w)} = \log \frac{#(w \in C_t)}{|C_t|}
+\label{eq:equation}
+\end{equation}
 
@@
 \textbf{Distributional} & diet, tape, plastic                            &  &  &  \\ \bottomrule
-\end{tabular}\label{tab:time-series-examples}
+\end{tabular}
+\caption{Time Series Analysis}
+\label{tab:time-series-examples}
 \end{table}
@@
 This changed the use of the word diet to mean a life-style of food consumption behavior and not only the food consumed by an individual or group.
+\\
 
@@
 \textbf{Twitter Tweets} & candy, myster, rally, sandy           &  &  &  \\ \bottomrule
-\end{tabular}\label{tab:sources-examples}
+\end{tabular}
+\caption{Cross Domain Analysis}
+\label{tab:sources-examples}
 \end{table}
@@
 
-Despite the fast pace of change of the web content, our method is able to detect the introduction of new products, movies and books.
+Despite the fast pace of change of the web content, their method is able to detect the introduction of new products, movies and books.
 This could help semantically aware web applications to better understand user intentions and requests.

### 2024-08-02 src/sections/relatedwork.tex
completed section 2 (DONE :) )

@@
 This dataset provides n-grams with corresponding yearly occurrences and frequencies.
+
+\begin{figure}[tbh]
+\vspace{-1em}
+\centering
+\includegraphics[scale=0.45]{figures/survey}
+\caption{Timeline of models used to study semantic shifts (taken from \cite{kutuzov-etal-2018-diachronic})}\label{fig:testsvg}
+\end{figure}
 
@@
 
+% ------------------------------------------------------------------ SEMANTIC ------------------------------------------------------------------
+\subsection{Semantic Drift} \label{subsec:semantic-drift}
 \para{Building on the extensive research on \emph{semantic} change.}
@@
 \end{table}
-\vspace{0mm}
-% ------------------------------------------------------------------ SEMANTIC ------------------------------------------------------------------
-\begin{itemize}
-    \item \textbf{Investigating Semantic Shifts Through Word Contexts.}
+\raggedbottom
+%\vspace{-1em}
+
+\para{Investigating Semantic Shifts Through Word Contexts.}
 Gulordava and Baroni (2011) detected semantic change by focusing on words used in the 1960s and 1990s.
 They compared the similarity of the surrounding words (words that co-occur with the target word) in these two time periods.
-    \begin{itemize}
-        \item To assess similarity, the researchers calculated the \emph{Local Mutual Information (LMI)} score between the central word and its surrounding words.
+
+To assess similarity, the researchers calculated the \emph{Local Mutual Information (LMI)} score between the central word and its surrounding words.
 A low LMI score between the target word and its surrounding words across the two time periods indicated a potential semantic shift.
-        \item The study demonstrated that their distributional similarity models were effective in capturing \emph{cultural shifts} in word meaning.
+The study demonstrated that their distributional similarity models were effective in capturing \emph{cultural shifts} in word meaning.
 For example, they found that the word `sleep' acquired more negative connotations related to sleep disorders when comparing its contexts in the 1960s to those in the 1990s.
-    \end{itemize}
 
-    \item \para{Tracking Meaning Evolution Through Neural Nets.}
+\para{Tracking Meaning Evolution Through Neural Nets.}
 Kim et al.\ (2014) focused on analyzing how word meanings evolved between 1900 and 2009.
-    \begin{itemize}
-        \item They developed the first method that employed \emph{prediction-based model} to trace semantic shifts.
+They developed the first method that employed \emph{prediction-based model} to trace semantic shifts.
 This involved training a model on data from a specific year $y_i$ and then using the resulting word vectors as the starting point for training the model on the next year's data $y_i+1$.
-        \item Their method analyzed \emph{global shifts} in a word's vector semantics.
+
+Their method analyzed \emph{global shifts} in a word's vector semantics.
 Additionally, by plotting the time series of a word's distance to its neighboring words in the model's vector space, they visualized the period during which the semantic shift occurred.
-        \item They demonstrated this for the word `cell' compared to its early neighbors, `closet' and `dungeon,' and the more recent neighbors, `phone' and `cordless.'
+They demonstrated this for the word `cell' compared to its early neighbors, `closet' and `dungeon,' and the more recent neighbors, `phone' and `cordless.'
 For `cell,' the identified period of change (1985-2009), which interestingly coincides with the introduction and widespread adoption of cell phones by the public.
-    \end{itemize}
 
-    \item \para{Pinpointing Significant Shifts Statistically.}
+\para{Pinpointing Significant Shifts Statistically.}
 Kulkarni et al.\ (2015) proposes a novel computational approach to identify and quantify the semantic and usage changes in words across various media (new products, movies and books).
-    \begin{itemize}
-        \item Building on the concept of distributed representations proposed by Hinton ~\cite{hinton1986learning}, they map words into a continuous vector space where words with similar meanings are positioned close together.
-        \item The approach hinges on constructing \emph{property time series} for each word.
-        They propose three methods for constructing these time series (\Cref{sec:paper_hamilton}).:
-        \begin{itemize}
+Building on the concept of distributed representations proposed by Hinton ~\cite{hinton1986learning}, they map words into a continuous vector space where words with similar meanings are positioned close together.
+The approach hinges on constructing \emph{property time series} for each word.
+They propose three methods for constructing these time series (\Cref{sec:paper_hamilton}):
+\begin{packed_enumerate}
     \item \textbf{Frequency:}
@@
     and it's utilized to construct distributional time series for a word's displacement.
-        \end{itemize}
-        \item Finally, they employed statistically sound \emph{change point detection algorithm} to identify significant moments in these time series, pinpointing the periods where word meaning or usage likely underwent a shift.
+\end{packed_enumerate}
+Finally, they employed statistically sound \emph{change point detection algorithm} to identify significant moments in these time series, pinpointing the periods where word meaning or usage likely underwent a shift.
 Their results indicated that computational methods for the detection of semantic shifts can be robustly applied to time spans less than a decade.
-    \end{itemize}
 
-    \item \para{Word Semantic Modelling of Polysemant.}
+\para{Word Semantic Modelling of Polysemant.}
 Liao and Cheng (2016) explored the semantic changes of words.
@@
 They focused on \emph{polysemous} words (words with multiple meanings) and aimed to detect when new meanings emerged.
-    \begin{itemize}
-        \item They used the \emph{skip-gram architecture with negative sampling} ~\cite{10.5555/2999792.2999959} to obtain word embeddings.
+
+They used the \emph{skip-gram architecture with negative sampling} ~\cite{10.5555/2999792.2999959} to obtain word embeddings.
 This technique helped in capturing the contextual meaning of words by representing them in a continuous vector space.
-        \item They employed \emph{DBSCAN} to group word embeddings.
+They employed \emph{DBSCAN} to group word embeddings.
 DBSCAN helped in identifying clusters of similar word contexts and distinguishing them from noise, which could indicate semantic changes.
-        \item To find similar words, they used a nearest neighbor search method called \emph{Random Project Forest}.
+To find similar words, they used a nearest neighbor search method called \emph{Random Project Forest}.
 This method helped in identifying words that are contextually similar to a given word.
-        \item Finally, they compared the stability of \emph{similar words} with the stability of their \emph{context words}.
-    \end{itemize}
+ Finally, they compared the stability of \emph{similar words} with the stability of their \emph{context words}.
 
-    \item \para{Distinguishing Cultural Shifts from Linguistic Drift.}
+\para{Distinguishing Cultural Shifts from Linguistic Drift.}
 Hamilton et al.\ (2016) addressed the challenge of distinguishing between \emph{cultural shifts} and \emph{linguistic drift}, both of which can contribute to semantic change.
-    \begin{itemize}
-        \item They proposed two distinct measures based on distributional semantics to distinguish between these two types of semantic change:
-        \begin{itemize}
-            \item \para{Local Neighborhood Measure:} This measure focused on the closest neighbors (most similar words) in a word's embedding.
+They proposed two distinct measures based on distributional semantics to distinguish between these two types of semantic change:
+\begin{packed_enumerate}
+    \item \textbf{Local Neighborhood Measure:}
+    This measure focused on the closest neighbors (most similar words) in a word's embedding.
     A drastic shift in these nearest neighbors suggests a significant change in core meaning, potentially driven by a \emph{cultural shift}
     (e.g., `gay' changing from carefree to referring to homosexuality).
-            \item \para{Global Measure:} This measure considered the overall distribution of a word's surrounding words in a larger context window.
+    \item \textbf{Global Measure:}
+    This measure considered the overall distribution of a word's surrounding words in a larger context window.
     Gradual changes in this broader distribution are more likely to reflect \emph{linguistic drift}, the natural evolution of language due to regular processes
     (e.g., `promise' expanding from a declaration to also suggesting a likelihood).
-        \end{itemize}
-        \item Prior research often treated semantic change as a single phenomenon.
+\end{packed_enumerate}
+Prior research often treated semantic change as a single phenomenon.
 Hamilton et al.\ offered a novel approach by distinguishing between cultural shifts and linguistic drift using their two measures.
-    \end{itemize}
 
-    \item \para{Diachronic Analysis on Historical Data: }
+\para{Diachronic Analysis on Historical Data: }
 The primary goal of Hamilton et al.\ (2016) is to track semantic changes and understand how the meanings of words shift in different historical contexts.
-    \begin{itemize}
-        \item They created word embeddings for different time periods using both the PPMI matrix with SVD and the SGNS model.
+They created word embeddings for different time periods using both the PPMI matrix with SVD and the SGNS model.
 These embeddings were generated from historical text corpora to capture the contextual usage of words in each period.
-        \item After creating the embeddings, they aligned the embeddings.
+After creating the embeddings, they aligned the embeddings.
 Once the embeddings are aligned, they calculate the \emph{semantic displacement} of a word.
 This essentially measured how much a word's vector representation has moved in the embedding space between two time periods.
-        \item The study demonstrates that their method effectively identified semantic change in words and uncovered two statistical `laws' of semantic change:
-        \begin{itemize}
-            \item \para{Law of Conformity}, which suggests that the rate of semantic change is inversely proportional to a word's frequency.
+The study demonstrates that their method effectively identified semantic change in words and uncovered two statistical `laws' of semantic change:
+\begin{packed_enumerate}
+    \item \textbf{Law of Conformity}, which suggests that the rate of semantic change is inversely proportional to a word's frequency.
     High-frequency words tend to change meaning more slowly.
-            \item \para{Law of Innovation}, on the other hand, proposes that words with multiple meanings are more likely to undergo semantic transformations over time.
-        \end{itemize}
-        \item They leveraged the rich information within word embeddings to quantify the degree of semantic change (semantic displacement) and identified potential patterns.
-    \end{itemize}
+    \item \textbf{Law of Innovation}, on the other hand, proposes that words with multiple meanings are more likely to undergo semantic transformations over time.
+\end{packed_enumerate}
+They leveraged the rich information within word embeddings to quantify the degree of semantic change (semantic displacement) and identified potential patterns.
 
-    \item \para{Semantic Shifts with Contextual Embeddings:}
+\para{Semantic Shifts with Contextual Embeddings:}
 Giulianelli et al.\ (2020) explored the phenomenon of lexical semantic change using an unsupervised approach using \emph{BERT}.
-        \begin{itemize}
-            \item They utilized the \emph{BERT} model to generate contextual word embeddings, which captured the meaning of a word based on its surrounding context.
-            \item The extracted word usage vectors are then clustered into different \emph{usage types} using the \emph{k-means clustering algorithm}.
+They utilized the \emph{BERT} model to generate contextual word embeddings, which captured the meaning of a word based on its surrounding context.
+The extracted word usage vectors are then clustered into different \emph{usage types} using the \emph{k-means clustering algorithm}.
 This helped to identify distinct senses or meanings of the words as they appear in various contexts.
-            \item They analyzed these clusters for a specific word across different time periods.
+They analyzed these clusters for a specific word across different time periods.
 The study proposed three metrics to quantify semantic change:
-            \begin{itemize}
+\begin{packed_enumerate}
     \item \textbf{Entropy Difference (ED)}: Measures the change in uncertainty (entropy) of a word’s usage distribution over time.
@@
     \item \textbf{Average Pairwise Distance (APD)}: Computes the average distance between word usage vectors from different periods, indicating shifts in word meaning.
-            \end{itemize}
-            \item The qualitative analysis indicated that the approach could capture various linguistic phenomena, including both synchronic (current usage) and diachronic (historical changes) aspects.
-        \end{itemize}
+\end{packed_enumerate}
+The qualitative analysis indicated that the approach could capture various linguistic phenomena, including both synchronic (current usage) and diachronic (historical changes) aspects.
 
-    \item \para{Semantic Similarity:} The impact of significant events like the COVID-19 pandemic on language and semantic change has also been a subject of study.
-    \begin{itemize}
-        \item Laurino et al.\ (2023) explores tracking fast semantic changes through a large-scale word association task, aiming to understand how the collective mental lexicon evolves in response to such global events.
+\para{Semantic Similarity:}
+The impact of significant events like the COVID-19 pandemic on language and semantic change has also been a subject of study.
+Laurino et al.\ (2023) explores tracking fast semantic changes through a large-scale word association task, aiming to understand how the collective mental lexicon evolves in response to such global events.
 Their research highlights the dynamic nature of language and how it incorporates new senses.
-        \item One key finding from their work is that words directly related to the pandemic exhibited a greater difference in semantic similarity between pre-pandemic and pandemic time periods.
+One key finding from their work is that words directly related to the pandemic exhibited a greater difference in semantic similarity between pre-pandemic and pandemic time periods.
 This suggests that these words underwent a more significant and rapid semantic shift compared to control words not associated with the pandemic.
@@
 For instance, words like `quarantine', `mask', and `social distancing' took on new and prominent meanings in everyday conversation.
-    \end{itemize}
-\end{itemize}
+
+\subsection{Takeaways}\label{subsec:semantic-takeaways}
+Existing research on semantic change detection employs a diverse range of methods.
+Traditional approaches rely on frequency analysis, such as log ratio computations, while more recent studies leverage advanced distributional semantics, including word embeddings generated by techniques like SGNS, PPMI, and SVD.
+Clustering algorithms like DBSCAN and contextual embeddings have also been explored.
+These methods have successfully identified semantic shifts in corpora spanning from the 19th to the 21st centuries, revealing changes in word meanings and usage.
+Despite these advancements, challenges persist, such as the limited availability of methods for languages other than English and the requirement for large datasets.
+Developing algorithms capable of detecting semantic change in smaller corpora is an important area for future research.
 
 % ------------------------------------------------------------------ SYNTACTIC ------------------------------------------------------------------
+\subsection{Syntactic Alterations} \label{subsec:syntactic-alterations}
 \para{Linguistic change studies have shifted to include \emph{syntax} alongside semantics.}
@@
 Nouns, due to their inherent flexibility in meaning, exhibit a greater tendency to undergo these changes driven by cultural shifts.
-While verbs are more likely to participate in gradual semantic changes that follow established linguistic patterns.
-\vspace{0mm}
-\begin{itemize}
-    \item \para{Acquiring a new POS.}
+While verbs are more likely to participate in gradual semantic changes that follow established linguistic patterns. \cite{hamilton-etal-2016-cultural, hamilton-etal-2016-diachronic}
+
+\para{Acquiring a new POS.}
 Kulkarni et al.\ assigned part-of-speech (POS) tags to a large collection of text.
 They calculated the likelihood of a word appearing in specific grammatical contexts over time.
-    \begin{itemize}
-        \item To quantify temporal change, they compared the probability distributions of POS tags for a particular word across different time periods.
+To quantify temporal change, they compared the probability distributions of POS tags for a particular word across different time periods.
 This essentially measured the divergence between these distributions.
-        \item An example they cited was the word `windows'.
+An example they cited was the word `windows'.
 Its POS tag shifted from a common noun (referring to doors and windows of a house) to a proper noun (`Microsoft Windows').
 This highlights how a word's grammatical function can change alongside its meaning.
-    \end{itemize}
 
-    \item \para{Capturing Differences Between Nouns and Verbs.}
+\para{Capturing Differences Between Nouns and Verbs.}
 Hamilton et al.\ (2016) highlight how \emph{cultural changes}, often influenced by new technologies, are closely tied to transformations within \emph{local neighborhoods}, particularly sensitive to shifts in nouns.
 On the other hand, \emph{linguistic changes} are more associated with \emph{global measures} and are particularly responsive to variations in verbs.
-     \begin{itemize}
-         \item To validate this hypothesis, the authors employed a statistical technique called a \emph{linear mixed model} where word type (noun or verb) is a fixed effect,
+
+To validate this hypothesis, the authors employed a statistical technique called a \emph{linear mixed model} where word type (noun or verb) is a fixed effect,
 amount of change measured by each metric (local or global) treated as the dependent variable.
 By analyzing the model's results, they could assess whether there's a significant difference in the way nouns and verbs exhibit change.
-         \item The evolution of words like `actually', `must', and `promise' demonstrate these changes.
+The evolution of words like `actually', `must', and `promise' demonstrate these changes.
 For instance, `must' has transitioned from expressing obligation to indicating necessity, showcasing a common pattern seen in modal verbs.
-     \end{itemize}
-\end{itemize}
+
+\subsection{Takeaways}
+While research in semantic change has advanced, syntactic change remains relatively understudied.
+Kulkarni et al.\ (2015) utilized syntactic analysis through part-of-speech (POS) tagging to observe changes in word usage patterns, emphasizing shifts in syntactic roles over time.
+Hamilton et al.\ (2016) linked syntactic changes to broader cultural shifts, analyzing the evolution of nouns and verbs.
+Their study demonstrated how cultural shifts influenced language use, with certain syntactic changes correlating with broader societal trends.
+However, these studies only scratch the surface of syntactic change, with areas like broadening, narrowing, and amelioration largely unexplored.
+A more comprehensive understanding of language evolution necessitates a deeper dive into syntactic processes and their interplay with semantic shifts.
 
 % ------------------------------------------------------------------ THREE PAPERS ------------------------------------------------------------------
+\subsection{Selected Papers}\label{subsec:selected-papers}
 Several studies from our literature review provide valuable insights for addressing data voids in our analysis.
-\begin{itemize}
-    \item The first being the work by Kulkarni et al.\ which introduces us to time series construction.
+The first being the work by Kulkarni et al.\ which introduces us to time series construction.
 Their distributional methods focuses on finding subtle semantic shits to determine the context where a word in used.
 This concept aligns perfectly with our goal of uncovering shifts in word usage when encountering data voids.
-    \item Second paper by Hamilton et al.\ introduces us to word emebeddings alignment.
+
+Second paper by Hamilton et al.\ introduces us to word emebeddings alignment.
 Since we want to compare word vectors from different time periods, vectors should be aligned in same coordinate axes.
 After aligning the embeddings for individual time periods, we can use the aligned word vectors to compute the semantic displacement that a word has undergone during a certain time-period.
-%    \item Third one by Laurino et al.\ investigates the impact of the COVID-19 pandemic on word meaning.
-%One key finding from their work is that words directly related to the pandemic exhibited a greater difference in semantic similarity between pre-pandemic and pandemic time periods.
-%This suggests that these words underwent a more significant and rapid semantic shift compared to control words not associated with the pandemic.
-%They also employ semantic similarity analysis to quantify the shifts in meaning for pandemic-related words and provides evidence that the COVID-19 pandemic acted as a catalyst for rapid semantic change.
-\end{itemize}
+Next two section \Cref{sec:paper_kulkarni}, \Cref{sec:paper_hamilton} focusses on the selected papers in detail.

### 2024-08-05 src/sections/paper_kulkarni.tex
RP1 DONE. (Section 3)

@@
 Kulkarni et al.\ (2015) explores the problem of detecting changes in word usage patterns over time.
-They introduce methods to identify words that have undergone significant shifts in \emph{meaning}, usage \emph{frequency}, or \emph{context}.
-This is achieved by analyzing historical corpora, which contain vast amounts of text data spanning multiple years or even centuries.
-The study utilizes diverse datasets, including Twitter posts, Amazon product reviews, and Google Book Ngrams, to construct time series for individual words.\\
+They introduce methods to identify words that have undergone significant shifts in \emph{meaning}, usage \emph{frequency}, and \emph{context}.
+This is achieved by analyzing historical corpora, which contain vast amounts of text data spanning multiple centuries.
+The study utilizes diverse datasets, including Twitter posts, Amazon product reviews, and Google Book Ngrams, to construct time series for individual words.
+
 The research conducted by Kulkarni et al.\ exhibits a compelling alignment with our own research objectives,
-specifically focusing on the identification of data voids characterized by shifting semantic meanings over time.
-Their exploration into these linguistic shifts has revealed interesting insights, such as the evolution of the word `gay,'
+specifically focusing on the identification of \emph{data voids} characterized by shifting semantic meanings over time.
+Their exploration into these linguistic shifts has revealed interesting insights, such as the evolution of the word `gay,' (\Cref{fig:kulkarni-example})
 which transitioned from being cheerful or dapper to signifying homosexual or lesbian aspect around the mid-20th century.
+
+\begin{figure}[htb]
+    \centering
+    \vspace{-1em}
+    \includegraphics[scale=0.17]{figures/example-gay}
+    \vspace*{-0.5cm}
+    \caption{Example of word 'gay' undergoing change in meaning over time (taken from \cite{kulkarni2014statisticallysignificantdetectionlinguistic}).}
+    \label{fig:kulkarni-example}
+\end{figure}
+
 To achieve understanding of this phenomenon, they studied the following research questions.
@@
 From these snapshots, a common vocabulary $V$ is created having words which are common across all snapshots.
-To analyze changes in these words, a time series $T(w)$ is constructed for each word $w\in V$.
+To analyze changes in these words, a time series $T(w)$ is constructed for each word $w\in V$ (\Cref{fig:kulkarni-overview}).
+
 Kulkarni et al.\ propose several methods to create this time series, aimed at understanding the evolution of words over time.
@@
     \centering
+    \vspace{-1em}
     \includegraphics[scale=0.8]{figures/kulkarni_drawio}
-    \caption{My caption}
-    \label{fig:label}
+    \vspace*{-0.7cm}
+    \caption{Overview of time series construction.}
+    \label{fig:kulkarni-overview}
 \end{figure}
@@
 
-\textbf{(1) Frequency-based method}: The simplest and most direct method for detecting sudden changes in word usage is by analyzing frequency trends.
-\begin{wrapfigure}{R}{0.25\textwidth}
+\begin{wrapfigure}{r}{0.3\textwidth}
     \centering
-    \includegraphics[width=0.25\textwidth]{figures/frequency-gay}
-    \caption{My caption}
+    \vspace{-1em}
+    \includegraphics[width=0.35\textwidth]{figures/frequency-gay}
+    \vspace*{-0.7cm}
+    \caption{Changes in frequency of 'gay' (taken from \cite{kulkarni2014statisticallysignificantdetectionlinguistic}).}
     \label{fig:example-gay}
@@
 
+\para{Frequency-based method.}
+The simplest and most direct method for detecting sudden changes in word usage is by analyzing frequency trends.
 This approach examines how often a word is used over a given period, providing insights into shifts in its popularity or relevance.
-Changes in frequency can indicate whether a word is gaining new meanings or losing old ones, reflecting broader cultural or societal trends.\\
+Changes in frequency can indicate whether a word is gaining new meanings or losing old ones, reflecting broader cultural or societal trends.
 For example, the word “gay” in \Cref{fig:example-gay} shows a noticeable spike in usage during the 1980s, signaling a shift in its meaning or cultural relevance.
+
+\\
 Tools such as the Google Books Ngram Viewer and Google Trends are used for this purpose, as they provide large datasets and visual representations of word usage across different timeframes.
@@
 
-\textbf{(2) Syntactic method:}
+\para{Syntactic method.}
 Frequency-based metrics, while simple to compute, are susceptible to errors caused by imbalances in the corpus's domain and genre distribution.
@@
 
-\textbf{(3) Distributional method}:
+\para{Distributional method.}
 For detecting subtle semantic changes, which are not changed either due to frequency or through change in part of speech,
@@
 
-\begin{itemize}
-    \item The researchers aimed to \emph{learn word embeddings} by training neural language models on a corpus at different time snapshots.
+The researchers aimed to \emph{learn word embeddings} by training neural language models on a corpus at different time snapshots.
 They initialized word vector representations randomly and optimized the model parameters using stochastic gradient descent.
@@
 After training, word embeddings were normalized by their L2 norm to ensure consistent representation.
-    \item The \emph{alignment} process involved aligning word embeddings from different time snapshots into a unified coordinate system to characterize changes between them.
+
+The \emph{alignment} process involved aligning word embeddings from different time snapshots into a unified coordinate system to characterize changes between them.
 Assumptions were made to aid the alignment process, including the equivalence of spaces under linear transformation and the preservation of local structure of most words over time.
 When the alignment model failed to align a word correctly, it indicated a potential linguistic shift, highlighting the importance of accurate alignment for tracking semantic changes over time.
-    \item By aligning embedding spaces across various time snapshots into a joint embedding space,
+By aligning embedding spaces across various time snapshots into a joint embedding space,
 the distributional method constructs a \emph{distributional time series} that captures the semantic evolution of words over time.
-    \item The paper shows the evolution of the word `tape' over time.
-        Initially, the word `tape' referred to an `adhesive tape' but underwent a semantic shift to also mean a `cassette tape' during the early 1970s.
-\end{itemize}
+
+The paper shows the evolution of the word `tape' over time.
+Initially, the word `tape' referred to an `adhesive tape' but underwent a semantic shift to also mean a `cassette tape'.
+
+\subsection{Takeaways}\label{subsec:takeaways}
+For the first research question, Kulkarni et al.\ (2015) explored multiple dimensions of word usage change over time.
+To address this, they constructed three types of time series: frequency-based, syntactic, and distributional.
+Among these, the distributional time series proved most insightful, as it captures the contextual nuances of word use.
+This approach utilized word embeddings to capture the semantic context of a word within specific time periods.
+Subsequently, these embeddings were aligned into a unified space,
+and it relied on assumptions about word stability and embedding space linearity.
+These limitations affect the method's ability to fully capture complex linguistic changes.
+Despite these challenges, the distributional approach provides a foundational understanding of language dynamics.
+
+%back-propogation alogithm
+%normalization factor
+%negative lof-likelihhod
+%stochastic gradient descent
+%L2 Norm
 
 \subsection{Given that a word's usage has changed, how can the precise moment or period of this shift be determined?} \label{subsec:kulkarni-rq2}
-After constructing time series data for each word in the corpus—whether using frequency-based, syntactic, or distributional methods—the next step is to determine if the word has experienced a significant change.
-If a significant shift is detected, the algorithm then identifies and returns the estimated change point (ECP), which marks the time when this change occurred.
-\begin{itemize}
-    \item Since language exhibits a stochastic drift.
+The authors introduce a method for identifying significant changes in word usage over time using a \emph{change point detection algorithm} based on the \emph{Mean Shift model}.
+The process begins by constructing time series data for each word in the corpus—whether using frequency-based, syntactic, or distributional methods.
+The next step is to determine if the word has experienced a significant change.
+If a significant shift is detected, the algorithm then identifies and returns the \emph{estimated change point (ECP)}, which marks the time when this change occurred.
+
+Since language exhibits a stochastic drift.
 In this context, `stochastic' refers to the randomness or probabilistic nature of the training process,
 where the models are trained on the same dataset but may produce different results each time due to this randomness.
-        To resolve this, time series was normalized for each word by transforming the time series into a Z-score series.
-    \item Then the algorithm utilizes a Mean Shift model to detect changes in the mean of the time series.
-    \item Change points can be thus identified by detecting significant shifts in the mean.
-\end{itemize}
+To resolve this, time series was normalized for each word by transforming the time series into a \emph{Z-score series}.
+
+The Mean Shift model is then applied, which involves calculating the mean of the time series data before and after each potential change point.
+By comparing these means, the algorithm determines if a statistically significant change has occurred at any given point in time.
+
+\para{Change Point Algorithm.}
+The first step is to normalize the time series data for the word $w$.
+After normalization, the algorithm computes a mean shift series, denoted as $K(Z(w))$.
+The mean shift is defined as the difference in the mean of a time series at different time points.
+The algorithm calculates this shift by segmenting the time series around a pivot point, denoted as time point $j$, and comparing the means of the segments before and after this point.
+To assess the significance of the computed mean shifts, the algorithm employs a bootstrapping method.
+This involves generating multiple bootstrap samples, denoted as $BS$.
+This is done by repeatedly drawing samples from the normalized data $Z(w)$ until the number of samples equals $B$, which is predefined.
+For each sample, the algorithm calculates a p-value for each time point $i$.
+This p-value indicates how likely it is that the observed change at that time point is significant.
+The algorithm identifies potential change points by looking for times $j$ where the Z-Score $Z_j(w)$ is greater than or equal to the threshold $\gamma$.
+It then checks the p-values for these change points and selects the one with the smallest p-value, indicating the most significant change.
+The algorithm returns the p-value and the estimated change point (ECP).
+
+This information helps understand exactly when the meaning or usage of the word has shifted over time.
+
+\subsection{Takeaways}\label{subsec:takeaways2}
+The algorithm (\Cref{subsec:kulkarni-rq2}) presented in the paper computes the mean shift series as part of its approach to detect significant changes in linguistic usage over time.
+The paper illustrates this with the example of the word `tape,' which transitioned from meaning adhesive tape to cassette tape, with the change detected in the 1970s.
+Similarly, the word `apple' shifted from being used primarily as a common noun to a proper noun associated with Apple Inc., with this change occurring around 1984.
+These examples illustrate the algorithm’s capability to pinpoint when a word’s usage and meaning undergo significant shifts.
+The paper emphasizes the importance of analyzing these changes to understand linguistic evolution better.
+
+However, the effectiveness of this method can be constrained by its sensitivity to data perturbations and the quality of annotations.
+The performance of the methods, particularly the Syntactic method, declines as the perturbation increases.
+This indicates that while the method may work well under minimal changes, it struggles to maintain accuracy when the changes in the data are more pronounced
 
@@
 
-\para{Time Series Analysis \& Historical Analysis.} \\
+\para{Time Series Analysis \& Historical Analysis.}
+The results presented in Kulkarni et al.\ (2015) illustrate how words acquire new meanings and how these meanings change over time through time series and historical analysis.
 
-\begin{table}[]
+\para{Frequency \& Distributional.}
+For words like `transmitted,' `bitch,' `sex,' and `her,' the frequency and distributional methods reveal significant insights.
+For example, the sharp increase in the frequency of the word `her' around the 1960s can be attributed to the concurrent rise of the feminist movement.
+However, the frequency method can produce many false positives due to the temporary popularity of specific social and political events.
+
+\begin{table}[tbh]
 \begin{tabular}{@{}lllll@{}}
@@
 \end{table}
+\raggedbottom
 
-Frequency \& Distributional - transmistted, bitch, sex, her
-The sharp increase of the word her in frequency around the 1960’s could be attributed to the concurrent rise and popularity of the feminist movement.
-Sudden temporary popularity of specific social and political events could lead the Frequency method to produce many false positives.
+\para{Syntactic Method.}
+The syntactic method uniquely detected the word `apple,' which saw its most frequent part of speech tag shift significantly from `Noun' to `Proper Noun.'
+This method has a low false positive rate but suffers from a high false negative rate, as evidenced by its detection of only two words in the study.
 
-Syntactic - apple, hug, sink, click, handle, windows, bush
-The word apple was detected uniquely by the Syntactic method as its most frequent part of speech tag changed significantly from “Noun” to “Proper Noun”
-The Syntactic method is indicative of having low false positive rate, but suffers from a high false negative rate, given that only two words in the table were detected.
+\para{Distributional.}
+For words like `diet,' `tape,' and `plastic,' the distributional method sheds light on significant changes in their usage.
+The popularity of dieting books, starting with the bestseller `Dr.
+Atkins’ Diet Revolution' by Robert C.
+Atkins in 1972, shifted the meaning of `diet' from simply referring to the food consumed to a lifestyle of food consumption behavior.
 
-Distributional - diet, tape, plastic
-The popularity of books on dieting started with the best selling book Dr. Atkins’ Diet Revolution by Robert C. Atkins in 1972.
-This changed the use of the word diet to mean a life-style of food consumption behavior and not only the food consumed by an individual or group.
-\\
+These findings demonstrate the nuanced ways in which word meanings evolve and the strengths and limitations of different methods in detecting these changes.
+\Cref{tab:time-series-examples} shows all the examples that the paper have shown for different methods.
+While frequency and distributional methods can highlight shifts in usage, they are prone to false positives.
+The syntactic method, though precise, may miss many significant changes.
+Therefore, a combined approach utilizing all methods may offer the most comprehensive insights into linguistic evolution
 
 \para{Cross Domain Analysis.}
-Amazon Reviews and Twitter (content that spans a much shorter time scale as compared to Google Books Ngram Corpus)
-What is the effect of applying distributional method on Amazon Review, Twitter dataset
+The study extends its analysis to datasets like Amazon Reviews and Twitter, which cover much shorter time scales compared to the Google Books Ngram Corpus.
+This cross-domain analysis explores the impact of applying the distributional method on these more temporally constrained datasets.
+\Cref{tab:sources-examples} presents words that exhibit semantic change across various datasets.
 
-\begin{table}[]
+\begin{table}[tbh]
 \begin{tabular}{@{}lllll@{}}
@@
 \end{table}
+\raggedbottom
 
-Words - streaming, ray, combo, rays, twilight
+In the Twitter dataset, the word `sandy' acquired a new sense following Hurricane Sandy hitting the East Coast of the USA\@.
+Similarly, the analysis on Amazon Reviews revealed changes in words associated with new products and technology trends.
 
-Twitter - candy, myster, rally, sandy
-The word sandy acquired a new sense after the incident of Hurricane Sandy hitting the East Coast of USA, which was also mentioned in the data voids example by danah boyd.
+These words demonstrate the method’s ability to capture emerging usages and meanings.
+The example of `sandy' highlights how a natural disaster can influence linguistic shifts, with the word gaining a new context and meaning after the hurricane event.
 
-These examples show that their method can detect the introduction of new products, movies, and books.
+These examples demonstrate that the method can effectively detect the introduction of new products, movies, and books, showcasing its adaptability to different domains and time scales.
+This cross-domain capability is particularly valuable for identifying and understanding contemporary linguistic changes in rapidly evolving contexts like social media and online reviews.
 
 \subsection{Takeaways} \label{subsec:kulkarni-takeaways}
-The paper by Kulkarni et al.\ (2015) contributes a novel and rigorous approach to detecting linguistic changes using statistical methods and word embeddings.
-Their contribution in using large historical corpora and word embeddings to capture the dynamic nature of language.
-A statistical framework that ensures the detected changes are significant, avoiding false positives due to noise or minor fluctuations.
-
-Despite the fast pace of change of the web content, their method is able to detect the introduction of new products, movies and books.
-This could help semantically aware web applications to better understand user intentions and requests.
-
-%Also, after knowing that a change has occurred, we would like to know when did that change really occur so that we can co-relate that the data void occured when a political event happened or how the COVID outbreak event,
-%which could help us identify that the term were introduced due to that event, which can be a future direction we can look into.
-
-%For example, the term "filmyourhospital" was introduced in during 2020 when coronavirus took place, which corresponds to a sudden spike.
-%And this term was never used before 2020, making it a new term that introduced to support the conspiracy theory regarding encouraged people to visit local hospitals to take pictures and videos of empty hospitals to help “prove” that the COVID-19 pandemic is an elaborate hoax.
-%Correlate with historical events.
+The paper by Kulkarni et al.\ (2015) presents an innovative approach to detecting significant linguistic changes over time using statistical methods and word embeddings.
+The goal behind their work is to systematically track and analyze how words evolve in meaning and usage across different temporal contexts.
+They construct time series for each word using three distinct methods:
+frequency, which captures the prevalence of word usage over time;
+syntactic, which examines shifts in part-of-speech tags;
+and distributional, which looks at changes in word co-occurrence patterns.
+They also employ a change point detection algorithm which identifies estimated change points, where significant linguistic shifts occur.
+Their analysis spans various datasets, including Google Books Ngram Viewer, Amazon Reviews, and Twitter,
+demonstrating the method’s adaptability and effectiveness in detecting both gradual and abrupt changes in word meanings across different domains and time scales.

### 2024-08-08 src/sections/introduction.tex
minor changes

@@
 \section{Introduction} \label{sec:introduction}
-In the digital age, online information seekers frequently encounter \emph{search results} laden with \emph{misinformation}, \emph{biased narratives}, and \emph{conspiracy theories}.
-This exposure can potentially lead users to accept and propagate false information.
-However, this phenomenon is not universal across all search queries.
-Certain searches, known as \textit{data voids} are particularly susceptible to these issues.
-\\
-\textit{Data voids} occur when searching for terms that yield limited, non-existent, or highly problematic relevant information.
-Unlike common searches that produce abundant data, queries falling into data voids may return no results or present irrelevant and often inaccurate information.
-Data voids can be mainly categorized into two main types:
-\begin{itemize}
-    \item newly coined terms with no established information base, and
-    \item existing terms whose meanings have evolved over time.
-\end{itemize}
-This report focuses on the latter, examining data voids where context drift occurs.
+Language is a dynamic, living entity that continually adapts to reflect the changing world around us.
+With the evolving language, words take on new meanings or connotations with technological advancements,
+social and cultural shifts, generational differences, globalization, political and social movements, etc.
+The particular words when get emerged in political context are known as data voids.
+Since, they have the behavior of changing meaning over time.
+There can be "voids" in the meaning of words or concepts, especially in specific contexts like politics.
+When a word's meaning shifts rapidly or becomes ambiguous, it can create a kind of "data void" for analysis or understanding.
+
+Data voids in political contexts often occur when terms or phrases take on new or altered meanings,
+particularly when these changes are driven by political actors or movements.
+
+This type of data void is especially significant because:
+
+Rapid meaning shifts: Political events or campaigns can cause sudden changes in how certain terms are used or interpreted.
+Intentional manipulation: Political actors may deliberately redefine terms or introduce new phrases to frame debates in their favor.
+Polarization of language: Words or phrases can take on drastically different meanings in different political contexts or for different groups.
+Dog whistles: Some terms may develop coded political meanings that aren't immediately apparent to all audiences.
+Weaponization of ambiguity: The lack of clear, authoritative information about a term's new political meaning can be exploited to spread misinformation or propaganda.
+
+Media manipulators often manipulate language to shape public opinion.
+Words are carefully chosen to present a particular viewpoint.
+Changes in political ideologies can alter the meaning of terms.
+Broader societal shifts influence political discourse.
+
+These elements can rapidly change the connotations and denotations of words, making them challenging to track and analyze.
+
+data void isn't just about a lack of information, but about the potential for confusion, manipulation, or misinformation as terms are redefined or recontextualized in the political sphere.
+
+This type of data void highlights the dynamic nature of language in political discourse and the challenges it presents for maintaining shared understanding in public debates.
+It also underscores the importance of being aware of how language can be used as a political tool.
+
 The following papers has been selected to explore this phenomenon.
@@
 Mainstream media, particularly those with far-left or far-right biases, may pick up and amplify these narratives, contributing to the misinformation cycle.
-
-\subsection{Example of a data void?(Migrant Caravan)}
-Let's understand data void through an example of a term - \emph{migrant caravan}
-The term ‘migrant caravan’ originally described a group of migrants traveling together for safety and mutual support.
-However, its usage, particularly in political discussions, has taken on a more negative tone.
-Today, it is commonly associated with large groups of migrants, often from Central America, journeying through Mexico to reach the United States.
-
-\subsection{Lifecycle of data void (from long tail to spike)}
-\subsection{Types of Data Void}
-\subsection{Technical Challenge}
-analyzing data void involves context drift over historical time. (How context drift happens with data void.)

### 2024-08-08 src/sections/paper_hamilton.tex
minor changes

@@
 
+In our work on data voids, we have recognized that word frequency significantly impacts their identification.
+For instance, the term `sandy' saw a spike in usage during the Hurricane Sandy event.
+Similarly, terms associated with political agendas also experience increased usage during relevant events.
+This paper provides insights into how frequency influences the evolution of word meanings.
+
+Additionally, our study identifies data voids involving polysemous terms, such as `migrant caravan.'
+This term began to change in meaning when FOX News used it to describe people moving from the Mexico border towards the US, and it continued to evolve with new groups of migrants.
+Understanding these kinds of data voids is crucial, and this paper helps us identify subtle semantic changes in words with multiple meanings,
+enhancing our understanding of how these changes relate to frequency and context.
+
 The paper aims to develop a robust methodology for quantifying semantic change using word embeddings and comparing different approaches to analyzing semantic change.
-This methodology is then applied in a large-scale cross-linguistic analysis spanning 200 years and four languages (English, German, French, and Chinese) to propose the above two statistical laws relating frequency and polysemy to semantic change.
+This methodology is then applied in a large-scale cross-linguistic analysis spanning 200 years and four languages
+(English, German, French, and Chinese) to propose the above two statistical laws relating frequency and polysemy to semantic change.
 
-\subsection{Constructing Word Embeddings.}\label{subsec:constructing-word-embeddings.}
+\subsection{Constructing Word Embeddings.}\label{subsec:constructing-word-embeddings}
 The authors employ three methods to construct word embeddings for different time periods.
@@
 
-\begin{enumerate}
-    \item
-\para{PPMI: }
+\para{Positive Pointwise Mutual Information (PPMI). }
 PPMI is a statistical measure used to quantify the association between a word and its context within a corpus.
@@
 
-\begin{itemize}
-    \item \para{Co-occurrence Matrix Construction.}
 First, a co-occurrence matrix is built, where each entry represents the frequency with which a word (target) and its context (typically a window of neighboring words) appear together in the corpus.
-    \item \para{Joint Probability Calculation.}
-    The joint probability of a target word $w_i$ and a context word $c_j$ is calculated by dividing the co-occurrence frequency of $w_i$ and $c_j$ by the total number of word pairs in the corpus.
-    \item \para{Marginal Probability Calculation.}
-    The marginal probabilities of the target word $w_i$ and the context word $c_j$ are calculated based on the total occurrences of each word in the corpus.
-    \item \para{Calculating PPMI.}
+The \emph{joint probability} of a target word $w_i$ and a context word $c_j$ is calculated by dividing the co-occurrence frequency of $w_i$ and $c_j$ by the total number of word pairs in the corpus.
+The \emph{marginal probability} of the target word $w_i$ and the context word $c_j$ are calculated based on the total occurrences of each word in the corpus.
 The Positive Pointwise Mutual Information (PPMI) between the target word $w_i$ and the context word $c_j$ is calculated using the formula:
-
 \begin{equation}
@@
 \begin{itemize}
-    \item $\hat{p}(w_i, c_j)$ is the estimated \emph{joint probability} of the target word $w_i$ and the context word $c_j$ occurring together.
-    \item $\hat{p}(w_i)$ and $\hat{p}(c_j)$ are the estimated \emph{marginal probabilities} of the target word $w_i$ and the context word $c_j$, respectively.
+\item $\hat{p}(w_i, c_j)$ is the \emph{joint probability} of the target word $w_i$ and the context word $c_j$ occurring together.
+\item $\hat{p}(w_i)$ and $\hat{p}(c_j)$ are the \emph{marginal probabilities} of the target word $w_i$ and the context word $c_j$, respectively.
 \item $\alpha > 0$ is a positive constant used to smooth the values and prevent extreme positive values in the PPMI calculation.
 \end{itemize}
+
 This formula refines the PPMI measure by incorporating $\alpha$ to adjust the log probability ratio.
 The inclusion of $\alpha$ helps manage the sparseness of the data and the impact of low-frequency events, making the measure more robust and stable.
-\end{itemize}
 
-    \item
-\para{SVD: }
+\para{Singular Value Decomposition (SVD). }
 SVD is a mathematical technique that factorizes a matrix into three other matrices, revealing important features of the original matrix.
@@
 SVD decomposes the co-occurrence matrix into three matrices in the following manner:
-
 \begin{equation}
@@
 
-    \item
-\para{SGNS: }
+\para{Skip-Gram with Negative Sampling (SGNS). }
 SGNS is a technique used to learn high-quality word vectors by training on large corpora.
 The primary goal of SGNS is to learn word embeddings such that words that appear in similar contexts have similar vector representations.
-\\
-For a given target word  w  and a context word  c , the model tries to maximize the probability  P(c|w)
-
+For a given target word $w$ and a context word $c$, the model tries to maximize the probability $P(c|w)$
 \begin{equation}
@@
 where:
-
 \begin{itemize}
@@
 
-\\
 The probability $\hat{p}(c_i \mid w_i)$ that a word $c_i$ is a context word of $w_i$ is proportional to the exponential of the dot product of their corresponding vector representations.
 The dot product $\mathbf{w}_i^{\text{SGNS}} \cdot \mathbf{c}_j^{\text{SGNS}}$ measures the similarity between the target word and the context word in the vector space.
-A higher dot product indicates that the words are more likely to co-occur, leading to a higher estimated probability.\\
+A higher dot product indicates that the words are more likely to co-occur, leading to a higher estimated probability.
 
-The model uses this probability estimation to distinguish true word-context pairs from randomly sampled "negative" pairs.
+The model uses this probability estimation to distinguish true word-context pairs from randomly sampled negative pairs.
 The training objective is to maximize the probability of observed (target, context) pairs and minimize the probability of randomly sampled pairs, thereby learning embeddings that reflect the semantic relationships between words.
 
-\end{enumerate}
-
-\subsection{Aligning Embeddings}\label{subsec:aligning-embeddings}
+\para{Aligning Embeddings.}
 Since the embeddings are trained independently for each time period, their vector spaces may not be directly comparable.
 The paper addresses this by aligning the embeddings from different time periods.
-They aligned the emebddings across time periods using orthogonal Procrustes, which transforms the embeddings to minimize the distances between corresponding words in different periods.
+They aligned the emebddings across time periods, which transforms the embeddings to minimize the distances between corresponding words in different periods.
 This step ensures that similar meanings are represented similarly across time.
 
-\subsection{Measuring Semantic Change}\label{subsec:measuring-semantic-change}
-Once the embeddings are aligned, following methods are performed to quantify how much the meaning of word changes over time.
-\begin{itemize}
-    \item \para{Pair-wise similarity time-series}
-        This method measures changes in the similarity between pairs of words over different time periods using cosine similarity.
-        \begin{equation}
-            s^{(t)}(w_i, w_j) = \text{cos-sim}(w_i^{(t)}, w_j^{(t)})
-            \label{eq:equation2}
-        \end{equation}
-        The Spearman correlation $(\rho)$ is employed to measure the relationship between the similarity scores of word pairs and time.
-        This non-parametric method assesses whether the similarity series shows a significant increase or decrease over time, which is crucial for understanding semantic shifts.
+\subsection{Takeaways}\label{subsec:hamilton_takeaways1}
+The authors created three types of word embeddings using PPMI, SVD, and SGNS, each capturing different information.
+PPMI measures how closely related two words are, resulting in sparse, high-dimensional vectors.
+SVD simplifies this information by reducing the dimensionality of the PPMI matrix into dense, low-dimensional vectors.
+Lastly, SGNS learns embeddings by predicting surrounding words in a given context, uncovering deeper semantic and contextual nuances that the other methods may not capture.
+Additionally, they employed alignment technique to align word embeddings across different time periods, ensuring consistency and enabling the analysis of semantic shifts over time.
 
-    \item \para{Measuring semantic displacement}
-        This method quantifies how much a word's meaning has changed by calculating the displacement of its vector representation across different time points.
-        They use the aligned word vectors to compute the \emph{semantic displacement} that a word has undergone during a certain time-period.
-        It is measured by calculating the distance between the vector representations of a word in different time periods.
-        For a given word $w$, if $\mathbf{w}^{(t1)}$ and $\mathbf{w}^{(t2)}$ are its embeddings in two time periods $t1$ and $t2$, respectively, then the semantic displacement $\Delta$ is computed as:
-        \begin{equation}
-            \Delta(w) = \text{cos-dist}(\mathbf{w}^{(t1)}, \mathbf{w}^{(t2)}).\label{eq:equation3}
-        \end{equation}
-\end{itemize}
+%\subsection{Measuring Semantic Change}\label{subsec:measuring-semantic-change}
+%Once the embeddings are aligned, following methods are performed to quantify how much the meaning of word changes over time.
+%\begin{itemize}
+%    \item \para{Pair-wise similarity time-series}
+%        This method measures changes in the similarity between pairs of words over different time periods using cosine similarity.
+%        \begin{equation}
+%            s^{(t)}(w_i, w_j) = \text{cos-sim}(w_i^{(t)}, w_j^{(t)})
+%            \label{eq:equation2}
+%        \end{equation}
+%        The Spearman correlation $(\rho)$ is employed to measure the relationship between the similarity scores of word pairs and time.
+%        This non-parametric method assesses whether the similarity series shows a significant increase or decrease over time, which is crucial for understanding semantic shifts.
+%
+%    \item \para{Measuring semantic displacement}
+%        This method quantifies how much a word's meaning has changed by calculating the displacement of its vector representation across different time points.
+%        They use the aligned word vectors to compute the \emph{semantic displacement} that a word has undergone during a certain time-period.
+%        It is measured by calculating the distance between the vector representations of a word in different time periods.
+%        For a given word $w$, if $\mathbf{w}^{(t1)}$ and $\mathbf{w}^{(t2)}$ are its embeddings in two time periods $t1$ and $t2$, respectively, then the semantic displacement $\Delta$ is computed as:
+%        \begin{equation}
+%            \Delta(w) = \text{cos-dist}(\mathbf{w}^{(t1)}, \mathbf{w}^{(t2)}).\label{eq:equation3}
+%        \end{equation}
+%\end{itemize}
 
-\subsection{Evaluation of various approaches}
-They compared the different approaches discussed in \Cref{subsec:methods-for-quantifying-semantic-change} on \emph{synchronic accuracy} (similarity within time-period) and \emph{diachronic validity} (semantic changes over time).
+\subsection{Evaluation \& Results}
+They compared the different approaches discussed in \Cref{subsec:constructing-word-embeddings} on \emph{synchronic accuracy} (similarity within time-period) and \emph{diachronic validity} (semantic changes over time).
 \begin{itemize}
@@
     It involves testing the methods against historical data to see if they can accurately identify known shifts in word meanings.
-    SGNS perfomed best on both of the tasks:
+    The results indicated that SGNS was particularly effective in capturing these shifts, followed by SVD and PPMI.
     \begin{itemize}
@@
         \item \textbf{Discovering Shifts from data.}
-        Tested whether the methods discover reasonable shifts by examining the top-10 words that changed the most from the 1900s to the 1990s according to the semantic displacement metric (\Cref{subsec:methods-for-quantifying-semantic-change}).
+        Tested whether the methods discover reasonable shifts by examining the top-10 words that changed the most from the 1900s to the 1990s according to the semantic displacement metric.
     \end{itemize}
@@
 This paper explores how different types of word embedding models, such as PPMI, SVD, and SGNS, can be used to study diachronic shifts in word meanings.
-It demonstrates the use of metrics like Spearman correlation to quantify semantic drift and employs Procrustes alignment techniques to align word embeddings across different time periods.
 By analyzing these aligned embeddings, the authors identify two key statistical laws of semantic change.

### 2024-08-08 src/sections/paper_kulkarni.tex
minor changes

@@
     \centering
-    \vspace{-1em}
+    \vspace{-3em}
     \includegraphics[width=0.35\textwidth]{figures/frequency-gay}
-    \vspace*{-0.7cm}
+    \vspace*{-1cm}
     \caption{Changes in frequency of 'gay' (taken from \cite{kulkarni2014statisticallysignificantdetectionlinguistic}).}
@@
 
-\\
 Tools such as the Google Books Ngram Viewer and Google Trends are used for this purpose, as they provide large datasets and visual representations of word usage across different timeframes.
@@
 To address this, they constructed three types of time series: frequency-based, syntactic, and distributional.
-Among these, the distributional time series proved most insightful, as it captures the contextual nuances of word use.
+Among these, the distributional time series proved most insightful, as it captures the contextual usage of word.
 This approach utilized word embeddings to capture the semantic context of a word within specific time periods.
@@
 and it relied on assumptions about word stability and embedding space linearity.
-These limitations affect the method's ability to fully capture complex linguistic changes.
-Despite these challenges, the distributional approach provides a foundational understanding of language dynamics.
 

### 2024-08-08 src/sections/relatedwork.tex
minor changes

@@
 
-\caption{\textbf{Comparative Overview of Linguistic Change Detection Approaches.}
+\caption{\textbf{Comparative overview of linguistic change detection approaches.}
 This table provides a comprehensive summary of various approaches and contributions to the study of linguistic changes,
-categorizing the type of changes (semantic or syntactic), and the methodologies used, including frequency and distributional approaches.
+categorizing the type of changes (semantic or syntactic), and the methodologies used, including frequency, distributional and part of speech approaches.
 It highlights each study’s methods, such as word embeddings, statistical analysis, and clustering, along with the time periods of change detection (TCD) and examples of words.}
@@
 
-\para{Investigating Semantic Shifts Through Word Contexts.}
+\para{Investigating semantic shifts through word contexts.}
 Gulordava and Baroni (2011) detected semantic change by focusing on words used in the 1960s and 1990s.
@@
 
-\para{Tracking Meaning Evolution Through Neural Nets.}
+\para{Tracking meaning evolution through neural nets.}
 Kim et al.\ (2014) focused on analyzing how word meanings evolved between 1900 and 2009.
@@
 
-\para{Pinpointing Significant Shifts Statistically.}
+\para{Pinpointing significant shifts statistically.}
 Kulkarni et al.\ (2015) proposes a novel computational approach to identify and quantify the semantic and usage changes in words across various media (new products, movies and books).
@@
 
-\para{Word Semantic Modelling of Polysemant.}
+\para{Word semantic modelling of polysemant.}
 Liao and Cheng (2016) explored the semantic changes of words.
@@
 
-\para{Distinguishing Cultural Shifts from Linguistic Drift.}
+\para{Distinguishing cultural shifts from linguistic drift.}
 Hamilton et al.\ (2016) addressed the challenge of distinguishing between \emph{cultural shifts} and \emph{linguistic drift}, both of which can contribute to semantic change.
-They proposed two distinct measures based on distributional semantics to distinguish between these two types of semantic change:
+They proposed two distinct measures based on distributional representations to distinguish between these two types of semantic change:
 \begin{packed_enumerate}
-    \item \textbf{Local Neighborhood Measure:}
+    \item \textbf{Local neighborhood masure:}
     This measure focused on the closest neighbors (most similar words) in a word's embedding.
@@
     (e.g., `gay' changing from carefree to referring to homosexuality).
-    \item \textbf{Global Measure:}
+    \item \textbf{Global measure:}
     This measure considered the overall distribution of a word's surrounding words in a larger context window.
@@
 
-\para{Diachronic Analysis on Historical Data: }
+\para{Diachronic analysis on historical data: }
 The primary goal of Hamilton et al.\ (2016) is to track semantic changes and understand how the meanings of words shift in different historical contexts.
@@
 
-\para{Semantic Shifts with Contextual Embeddings:}
+\para{Semantic shifts with contextual embeddings:}
 Giulianelli et al.\ (2020) explored the phenomenon of lexical semantic change using an unsupervised approach using \emph{BERT}.
@@
 
-\para{Semantic Similarity:}
+\para{Semantic similarity:}
 The impact of significant events like the COVID-19 pandemic on language and semantic change has also been a subject of study.
@@
 Existing research on semantic change detection employs a diverse range of methods.
-Traditional approaches rely on frequency analysis, such as log ratio computations, while more recent studies leverage advanced distributional semantics, including word embeddings generated by techniques like SGNS, PPMI, and SVD.
-Clustering algorithms like DBSCAN and contextual embeddings have also been explored.
+Traditional approaches rely on frequency analysis, such as log ratio computations\cite{gulordava-baroni-2011-distributional},
+while more recent studies utilized distributional representations\cite{kim-etal-2014-temporal, kulkarni2014statisticallysignificantdetectionlinguistic, hamilton-etal-2016-cultural},
+including word embeddings generated by techniques like SGNS, PPMI, and SVD\cite{hamilton-etal-2016-diachronic}.
+Clustering algorithms like DBSCAN\cite{10.1007/978-3-319-50496-4_18} was applied to detect semantic shifts in polysemous words.
+By clustering words based on their contextual similarity, they were able to identify when a word’s meaning had shifted, as changes in clusters indicated emerging or evolving senses.
+BERT-based contextual embeddings\cite{giulianelli-etal-2020-analysing} measured lexical semantic change by capturing how word meanings shift subtly across different contexts.
+The advent of word embeddings provided a quantitative framework for measuring semantic change,
+leading to the identification of patterns like the \emph{law of conformity} and \emph{law of innovation} as proposed by Hamilton et al.\ (2016).
 These methods have successfully identified semantic shifts in corpora spanning from the 19th to the 21st centuries, revealing changes in word meanings and usage.
-Despite these advancements, challenges persist, such as the limited availability of methods for languages other than English and the requirement for large datasets.
-Developing algorithms capable of detecting semantic change in smaller corpora is an important area for future research.
+Recent studies, such as Laurino et al.\ (2023), have begun to explore the rapid evolution of language in response to global events like the COVID-19 pandemic,
+highlighting the dynamic nature of semantic change.
+%Despite these advancements, challenges persist, such as the limited availability of methods for languages other than English and the requirement for large datasets.
+%Developing algorithms capable of detecting semantic change in smaller corpora is an important area for future research.
 
@@
 
-\para{Acquiring a new POS.}
+\para{Acquiring a new part of speech.}
 Kulkarni et al.\ assigned part-of-speech (POS) tags to a large collection of text.
@@
 
-\para{Capturing Differences Between Nouns and Verbs.}
+\para{Capturing differences between nouns and verbs.}
 Hamilton et al.\ (2016) highlight how \emph{cultural changes}, often influenced by new technologies, are closely tied to transformations within \emph{local neighborhoods}, particularly sensitive to shifts in nouns.
@@
 
-\subsection{Takeaways}
+\subsection{Takeaways}\label{subsec:takeaways3}
 While research in semantic change has advanced, syntactic change remains relatively understudied.
@@
 However, these studies only scratch the surface of syntactic change, with areas like broadening, narrowing, and amelioration largely unexplored.
-A more comprehensive understanding of language evolution necessitates a deeper dive into syntactic processes and their interplay with semantic shifts.
 

### 2024-08-11 src/sections/introduction.tex
changed title, modified RP1

--- 2024-07-31 src/sections/introduction.tex
@@
 \section{Introduction} \label{sec:introduction}
-In the digital age, online information seekers frequently encounter \emph{search results} laden with \emph{misinformation}, \emph{biased narratives}, and \emph{conspiracy theories}.
-This exposure can potentially lead users to accept and propagate false information.
-However, this phenomenon is not universal across all search queries.
-Certain searches, known as \textit{data voids} are particularly susceptible to these issues.
-\\
-\textit{Data voids} occur when searching for terms that yield limited, non-existent, or highly problematic relevant information.
-Unlike common searches that produce abundant data, queries falling into data voids may return no results or present irrelevant and often inaccurate information.
-Data voids can be mainly categorized into two main types:
-\begin{itemize}
-    \item newly coined terms with no established information base, and
-    \item existing terms whose meanings have evolved over time.
-\end{itemize}
-This report focuses on the latter, examining data voids where context drift occurs.
-The following papers has been selected to explore this phenomenon.
+Language is a dynamic, living entity that continually adapts to reflect the changing world around us.
+With the evolving language, words take on new meanings through technological advancements, cultural shifts, and political movements.
+
+Particularly in political contexts, certain words or phrases can experience rapid shifts in meaning or become ambiguous.
+These are known as `data voids' - situations where the meaning of a term has become unstable or contested, often due to deliberate efforts by political actors to redefine it.
+
+Unlike more general language evolution, where new terms emerge to describe novel concepts, data voids in politics involve the strategic manipulation of existing vocabulary.
+Political actors may repurpose words or introduce new phrases to frame debates in their favor, creating confusion and a lack of shared understanding.
+
+This phenomenon highlights how language can be weaponized as a political tool.
+As terms take on drastically different meanings for different groups, it can become challenging to maintain clear, authoritative information about their usage.
+The resulting `data voids' can then be exploited to spread misinformation or propaganda.
+
+My research goal is to find data voids in political context automatically and systematically study how word shifts their meaning over time.
+To achieve this, we have selected following research papers as a foundation for our research.
+
 \begin{itemize}
     \item Statistically Significant Detection of Linguistic Change.
-    ACM WWW 2014 ~\cite{kulkarni2014statisticallysignificantdetectionlinguistic}
+    ACM WWW 2015 ~\cite{kulkarni2014statisticallysignificantdetectionlinguistic}
     \item Diachronic Word Embeddings Reveal Statistical Laws of Semantic Change.
@@
 
-\subsection{What is a Data Void?}\label{subsec:what-is-a-data-void?}
-Data voids, as first introduced by Golebiewski and Boyd \cite{unknown},
-are identified while searching terms for which relevant information is either limited, non-existent, or deeply problematic.
-\\
-For example, unlike a typical web search for `basketball' that yields a vast amount of data,
-data voids present a problematic scenario such as searches on these terms may return no results or irrelevant, often inaccurate information.
-This situation could prompt users to pursue information, click on misleading search outcomes, and be exposed to false or misleading content,
-ultimately shaping their perceptions and beliefs with misinformation.
+The concept of context drift, as presented in these papers, refers to the change in the contextual usage of words over time.
+This can be particularly useful for detecting and analyzing data voids in political contexts.
 
-Data Voids often originate in fringe communities.
-When vulnerable users search for information using this term, they may encounter results that are misrepresentations of the topic, as these fringe communities have already shaped the narrative.
-Mainstream media, particularly those with far-left or far-right biases, may pick up and amplify these narratives, contributing to the misinformation cycle.
+Finding data voids is significantlly important because
+political events or campaigns can cause sudden changes in how certain terms are used or interpreted.
+Political actors may deliberately redefine terms or introduce new phrases to frame debates in their favor.
+Words or phrases can take on drastically different meanings in different political contexts or for different groups.
+Some terms may develop coded political meanings that aren't immediately apparent to all audiences.
+The lack of clear, authoritative information about a term's new political meaning can be exploited to spread misinformation or propaganda.
 
-\subsection{Example of a data void?(Migrant Caravan)}
-Let's understand data void through an example of a term - \emph{migrant caravan}
-The term ‘migrant caravan’ originally described a group of migrants traveling together for safety and mutual support.
-However, its usage, particularly in political discussions, has taken on a more negative tone.
-Today, it is commonly associated with large groups of migrants, often from Central America, journeying through Mexico to reach the United States.
+These elements can rapidly change the basic meaning (denotation) and the feelings or ideas they evoke (connotation) of words, making them challenging to track and analyze.
 
-\subsection{Lifecycle of data void (from long tail to spike)}
-\subsection{Types of Data Void}
-\subsection{Technical Challenge}
-analyzing data void involves context drift over historical time. (How context drift happens with data void.)
+\para{Example.}
+The term \emph{crisis actor} has undergone a significant shift in meaning,
+exemplifying the phenomenon of context drift we've been discussing.
+
+Originally this term was used in the entertainment and training industries,
+which refers to individuals hired to portray victims or other roles in simulated emergency scenarios.
+These simulations were used for training first responders, military personnel, or for other educational purposes.
+
+However, in recent years, the term has been full of misinformation which refers to genuine victims of violent events (mass shooting) in an attempt to prove that an event was staged or did not happen.
+This term got prominent during Sandy Hook Elementary School shooting (2012), Los Angeles International Airport (LAX) shooting (2013), Pulse Nightclub shooting in Orlando (2016), Las Vegas shooting (2017).
+
+This example clearly demonstrates how a term with a specific, legitimate meaning can be repurposed to spread misinformation and conspiracy theories.
+The shift creates a potential data void where searches for \emph{crisis actor} might lead to a mix of factual information about emergency training and harmful conspiracy theories,
+making it difficult for users to distinguish accurate information.

### 2024-08-11 src/sections/paper_hamilton.tex
changed title, modified RP1

@@
 
-\subsection{Evaluation \& Results}
+\subsection{Evaluation \& Results}\label{subsec:evaluation&results}
 They compared the different approaches discussed in \Cref{subsec:constructing-word-embeddings} on \emph{synchronic accuracy} (similarity within time-period) and \emph{diachronic validity} (semantic changes over time).
@@
     It involves testing the methods against historical data to see if they can accurately identify known shifts in word meanings.
-    The results indicated that SGNS was particularly effective in capturing these shifts, followed by SVD and PPMI.
+    The results indicated that SGNS was particularly effective in capturing these shifts, followed by SVD and PPMI\@.
     \begin{itemize}
@@
 
-\subsection{Statistical Laws of Semantic Change}
+\subsection{Statistical Laws of Semantic Change}\label{subsec:statistical-laws-of-semantic-change}
 The authors establish two statistical laws of semantic change by analyzing a large dataset across multiple languages and time periods.
@@
 
-\subsection{Takeaways}
+\subsection{Takeaways}\label{subsec:takeaways4}
 This paper explores how different types of word embedding models, such as PPMI, SVD, and SGNS, can be used to study diachronic shifts in word meanings.

### 2024-08-11 src/sections/paper_kulkarni.tex
changed title, modified RP1

@@
 
-\para{Construction of Time Series.}
+\para{Construction of time series.}
 In this paper, three methods are explored for statistically modeling the evolution of words over time:
@@
 
-\para{Frequency-based method.}
+\para{Frequency method.}
 The simplest and most direct method for detecting sudden changes in word usage is by analyzing frequency trends.
@@
 To address this, they constructed three types of time series: frequency-based, syntactic, and distributional.
-Among these, the distributional time series proved most insightful, as it captures the contextual usage of word.
+
+Frequency-based method relies on analyzing the frequency of word usage over time.
+It provides immediate insights into trends and spikes in word usage.
+However, the frequency method is prone to sampling errors, it may indicate a change in frequency without a significant shift in meaning,
+as seen with the word `Hurricane' during events like Hurricane Sandy, where frequency increased but meaning remained stable.
+Syntactic-based method analyzes the part of speech (POS) tags of words to detect changes in their syntactic roles over time.
+Distributional-based method is based on the fact that words appearing in similar contexts are semantically similar.
+
+Among the three methods employed by Kulkarni et al.\ (2015) for constructing time series,
+the distributional time series proved most insightful, as it captures the contextual usage of word.
 This approach utilized word embeddings to capture the semantic context of a word within specific time periods.
-Subsequently, these embeddings were aligned into a unified space,
-and it relied on assumptions about word stability and embedding space linearity.
+
+The syntactic method, while reliable in detecting changes in part-of-speech tags, may overlook significant shifts due to its reliance on linguistic taggers.
+At the time of the study, this posed a challenge due to the limitations of tagging accuracy, which has since improved significantly with the advent of more advanced methods for annotating datasets with POS tags.
 
@@
 
-\para{Change Point Algorithm.}
-The first step is to normalize the time series data for the word $w$.
+\begin{figure}[htb]
+    \centering
+    \vspace{-1em}
+    \includegraphics[scale=0.17]{figures/changepoint}
+    \vspace*{-0.5cm}
+    \caption{Change Point Algorithm (taken from \cite{kulkarni2014statisticallysignificantdetectionlinguistic}).}
+    \label{fig:kulkarni-changepoint}
+\end{figure}
+
+\para{Change point algorithm.}
+After constructing time series data for each word in the corpus—whether using frequency-based, syntactic, or distributional methods (\Cref{subsec:constructing-word-embeddings}), the first step is to normalize the time series data for the word $w$.
 After normalization, the algorithm computes a mean shift series, denoted as $K(Z(w))$.
@@
 The algorithm returns the p-value and the estimated change point (ECP).
-
 This information helps understand exactly when the meaning or usage of the word has shifted over time.
+\Cref{fig:kulkarni-changepoint} illustrates the important aspects of the algorithm.
 
 \subsection{Takeaways}\label{subsec:takeaways2}
-The algorithm (\Cref{subsec:kulkarni-rq2}) presented in the paper computes the mean shift series as part of its approach to detect significant changes in linguistic usage over time.
-The paper illustrates this with the example of the word `tape,' which transitioned from meaning adhesive tape to cassette tape, with the change detected in the 1970s.
-Similarly, the word `apple' shifted from being used primarily as a common noun to a proper noun associated with Apple Inc., with this change occurring around 1984.
+The change point detection algorithm (\Cref{subsec:kulkarni-rq2}) presented in the paper detects significant linguistic shifts over time by analyzing word time series data.
+To detect changes, they first convert these time series into Z-score series, which normalize the data and make it easier to identify shifts.
+The Mean Shift model is then applied, which involves calculating the mean of the time series data before and after each potential change point.
+By comparing these means, the algorithm determines if a statistically significant change has occurred at any given point in time.
+
+The paper illustrates this with the example of the word `tape,' which transitioned from meaning adhesive tape to cassette tape, with the change detected in the \emph{1970s}.
+Similarly, the word `apple' shifted from being used primarily as a common noun to a proper noun associated with Apple Inc., with this change occurring around \emph{1984}.
+
 These examples illustrate the algorithm’s capability to pinpoint when a word’s usage and meaning undergo significant shifts.
@@
 
-However, the effectiveness of this method can be constrained by its sensitivity to data perturbations and the quality of annotations.
-The performance of the methods, particularly the Syntactic method, declines as the perturbation increases.
-This indicates that while the method may work well under minimal changes, it struggles to maintain accuracy when the changes in the data are more pronounced
-
 \subsection{Results} \label{subsec:kulkarni-results}
 
-\para{Time Series Analysis \& Historical Analysis.}
+\para{Time series analysis \& historical analysis.}
 The results presented in Kulkarni et al.\ (2015) illustrate how words acquire new meanings and how these meanings change over time through time series and historical analysis.
 
-\para{Frequency \& Distributional.}
+\para{Frequency method.}
 For words like `transmitted,' `bitch,' `sex,' and `her,' the frequency and distributional methods reveal significant insights.
@@
 
-\para{Syntactic Method.}
+\para{Syntactic method.}
 The syntactic method uniquely detected the word `apple,' which saw its most frequent part of speech tag shift significantly from `Noun' to `Proper Noun.'
@@
 
-\para{Distributional.}
+\para{Distributional method.}
 For words like `diet,' `tape,' and `plastic,' the distributional method sheds light on significant changes in their usage.
@@
 
-\para{Cross Domain Analysis.}
+\para{Cross domain analysis.}
 The study extends its analysis to datasets like Amazon Reviews and Twitter, which cover much shorter time scales compared to the Google Books Ngram Corpus.

### 2024-08-11 src/sections/relatedwork.tex
changed title, modified RP1

@@
 to track and examine semantic shifts across historical periods.
+\Cref{fig:timeline} presents a chronological overview of models and events in this domain.
 Much of this research utilized the \emph{Google Books Ngram corpus}
@@
 \includegraphics[scale=0.45]{figures/survey}
-\caption{Timeline of models used to study semantic shifts (taken from \cite{kutuzov-etal-2018-diachronic})}\label{fig:testsvg}
+\caption{Timeline of models used to study semantic shifts (taken from \cite{kutuzov-etal-2018-diachronic})}\label{fig:timeline}
 \end{figure}
@@
 including word embeddings generated by techniques like SGNS, PPMI, and SVD\cite{hamilton-etal-2016-diachronic}.
-Clustering algorithms like DBSCAN\cite{10.1007/978-3-319-50496-4_18} was applied to detect semantic shifts in polysemous words.
+Clustering algorithm like DBSCAN\cite{10.1007/978-3-319-50496-4_18} was applied to detect semantic shifts in polysemous words.
 By clustering words based on their contextual similarity, they were able to identify when a word’s meaning had shifted, as changes in clusters indicated emerging or evolving senses.
@@
 The advent of word embeddings provided a quantitative framework for measuring semantic change,
-leading to the identification of patterns like the \emph{law of conformity} and \emph{law of innovation} as proposed by Hamilton et al.\ (2016).
+leading to the identification of laws like the \emph{law of conformity} and \emph{law of innovation} as proposed by Hamilton et al.\ (2016).
 These methods have successfully identified semantic shifts in corpora spanning from the 19th to the 21st centuries, revealing changes in word meanings and usage.

### 2024-08-13 src/sections/introduction.tex
report writing done, conclusion pending

@@
 
-Particularly in political contexts, certain words or phrases can experience rapid shifts in meaning or become ambiguous.
-These are known as `data voids' - situations where the meaning of a term has become unstable or contested, often due to deliberate efforts by political actors to redefine it.
+\para{Lanaguage evolution in political contexts.}
+The evolution of language is especially evident in political contexts, certain words or phrases can experience rapid shifts in meaning or become ambiguous.
+These are known as \emph{data voids} - situations where the meaning of a term has become unstable or contested, often due to deliberate efforts by political actors to redefine it.
 
@@
 
+\para{Research goal.}
 My research goal is to find data voids in political context automatically and systematically study how word shifts their meaning over time.
@@
     \item Statistically Significant Detection of Linguistic Change.
-    ACM WWW 2015 ~\cite{kulkarni2014statisticallysignificantdetectionlinguistic}
+    ACM WWW 2015 ~\cite{kulkarni2014statisticallysignificantdetectionlinguistic} (\Cref{sec:paper_kulkarni})
     \item Diachronic Word Embeddings Reveal Statistical Laws of Semantic Change.
-    ACL 2016 ~\cite{hamilton-etal-2016-diachronic}
+    ACL 2016 ~\cite{hamilton-etal-2016-diachronic} (\Cref{sec:paper_hamilton})
 \end{itemize}
@@
 The concept of context drift, as presented in these papers, refers to the change in the contextual usage of words over time.
-This can be particularly useful for detecting and analyzing data voids in political contexts.
+This can be particularly useful for detecting and analyzing patterns followed by data voids in political contexts.
 
-Finding data voids is significantlly important because
-political events or campaigns can cause sudden changes in how certain terms are used or interpreted.
+\para{Finding data voids is significantlly important.}
+Political events or campaigns can cause sudden changes in how certain terms are used or interpreted.
 Political actors may deliberately redefine terms or introduce new phrases to frame debates in their favor.
@@
 making it difficult for users to distinguish accurate information.
+\\
+This report introduces methodologies for analyzing data voids based on two key papers.
+To contextualize our work, we present a thorough literature review examining the evolution of word meaning.
+Subsequently, Section \Cref{sec:paper_kulkarni} delves into the first paper, focusing on significant linguistic change detection,
+while Section \Cref{sec:paper_hamilton} explores the second paper on diachronic word embeddings.

### 2024-08-13 src/sections/paper_hamilton.tex
report writing done, conclusion pending

@@
 In our work on data voids, we have recognized that word frequency significantly impacts their identification.
-For instance, the term `sandy' saw a spike in usage during the Hurricane Sandy event.
+For instance, the term `sandy' saw a spike in usage during the Hurricane Sandy event (\Cref{subsec:kulkarni-results}).
 Similarly, terms associated with political agendas also experience increased usage during relevant events.
@@
 \para{Aligning Embeddings.}
-Since the embeddings are trained independently for each time period, their vector spaces may not be directly comparable.
-The paper addresses this by aligning the embeddings from different time periods.
-They aligned the emebddings across time periods, which transforms the embeddings to minimize the distances between corresponding words in different periods.
-This step ensures that similar meanings are represented similarly across time.
+Words that have similar meanings across different time periods should ideally have similar embeddings.
+Alignment ensures that the embeddings from different periods are comparable, making it possible to observe how a word’s meaning shifts over time.
+
+This alignment process involves finding the best way to rotate and transform the embeddings from one time period so that they closely match the embeddings from the next period.
+This is done by finding an orthogonal transformation matrix (let’s call it $\mathbf{R}^{(t)}$) that minimizes the difference between the embeddings from two consecutive time periods.
+The transformation ensures that distances and angles between word vectors (which represent their meanings) are preserved, allowing for an accurate comparison.
+By aligning the embeddings, the authors could measure the semantic displacement of words—how much the meaning of a word has shifted from one time period to another.
 
@@
 
-%\subsection{Measuring Semantic Change}\label{subsec:measuring-semantic-change}
-%Once the embeddings are aligned, following methods are performed to quantify how much the meaning of word changes over time.
-%\begin{itemize}
-%    \item \para{Pair-wise similarity time-series}
-%        This method measures changes in the similarity between pairs of words over different time periods using cosine similarity.
-%        \begin{equation}
-%            s^{(t)}(w_i, w_j) = \text{cos-sim}(w_i^{(t)}, w_j^{(t)})
-%            \label{eq:equation2}
-%        \end{equation}
-%        The Spearman correlation $(\rho)$ is employed to measure the relationship between the similarity scores of word pairs and time.
-%        This non-parametric method assesses whether the similarity series shows a significant increase or decrease over time, which is crucial for understanding semantic shifts.
-%
-%    \item \para{Measuring semantic displacement}
-%        This method quantifies how much a word's meaning has changed by calculating the displacement of its vector representation across different time points.
-%        They use the aligned word vectors to compute the \emph{semantic displacement} that a word has undergone during a certain time-period.
-%        It is measured by calculating the distance between the vector representations of a word in different time periods.
-%        For a given word $w$, if $\mathbf{w}^{(t1)}$ and $\mathbf{w}^{(t2)}$ are its embeddings in two time periods $t1$ and $t2$, respectively, then the semantic displacement $\Delta$ is computed as:
-%        \begin{equation}
-%            \Delta(w) = \text{cos-dist}(\mathbf{w}^{(t1)}, \mathbf{w}^{(t2)}).\label{eq:equation3}
-%        \end{equation}
-%\end{itemize}
+\subsection{Measuring Semantic Change}\label{subsec:measuring-semantic-change}
+Once the embeddings are aligned, following methods are performed to quantify how much the meaning of word changes over time.
+\begin{itemize}
+    \item \para{Pair-wise similarity time-series}
+        This method measures changes in the similarity between pairs of words over different time periods using cosine similarity.
+        \begin{equation}
+            s^{(t)}(w_i, w_j) = \text{cos-sim}(w_i^{(t)}, w_j^{(t)})
+            \label{eq:equation2}
+        \end{equation}
+        The Spearman correlation $(\rho)$ is employed to measure the relationship between the similarity scores of word pairs and time.
+        This non-parametric method assesses whether the similarity series shows a significant increase or decrease over time, which helps us identify specific linguistic or cultural shifts.
+
+    \item \para{Measuring semantic displacement}
+        This method quantifies how much a word's meaning has changed by calculating the displacement of its vector representation across different time points.
+        They use the aligned word vectors to compute the \emph{semantic displacement} that a word has undergone during a certain time-period.
+        It is measured by calculating the distance between the vector representations of a word in different time periods.
+        For a given word $w$, if $\mathbf{w}^{(t1)}$ and $\mathbf{w}^{(t2)}$ are its embeddings in two time periods $t1$ and $t2$, respectively, then the semantic displacement $\Delta$ is computed as:
+        \begin{equation}
+            \Delta(w) = \text{cos-dist}(\mathbf{w}^{(t1)}, \mathbf{w}^{(t2)}).\label{eq:equation3}
+        \end{equation}
+\end{itemize}
 
 \subsection{Evaluation \& Results}\label{subsec:evaluation&results}
-They compared the different approaches discussed in \Cref{subsec:constructing-word-embeddings} on \emph{synchronic accuracy} (similarity within time-period) and \emph{diachronic validity} (semantic changes over time).
-\begin{itemize}
-    \item \para{Synchronic Accuracy}
+In this paper, the authors compared different word embedding approaches — PPMI, SVD, and SGNS \Cref{subsec:constructing-word-embeddings} by evaluating them
+on \emph{synchronic accuracy} (similarity within time-period) and \emph{diachronic validity} (semantic changes over time).
+
+%\begin{figure}[htb]
+%    \centering
+%    \vspace{-1em}
+%    \includegraphics[scale=0.17]{figures/hamilton}
+%    \vspace*{-0.5cm}
+%    \caption{Example of word 'gay' undergoing change in meaning over time (taken from \cite{hamilton-etal-2016-diachronic}).}
+%    \label{fig:hamilton-example}
+%\end{figure}
+
+\para{Synchronic Accuracy.}
 Synchronic accuracy refers to the accuracy with which word embeddings capture the semantic relationships and meanings of words at a \emph{specific point in time}.
 It assesses how well the models represent the semantic relationships among words in a given time period.
-    SVD performed best here.
-    \item \para{Diachronic Validity}
+SVD performed best here followed by PPMI and SGNS\@.
+
+\para{Diachronic Validity.}
 Diachronic validity evaluates how effectively the methods can detect and quantify changes in \emph{meaning over time}.
-    It involves testing the methods against historical data to see if they can accurately identify known shifts in word meanings.
-    The results indicated that SGNS was particularly effective in capturing these shifts, followed by SVD and PPMI\@.
-    \begin{itemize}
-        \item \textbf{Detecting Known Shifts.}
-        The goal in this task is for the methods to correctly capture whether pairs of words moved closer or further apart in semantic space during a pre-determined time-period.
+The authors evaluate the diachronic validity of their methods through two main tasks:
 
-        \item \textbf{Discovering Shifts from data.}
-        Tested whether the methods discover reasonable shifts by examining the top-10 words that changed the most from the 1900s to the 1990s according to the semantic displacement metric.
-    \end{itemize}
+\begin{figure}[tbh]
+    \centering
+    \vspace{-1em}
+    \includegraphics[scale=0.27]{figures/hamilton_known_shifts}
+    \vspace*{-0.5cm}
+    \caption{Known historical shifts from previous works (taken from \cite{hamilton-etal-2016-diachronic}).}
+    \label{fig:hamilton-known-shifts}
+\end{figure}
 
-\end{itemize}
+\textbf{Detecting Known Shifts:}
+This task involves testing whether the methods can accurately capture historical shifts in word meanings that are already documented.
+The goal is to determine if the methods can identify whether pairs of words have moved closer or further apart in semantic space over a specified time period.
+The authors used a set of attested historical shifts as an evaluation set (\Cref{fig:hamilton-known-shifts}) to validate how well the embeddings capture these known changes.
+These examples are taken from previous works on semantic change.
+
+The results showed that all methods performed well in capturing the correct directionality of shifts, meaning they accurately identified whether word pairs became more similar or dissimilar over time.
+
+\para{Discovering Shifts from data.}
+In this task, the authors aimed to see if the methods could discover reasonable shifts in word meanings without prior knowledge of those shifts.
+They examined the top-10 words that changed the most from the 1900s to the 1990s (\Cref{tab:discovering-shifts-table}) based on semantic displacement metric (\Cref{subsec:measuring-semantic-change}).
+These words were categorized as \emph{genuine, borderline, or artifacts of the corpus} by the authors.
+
+SGNS outperformed the other methods, with 70\% of its top-10 list corresponding to genuine semantic shifts, while SVD and PPMI had lower rates of genuine shifts.
+
+\begin{table}[tbh]
+\begin{tabular}{@{}llll@{}}
+\toprule
+\rowcolor[HTML]{FFFFFF}
+\multicolumn{1}{c}{\cellcolor[HTML]{FFFFFF}}                                          & \multicolumn{3}{c}{\cellcolor[HTML]{FFFFFF}\textbf{Examples}}                                                                                                                                                                                                                               \\ \cmidrule(l){2-4}
+\rowcolor[HTML]{FFFFFF}
+\multicolumn{1}{c}{\multirow{-2}{*}{\cellcolor[HTML]{FFFFFF}\textbf{Semantic Shift}}} & \multicolumn{1}{c}{\cellcolor[HTML]{FFFFFF}\textbf{PPMI}}                                                    & \multicolumn{1}{c}{\cellcolor[HTML]{FFFFFF}\textbf{SVD}}                  & \multicolumn{1}{c}{\cellcolor[HTML]{FFFFFF}\textbf{SGNS}}                                        \\ \midrule
+\rowcolor[HTML]{EFEFEF}
+\textbf{Genuine Shift}                                                                & started                                                                                                      & \begin{tabular}[c]{@{}l@{}}headed, calls,\\ gay, actually\end{tabular}    & \begin{tabular}[c]{@{}l@{}}wanting, gay, check, starting,\\ major, actually, headed\end{tabular} \\
+\cellcolor[HTML]{FFFFFF}\textbf{Borderline}                                           & \begin{tabular}[c]{@{}l@{}}know, got, would, \\ decided, think, stop, \\ remember, must, wanted\end{tabular} & male, naturally                                                           & touching                                                                                         \\
+\rowcolor[HTML]{EFEFEF}
+\textbf{Corpus Artifact}                                                              & -                                                                                                             & \begin{tabular}[c]{@{}l@{}}harry, wherever,\\ special, cover\end{tabular} & harry, romance                                                                                   \\ \bottomrule
+\end{tabular}
+\caption{The top 10 words identified by each embedding method were examined, and after reviewing the literature and analyzing their nearest neighbors,
+the authors categorized these words into different types of semantic shifts.
+For instance, the word “headed” shifted from primarily meaning the “top of a body or entity” to signifying “a direction of travel” indicated a \emph{genuine} shift.
+Additionally, some \emph{borderline} cases were identified, which were largely attributed to global shifts in genre or discourse, as well as corpus artifacts.
+For example, words like “special,” “cover,” and “romance” appeared to be \emph{artifacts} from the covers of fiction books, occasionally influenced by advertisements,
+taken from \cite{hamilton-etal-2016-diachronic}}
+\label{tab:discovering-shifts-table}
+\end{table}
 
 \subsection{Statistical Laws of Semantic Change}\label{subsec:statistical-laws-of-semantic-change}
-The authors establish two statistical laws of semantic change by analyzing a large dataset across multiple languages and time periods.
-They demonstrate these laws by involving the comparison of different word embedding models (PPMI, SVD, SGNS) against known historical semantic shifts and novel benchmarks.
+The authors utilized diachronic embeddings to perform a large-scale cross-linguistic analysis,
+revealing statistical laws that connect word frequency and polysemy to the rate of semantic change.
 
-How diachronic embeddings can be used to reveal statistical laws that relate frequency and polysemy to semantic change.
+The authors quantified semantic change using the cosine distance metric between word embeddings across consecutive time periods, expressed as:
+\begin{equation}
+\Delta^{(t)}(w_i) = \text{cos-dist}(\mathbf{w}_i^{(t)}, \mathbf{w}_i^{(t+1)})
+\label{eq:equation4}
+\end{equation}
 
-\begin{itemize}
-    \item \para{Law of conformity:} \emph{Frequently used words change at slower rates.}\\
-    Words with higher frequencies tend to experience slower rates of semantic change.
-    This is statistically supported by the observation that high-frequency words have smaller semantic displacements over time.
-    \item \para{Law of innovation:} \emph{Polysemous words change at faster rates.}\\
-    Polysemous words tend to change more rapidly.
-    The study finds that these words show larger semantic displacements, indicating more significant changes in meaning over time.
-\end{itemize}
+This equation measures how much a word’s meaning has shifted between two time periods t and t+1.
+The value $\Delta^{(t)}(w_i)$ indicates the rate of semantic change for the word $w_i$.
+
+To investigate the relationship between word frequency, polysemy, and semantic displacement, the authors performed regression analysis.
+They considered data points for each word across pairs of consecutive decades.
+Specifically, they analyzed how a word’s frequency and polysemy at time t correlated with its semantic displacement in the subsequent decade.
+
+The analysis was restricted to non-stop words that appeared more than 500 times in both decades contributing to a change.
+This criterion ensured that only words with sufficient co-occurrence data across years were included.
+
+The authors employed a linear mixed model with random intercepts per word and fixed effects for frequency, polysemy, and per decade.
+This model allowed them to estimate the effects of frequency and polysemy on semantic change while controlling for temporal trends and correcting for correlations between measurements on the same word across time.
+
+\para{Law of conformity:} \emph{Frequently used words change at slower rates.}\\
+The Law of Conformity states that the rates of semantic change scale with a negative power of word frequency.
+This means that words that are used more frequently tend to change their meanings at a slower rate compared to less frequently used words.
+
+This finding suggests that words with higher frequencies are more resistant to semantic shifts, confirming the Law of Conformity.
+The authors demonstrated this law by presenting results using SGNS embeddings.
+
+\para{Law of innovation:} \emph{Polysemous words change at faster rates.}\\
+The Law of Innovation posits that words with higher levels of polysemy—meaning they have multiple meanings—experience higher rates of semantic change.
+The authors measured a word’s polysemy by examining its neighborhood in an empirical co-occurrence network.
+They constructed these networks for the top 10,000 non-stop words in each language using the Positive Pointwise Mutual Information (PPMI) measure.
+In these networks, words are connected if they co-occur more often than would be expected by chance.
+The polysemy of a word is then quantified by its local clustering coefficient within this network.
+The clustering coefficient  $d(w_i)$  measures the proportion of a word  $w_i$’s neighbors that are also neighbors of each other.
+
+A high clustering coefficient (and thus a low polysemy score) indicates that the words a given word co-occurs with also tend to co-occur with each other.
+Conversely, polysemous words that appear in disjoint or unrelated contexts will have low clustering coefficients.
+
+The authors found that the logarithm of the polysemy score exhibits a strong positive effect on rates of semantic change.
+This finding supports the Law of Innovation by showing that words with more meanings (higher polysemy) change their meanings more rapidly.
 
@@
 By analyzing these aligned embeddings, the authors identify two key statistical laws of semantic change.
-These findings provide a quantitative framework for understanding how and why languages evolve over time.
+Law of conformity shows that high-frequency words have smaller semantic displacements over time.
+They used statistical models to assess the relationship between word frequency and the rate of semantic change.
+Specifically, they found that the logarithm of a word's frequency had a significant negative effect on the rates of semantic change, indicating that higher frequency words change less rapidly
+The Law of Innovation is established by demonstrating that polysemous words—those with diverse meanings—tend to experience greater semantic change.
+The clustering coefficient was used as a key metric for measuring polysemy within co-occurrence networks, revealing that words with lower clustering coefficients (higher polysemy) are more prone to semantic shifts.
+This law highlights the dynamic nature of words with multiple meanings and their tendency to innovate in the face of linguistic evolution.

### 2024-08-14 src/sections/conclusion.tex
added conclusion section, minor changes

%! Author = mkeim
%! Date = 8/13/24

\section{Conclusion}\label{sec:}
In this report, we explored the mechanisms by which words change their meanings over time and how this process can be leveraged to identify data voids.
We examined two key studies to gain insights into this phenomenon.

The first study by Kulkarni et al.\ (2015) presents a robust framework for detecting significant shifts in word usage over time.
Their methodology involves constructing time series for individual words based on frequency, syntactic, and distributional methods, followed by the application of a change point detection algorithm using the Mean Shift model.
This approach effectively identifies moments in time where notable changes in word usage or meaning occur, as illustrated by examples like the evolving meanings of “tape” and “apple,” and the contextual shift of “sandy” post-Hurricane Sandy.

The second study by Hamilton et al.\ (2016) advances the study of semantic change through the development and application of various word embedding models — PPMI, SVD, and SGNS — to track the evolution of word meanings over time.
By aligning embeddings across different periods and employing metrics like Spearman correlation and Procrustes alignment,
they uncover two important statistical laws: the Law of Conformity, which indicates that more frequent words undergo smaller semantic changes,
and the Law of Innovation, which shows that words with higher polysemy are more likely to experience significant shifts in meaning.
Their findings provide valuable tools and insights into the dynamics of language evolution.

Through these discussions, we have gained a deeper understanding of how semantic change can be detected and analyzed,
offering powerful tools for identifying data voids in evolving linguistic landscapes.

### 2024-08-14 src/sections/introduction.tex
added conclusion section, minor changes

@@
 The shift creates a potential data void where searches for \emph{crisis actor} might lead to a mix of factual information about emergency training and harmful conspiracy theories,
-making it difficult for users to distinguish accurate information.
-\\
+making it difficult for users to distinguish accurate information.\\
+
 This report introduces methodologies for analyzing data voids based on two key papers.
-To contextualize our work, we present a thorough literature review examining the evolution of word meaning.
+To contextualize our work, we present a thorough literature review examining the evolution of word meaning (\Cref{sec:relatedwork}).
 Subsequently, Section \Cref{sec:paper_kulkarni} delves into the first paper, focusing on significant linguistic change detection,

### 2024-08-14 src/sections/paper_hamilton.tex
added conclusion section, minor changes

@@
 \begin{table}[tbh]
+\centering
+\small
 \begin{tabular}{@{}llll@{}}
@@
 For example, words like “special,” “cover,” and “romance” appeared to be \emph{artifacts} from the covers of fiction books, occasionally influenced by advertisements,
-taken from \cite{hamilton-etal-2016-diachronic}}
+taken from \cite{hamilton-etal-2016-diachronic}.}
 \label{tab:discovering-shifts-table}
@@
 The clustering coefficient was used as a key metric for measuring polysemy within co-occurrence networks, revealing that words with lower clustering coefficients (higher polysemy) are more prone to semantic shifts.
-This law highlights the dynamic nature of words with multiple meanings and their tendency to innovate in the face of linguistic evolution.
+This law highlights the dynamic nature of words with multiple meanings and their tendency to innovate in linguistic evolution.

### 2024-08-14 src/sections/relatedwork.tex
added conclusion section, minor changes

@@
 They analyzed these clusters for a specific word across different time periods.
-The study proposed three metrics to quantify semantic change:
-\begin{packed_enumerate}
-    \item \textbf{Entropy Difference (ED)}: Measures the change in uncertainty (entropy) of a word’s usage distribution over time.
-    \item \textbf{Jensen-Shannon Divergence (JSD)}: Compares the similarity of word usage distributions across time intervals.
-    \item \textbf{Average Pairwise Distance (APD)}: Computes the average distance between word usage vectors from different periods, indicating shifts in word meaning.
-\end{packed_enumerate}
+The study proposed three metrics to quantify semantic change:\\
+\textbf{Entropy Difference (ED)}: Measures the change in uncertainty (entropy) of a word’s usage distribution over time.\\
+\textbf{Jensen-Shannon Divergence (JSD)}: Compares the similarity of word usage distributions across time intervals.\\
+\textbf{Average Pairwise Distance (APD)}: Computes the average distance between word usage vectors from different periods, indicating shifts in word meaning.
+
 The qualitative analysis indicated that the approach could capture various linguistic phenomena, including both synchronic (current usage) and diachronic (historical changes) aspects.
@@
 Nouns, due to their inherent flexibility in meaning, exhibit a greater tendency to undergo these changes driven by cultural shifts.
-While verbs are more likely to participate in gradual semantic changes that follow established linguistic patterns. \cite{hamilton-etal-2016-cultural, hamilton-etal-2016-diachronic}
+While verbs are more likely to participate in gradual semantic changes. \cite{hamilton-etal-2016-cultural, hamilton-etal-2016-diachronic}
 
@@
 This concept aligns perfectly with our goal of uncovering shifts in word usage when encountering data voids.
-
 Second paper by Hamilton et al.\ introduces us to word emebeddings alignment.

### 2024-08-15 src/main.tex
completed related work section w/o table (due review)
changes in related work, table pending
added related work section, pending:table citation, caption
modified the table, added citations
worked on RP1 section
added context for hamilton paper
completed section 2 (DONE :) )
minor changes
changed title, modified RP1
report writing done, conclusion pending
added conclusion section, minor changes
minor edits before review

@@
-\documentclass[manuscript, screen, acmsmall]{acmart}
+\documentclass[screen, acmsmall]{acmart}
 \usepackage[noend]{algorithm2e}
@@
 \usepackage{colortbl}
+\usepackage[table,xcdraw]{xcolor}
+\usepackage[normalem]{ulem}
+\usepackage{wrapfig}
+\usepackage{svg}
+\useunder{\uline}{\ul}{}
 
@@
 \PassOptionsToPackage{hyphens}{url}
-\usepackage{hyperref} \urlstyle{tt}
-\newcolumntype{P}[1]{>{\arraybackslash}p{#1}}
+\usepackage{hyperref}
+\usepackage{siunitx}
+\usepackage{wasysym} \urlstyle{tt}
+%\newcolumntype{P}[1]{>{\arraybackslash}p{#1}}
 %\newcolumntype{X}[1]{>{\centering\arraybackslash}p{#1}}
 
-\crefformat{section}{\S#2#1#3}
-\crefformat{subsection}{\S#2#1#3}
-\crefformat{subsubsection}{\S#2#1#3}
+%\expandafter\def\expandafter\UrlBreaks\expandafter{\UrlBreaks%  save the current one
+%  \do\a\do\b\do\c\do\d\do\e\do\f\do\g\do\h\do\i\do\j%
+%  \do\k\do\l\do\m\do\n\do\o\do\p\do\q\do\r\do\s\do\t%
+%  \do\u\do\v\do\w\do\x\do\y\do\z\do\A\do\B\do\C\do\D%
+%  \do\E\do\F\do\G\do\H\do\I\do\J\do\K\do\L\do\M\do\N%
+%  \do\O\do\P\do\Q\do\R\do\S\do\T\do\U\do\V\do\W\do\X%
+%  \do\Y\do\Z}
+
+%\crefformat{section}{\S#2#1#3}
+%\crefformat{subsection}{\S#2#1#3}
+%\crefformat{subsubsection}{\S#2#1#3}
 
@@
   pdftitle={Report for Qualifying Exam},
-  pdfauthor={},
+  pdfauthor={Manisha Keim},
   pdfkeywords={},
@@
 
-\newcommand{\red}[1]{\textcolor{red}{\bf #1}}
-\newcommand{\blue}[1]{\textcolor{blue}{\bf #1}}
+%\newcommand{\red}[1]{\textcolor{red}{\bf #1}}
+%\newcommand{\blue}[1]{\textcolor{blue}{\bf #1}}
 
@@
 \newcommand{\X}{$\mathbb{X}$\xspace}
-\newcommand\mycommfont[1]{\footnotesize\ttfamily\textcolor{blue}{#1}}
+%\newcommand\mycommfont[1]{\footnotesize\ttfamily\textcolor{blue}{#1}}
 
 \begin{document}
-
 \renewcommand\footnotetextcopyrightpermission[1]{} % removes footnote with conference information in first column
 \pagestyle{plain} % removes running headers
-
 \renewcommand{\sectionautorefname}{\S}
@@
 \tableofcontents
+%\input{sections/abstract}
 \newpage
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%% Main document
-%\maketitle
-%\sloppy
-\input{sections/abstract}
+%  \title{Understanding the flow of Temporal Context Shift in Data Voids}
+%  \maketitle
+%  \sloppy
+
+%\newpage
 \input{sections/introduction}
@@
 \newpage
-\input{sections/papers}
+\input{sections/paper_kulkarni}
+\newpage
+\input{sections/paper_hamilton}
+\newpage
+\input{sections/conclusion}
 \balance
-
 \newpage
@@
 \bibliography{main}
-
 \end{document}

### 2024-08-15 src/sections/relatedwork.tex
minor edits before review

--- 2024-08-11 src/sections/relatedwork.tex
@@
 \begin{packed_enumerate}
-    \item \textbf{Entropy Difference (ED)}: Measures the change in uncertainty (entropy) of a word’s usage distribution over time.
-    \item \textbf{Jensen-Shannon Divergence (JSD)}: Compares the similarity of word usage distributions across time intervals.
-    \item \textbf{Average Pairwise Distance (APD)}: Computes the average distance between word usage vectors from different periods, indicating shifts in word meaning.
+    \item Entropy Difference (ED): Measures the change in uncertainty (entropy) of a word’s usage distribution over time.
+    \item Jensen-Shannon Divergence (JSD): Compares the similarity of word usage distributions across time intervals.
+    \item Average Pairwise Distance (APD): Computes the average distance between word usage vectors from different periods, indicating shifts in word meaning.
 \end{packed_enumerate}

### 2024-08-19 src/sections/conclusion.tex
changed introduction and minor changes

@@
 
-Through these discussions, we have gained a deeper understanding of how semantic change can be detected and analyzed,
-offering powerful tools for identifying data voids in evolving linguistic landscapes.
+Through these papers, we have gained a deeper understanding of how semantic change can be detected and analyzed,
+offering powerful methodologies for identifying data voids in evolving linguistic landscapes.

### 2024-08-19 src/sections/introduction.tex
changed introduction and minor changes

@@
 Language is a dynamic, living entity that continually adapts to reflect the changing world around us.
-With the evolving language, words take on new meanings through technological advancements, cultural shifts, and political movements.
-
-\para{Lanaguage evolution in political contexts.}
-The evolution of language is especially evident in political contexts, certain words or phrases can experience rapid shifts in meaning or become ambiguous.
-These are known as \emph{data voids} - situations where the meaning of a term has become unstable or contested, often due to deliberate efforts by political actors to redefine it.
-
-Unlike more general language evolution, where new terms emerge to describe novel concepts, data voids in politics involve the strategic manipulation of existing vocabulary.
-Political actors may repurpose words or introduce new phrases to frame debates in their favor, creating confusion and a lack of shared understanding.
-
-This phenomenon highlights how language can be weaponized as a political tool.
-As terms take on drastically different meanings for different groups, it can become challenging to maintain clear, authoritative information about their usage.
-The resulting `data voids' can then be exploited to spread misinformation or propaganda.
+As language evolves, words shift in meaning, acquire new connotations, or are repurposed for entirely new contexts.
+These changes occur as society, technology, and culture change.
+When words change meaning, there is often a delay before search engines and online platforms fully reflect these new meanings.
+This creates a \emph{data void} — an absence of relevant, up-to-date information based on the current understanding of the word.
+This void represents a gap between evolving language and existing online content.
 
 \para{Research goal.}
-My research goal is to find data voids in political context automatically and systematically study how word shifts their meaning over time.
+Understanding how words evolve is crucial because it allows us to anticipate where data voids might occur.
+By tracking linguistic changes, our goal is to predict when a word might create a gap in digital information which could lead to a potential data void.
 To achieve this, we have selected following research papers as a foundation for our research.
@@
 
-The concept of context drift, as presented in these papers, refers to the change in the contextual usage of words over time.
-This can be particularly useful for detecting and analyzing patterns followed by data voids in political contexts.
+The concept of semantic change, as presented in these papers, refers to the change in the contextual usage of words over time.
+This can be particularly useful for analyzing how words evolve and ensuring that, as meanings shift, data voids are promptly recognized and addressed.
 
-\para{Finding data voids is significantlly important.}
-Political events or campaigns can cause sudden changes in how certain terms are used or interpreted.
-Political actors may deliberately redefine terms or introduce new phrases to frame debates in their favor.
-Words or phrases can take on drastically different meanings in different political contexts or for different groups.
-Some terms may develop coded political meanings that aren't immediately apparent to all audiences.
-The lack of clear, authoritative information about a term's new political meaning can be exploited to spread misinformation or propaganda.
+\para{Finding data voids is significantly important.}
+Data voids have real-world implications, particularly in socio-political and sociotechnical domains.
+When important terms, especially those tied to politics, social issues, or technology, develop new meanings,
+data voids can be exploited to shape public perception.
 
-These elements can rapidly change the basic meaning (denotation) and the feelings or ideas they evoke (connotation) of words, making them challenging to track and analyze.
+Data voids, left unaddressed, can distort debates, influence elections, and shape collective beliefs, often in misleading or harmful directions.
+Once a data void is identified, it’s critical to fill that gap with accurate, authoritative information.
+Otherwise, bad actors can capitalize on the absence of reliable data to promote their own agenda.
+By understanding linguistic changes, researchers and technology platforms can work together to prevent these voids from becoming filled with harmful or misleading content.
 
@@
 The term \emph{crisis actor} has undergone a significant shift in meaning,
-exemplifying the phenomenon of context drift we've been discussing.
+exemplifying the phenomenon of semantic change we've been discussing.
 
-Originally this term was used in the entertainment and training industries,
-which refers to individuals hired to portray victims or other roles in simulated emergency scenarios.
-These simulations were used for training first responders, military personnel, or for other educational purposes.
+Historically, `crisis actor' referred to actors employed in emergency preparedness training exercises.
+These actors simulate victims or other roles in staged disaster drills to help first responders practice their response to crises like natural disasters or terrorist attacks.
+It was a technical, niche term with a specific and legitimate use.
 
-However, in recent years, the term has been full of misinformation which refers to genuine victims of violent events (mass shooting) in an attempt to prove that an event was staged or did not happen.
-This term got prominent during Sandy Hook Elementary School shooting (2012), Los Angeles International Airport (LAX) shooting (2013), Pulse Nightclub shooting in Orlando (2016), Las Vegas shooting (2017).
+However, in recent years, it is used to describe the idea that victims of real-world tragedies are actually actors.
+For example, in mass shootings, the victims have been referred to in propaganda and misinformation campaigns as `crisis actors,'
+falsely implying that the events were staged or fabricated.
 
-This example clearly demonstrates how a term with a specific, legitimate meaning can be repurposed to spread misinformation and conspiracy theories.
-The shift creates a potential data void where searches for \emph{crisis actor} might lead to a mix of factual information about emergency training and harmful conspiracy theories,
-making it difficult for users to distinguish accurate information.\\
+This example clearly demonstrates how a data void occurs when a word’s meaning is co-opted and evolves rapidly,
+leaving a gap in reliable, accurate content.
+This allows misinformation to flourish, having a profound impact on both political discourse and social trust.\\
 
-This report introduces methodologies for analyzing data voids based on two key papers.
+This report introduces methodologies to identify shifts in words and this will be useful in my research to proactively identify data voids based on two key papers.
 To contextualize our work, we present a thorough literature review examining the evolution of word meaning (\Cref{sec:relatedwork}).

### 2024-08-19 src/sections/paper_hamilton.tex
minor edits before review
changed introduction and minor changes

@@
 
-\subsection{Constructing Word Embeddings.}\label{subsec:constructing-word-embeddings}
+\subsection{Constructing Word Embeddings}\label{subsec:constructing-word-embeddings}
 The authors employ three methods to construct word embeddings for different time periods.
@@
 \subsection{Evaluation \& Results}\label{subsec:evaluation&results}
-In this paper, the authors compared different word embedding approaches — PPMI, SVD, and SGNS \Cref{subsec:constructing-word-embeddings} by evaluating them
+In this paper, the authors compared different word embedding approaches — PPMI, SVD, and SGNS (\Cref{subsec:constructing-word-embeddings}) by evaluating them
 on \emph{synchronic accuracy} (similarity within time-period) and \emph{diachronic validity} (semantic changes over time).

### 2024-08-19 src/sections/relatedwork.tex
changed introduction and minor changes

@@
 \subsection{Takeaways}\label{subsec:semantic-takeaways}
-Existing research on semantic change detection employs a diverse range of methods.
-Traditional approaches rely on frequency analysis, such as log ratio computations\cite{gulordava-baroni-2011-distributional},
-while more recent studies utilized distributional representations\cite{kim-etal-2014-temporal, kulkarni2014statisticallysignificantdetectionlinguistic, hamilton-etal-2016-cultural},
-including word embeddings generated by techniques like SGNS, PPMI, and SVD\cite{hamilton-etal-2016-diachronic}.
-Clustering algorithm like DBSCAN\cite{10.1007/978-3-319-50496-4_18} was applied to detect semantic shifts in polysemous words.
+Existing research on semantic change detection employed a diverse range of methods.
+Traditional approaches relied on frequency analysis, such as log ratio computations~\cite{gulordava-baroni-2011-distributional},
+while more recent studies utilized distributional representations ~\cite{kim-etal-2014-temporal, kulkarni2014statisticallysignificantdetectionlinguistic, hamilton-etal-2016-cultural},
+including word embeddings generated by techniques like SGNS, PPMI, and SVD ~\cite{hamilton-etal-2016-diachronic}.
+Clustering algorithm like DBSCAN ~\cite{10.1007/978-3-319-50496-4_18} was applied to detect semantic shifts in polysemous words.
 By clustering words based on their contextual similarity, they were able to identify when a word’s meaning had shifted, as changes in clusters indicated emerging or evolving senses.
-BERT-based contextual embeddings\cite{giulianelli-etal-2020-analysing} measured lexical semantic change by capturing how word meanings shift subtly across different contexts.
+BERT-based contextual embeddings ~\cite{giulianelli-etal-2020-analysing} measured lexical semantic change by capturing how word meanings shift subtly across different contexts.
 The advent of word embeddings provided a quantitative framework for measuring semantic change,
@@
 After aligning the embeddings for individual time periods, we can use the aligned word vectors to compute the semantic displacement that a word has undergone during a certain time-period.
-Next two section \Cref{sec:paper_kulkarni}, \Cref{sec:paper_hamilton} focusses on the selected papers in detail.
+Next two sections \Cref{sec:paper_kulkarni}, \Cref{sec:paper_hamilton} focusses on the selected papers in detail.

### 2024-08-20 src/sections/introduction.tex
introduction changes

@@
 These changes occur as society, technology, and culture change.
+In the digital age, search engines and online platforms play a crucial role in shaping access to knowledge and information.
 When words change meaning, there is often a delay before search engines and online platforms fully reflect these new meanings.
@@
 Understanding how words evolve is crucial because it allows us to anticipate where data voids might occur.
-By tracking linguistic changes, our goal is to predict when a word might create a gap in digital information which could lead to a potential data void.
-To achieve this, we have selected following research papers as a foundation for our research.
+By tracking linguistic changes, we aim to predict when a word might create a gap in digital information which could lead to a potential data void.
+To achieve this, we have selected the following research papers as a foundation for our study.
 
@@
 
-The concept of semantic change, as presented in these papers, refers to the change in the contextual usage of words over time.
-This can be particularly useful for analyzing how words evolve and ensuring that, as meanings shift, data voids are promptly recognized and addressed.
+These papers present the concept of semantic change, which refers to the evolution of a word’s contextual usage over time.
+This approach is particularly useful for analyzing how words evolve and ensuring that, as meanings shift, data voids are promptly recognized and addressed.
 
@@
 Data voids have real-world implications, particularly in socio-political and sociotechnical domains.
-When important terms, especially those tied to politics, social issues, or technology, develop new meanings,
-data voids can be exploited to shape public perception.
+When important terms, especially those tied to politics, social issues, or technology, gain new meanings,
+people can exploit data voids to shape public perception.
 
-Data voids, left unaddressed, can distort debates, influence elections, and shape collective beliefs, often in misleading or harmful directions.
+Data voids, left unaddressed, can distort debates, influence elections, and shape collective beliefs, often in misleading or harmful ways.
 Once a data void is identified, it’s critical to fill that gap with accurate, authoritative information.
-Otherwise, bad actors can capitalize on the absence of reliable data to promote their own agenda.
-By understanding linguistic changes, researchers and technology platforms can work together to prevent these voids from becoming filled with harmful or misleading content.
+Otherwise, bad actors can take advantage of the absence of reliable data to promote their own agendas.
+By understanding linguistic changes, researchers and technology platforms can work together to prevent harmful or misleading content from filling these voids.
 
 \para{Example.}
-The term \emph{crisis actor} has undergone a significant shift in meaning,
+The term \emph{crisis actor} has significantly shifted in meaning,
 exemplifying the phenomenon of semantic change we've been discussing.
@@
 Historically, `crisis actor' referred to actors employed in emergency preparedness training exercises.
-These actors simulate victims or other roles in staged disaster drills to help first responders practice their response to crises like natural disasters or terrorist attacks.
+These actors simulated victims or other roles in staged disaster drills to help first responders practice their response to crises like natural disasters or terrorist attacks.
 It was a technical, niche term with a specific and legitimate use.
 
-However, in recent years, it is used to describe the idea that victims of real-world tragedies are actually actors.
-For example, in mass shootings, the victims have been referred to in propaganda and misinformation campaigns as `crisis actors,'
-falsely implying that the events were staged or fabricated.
+However, in recent years, it is used to describe the false idea that victims of real-world tragedies are actually actors.
+For example, in mass shootings, propaganda and misinformation campaigns have referred to victims as “crisis actors,” falsely implying that the events were staged or fabricated.
 
-This example clearly demonstrates how a data void occurs when a word’s meaning is co-opted and evolves rapidly,
+This example clearly demonstrates how a data void occurs when people co-opt a word’s meaning and cause it to evolve rapidly,
 leaving a gap in reliable, accurate content.
@@
 
-This report introduces methodologies to identify shifts in words and this will be useful in my research to proactively identify data voids based on two key papers.
-To contextualize our work, we present a thorough literature review examining the evolution of word meaning (\Cref{sec:relatedwork}).
+This report introduces methodologies to identify shifts in words and this will be useful in our research to proactively identify data voids based on two key papers.
+To contextualize our work, we provide a thorough literature review examining the evolution of word meaning (\Cref{sec:relatedwork}).
 Subsequently, Section \Cref{sec:paper_kulkarni} delves into the first paper, focusing on significant linguistic change detection,

### 2024-08-20 src/sections/paper_hamilton.tex
introduction changes

@@
 They examined the top-10 words that changed the most from the 1900s to the 1990s (\Cref{tab:discovering-shifts-table}) based on semantic displacement metric (\Cref{subsec:measuring-semantic-change}).
-These words were categorized as \emph{genuine, borderline, or artifacts of the corpus} by the authors.
+The authors categorize the words identified as having shifted meanings into three distinct categories:\\
+\emph{Genuine shift:} Words that have undergone a clear and significant change in meaning over time.
+An example provided in the paper is the word `headed,' which shifted from primarily referring to the `top of a body/entity' to referring to `a direction of travel.'\\
+\emph{Borderline:} These are cases where the shift is not as clear-cut, often due to global genre or discourse shifts.
+These shifts might be more subtle and may not represent a complete change in meaning but rather a shift in usage within specific contexts.\\
+\emph{Corpus artifacts:} These are shifts that arise not from a true change in meaning but rather from peculiarities or biases in the corpus itself.
+For example, words like `special,' `cover,' and `romance' might appear to shift in meaning due to their frequent use in certain contexts like book covers or advertisements, which do not necessarily indicate a genuine semantic change.
 
@@
 \caption{The top 10 words identified by each embedding method were examined, and after reviewing the literature and analyzing their nearest neighbors,
-the authors categorized these words into different types of semantic shifts.
-For instance, the word “headed” shifted from primarily meaning the “top of a body or entity” to signifying “a direction of travel” indicated a \emph{genuine} shift.
-Additionally, some \emph{borderline} cases were identified, which were largely attributed to global shifts in genre or discourse, as well as corpus artifacts.
-For example, words like “special,” “cover,” and “romance” appeared to be \emph{artifacts} from the covers of fiction books, occasionally influenced by advertisements,
 taken from \cite{hamilton-etal-2016-diachronic}.}

### 2024-08-20 src/sections/paper_kulkarni.tex
changed introduction and minor changes
introduction changes

@@
 
-\begin{figure}[htb]
-    \centering
-    \vspace{-1em}
-    \includegraphics[scale=0.17]{figures/changepoint}
-    \vspace*{-0.5cm}
-    \caption{Change Point Algorithm (taken from \cite{kulkarni2014statisticallysignificantdetectionlinguistic}).}
-    \label{fig:kulkarni-changepoint}
-\end{figure}
+%\begin{figure}[htb]
+%    \centering
+%    \vspace{-1em}
+%    \includegraphics[scale=0.17]{figures/changepoint}
+%    \vspace*{-0.5cm}
+%    \caption{Change Point Algorithm (taken from \cite{kulkarni2014statisticallysignificantdetectionlinguistic}).}
+%    \label{fig:kulkarni-changepoint}
+%\end{figure}
 
@@
 After normalization, the algorithm computes a mean shift series, denoted as $K(Z(w))$.
-The mean shift is defined as the difference in the mean of a time series at different time points.
+The \emph{mean shift} is defined as the difference in the mean of a time series at different time points.
 The algorithm calculates this shift by segmenting the time series around a pivot point, denoted as time point $j$, and comparing the means of the segments before and after this point.
-To assess the significance of the computed mean shifts, the algorithm employs a bootstrapping method.
-This involves generating multiple bootstrap samples, denoted as $BS$.
+To assess the significance of the computed mean shifts, the algorithm employs a \emph{bootstrapping} method.
+The algorithm uses bootstrap sampling to ensure that the detected change point is not just due to random noise.
+It generates multiple synthetic time series by randomly resampling the original time series, denoted as $BS$.
+These resampled sets serve as a reference to evaluate whether the detected shifts in the real data are significant or not.
 This is done by repeatedly drawing samples from the normalized data $Z(w)$ until the number of samples equals $B$, which is predefined.
@@
 This information helps understand exactly when the meaning or usage of the word has shifted over time.
-\Cref{fig:kulkarni-changepoint} illustrates the important aspects of the algorithm.
+%\Cref{fig:kulkarni-changepoint} illustrates the important aspects of the algorithm.
 

### 2024-08-20 src/sections/relatedwork.tex
introduction changes

@@
 \subsection{Selected Papers}\label{subsec:selected-papers}
-Several studies from our literature review provide valuable insights for addressing data voids in our analysis.
-The first being the work by Kulkarni et al.\ which introduces us to time series construction.
-Their distributional methods focuses on finding subtle semantic shits to determine the context where a word in used.
-This concept aligns perfectly with our goal of uncovering shifts in word usage when encountering data voids.
-Second paper by Hamilton et al.\ introduces us to word emebeddings alignment.
-Since we want to compare word vectors from different time periods, vectors should be aligned in same coordinate axes.
-After aligning the embeddings for individual time periods, we can use the aligned word vectors to compute the semantic displacement that a word has undergone during a certain time-period.
-Next two sections \Cref{sec:paper_kulkarni}, \Cref{sec:paper_hamilton} focusses on the selected papers in detail.
+Several studies from our literature review provided valuable insights for addressing data voids in our analysis.
+The first was the work by Kulkarni et al.\ which introduced us to time series construction.
+Their distributional methods focused on finding subtle semantic shits to determine the context in which a word in used.
+This concept aligned perfectly with our goal of uncovering shifts in word usage when encountering data voids.
+The second paper by Hamilton et al.\ introduces us to word emebeddings alignment.
+Since we wanted to compare word vectors from different time periods, the vectors had to be aligned on the same coordinate axes.
+After aligning the embeddings for individual time periods,
+we could use the aligned word vectors to compute the semantic displacement that a word has underwent during a certain time-period.
+The next two sections \Cref{sec:paper_kulkarni} and \Cref{sec:paper_hamilton} focused on the selected papers in detail.

### 2024-08-23 src/sections/conclusion.tex
changes in sec 4.5, conclusion, front page

--- 2024-08-14 src/sections/conclusion.tex
@@
 \section{Conclusion}\label{sec:}
-In this report, we explored the mechanisms by which words change their meanings over time and how this process can be leveraged to identify data voids.
+In this report, we explored the mechanisms by which words change their meanings over time and how this process can be leveraged to identify \emph{data voids}.
 We examined two key studies to gain insights into this phenomenon.
 
-The first study by Kulkarni et al.\ (2015) presents a robust framework for detecting significant shifts in word usage over time.
-Their methodology involves constructing time series for individual words based on frequency, syntactic, and distributional methods, followed by the application of a change point detection algorithm using the Mean Shift model.
-This approach effectively identifies moments in time where notable changes in word usage or meaning occur, as illustrated by examples like the evolving meanings of “tape” and “apple,” and the contextual shift of “sandy” post-Hurricane Sandy.
+Kulkarni et al.\ (2015) introduced the concept of creating \emph{word embeddings} to measure the context of a word over time.
+By constructing these embeddings and applying \emph{change point detection}, they provided a powerful method for identifying significant shifts in word meanings.
+This approach can be instrumental in detecting data voids, as abrupt changes in word usage may signal emerging topics or areas where existing data is insufficient or outdated.
 
-The second study by Hamilton et al.\ (2016) advances the study of semantic change through the development and application of various word embedding models — PPMI, SVD, and SGNS — to track the evolution of word meanings over time.
-By aligning embeddings across different periods and employing metrics like Spearman correlation and Procrustes alignment,
-they uncover two important statistical laws: the Law of Conformity, which indicates that more frequent words undergo smaller semantic changes,
-and the Law of Innovation, which shows that words with higher polysemy are more likely to experience significant shifts in meaning.
-Their findings provide valuable tools and insights into the dynamics of language evolution.
+Hamilton et al.\ (2016) built on this foundation by applying variations of the word2vec model, particularly using the Skip-Gram with Negative Sampling (SGNS) method,
+to track the evolution of word meanings across different time periods.
+Hamilton’s unique contribution lies in the formulation of statistical laws governing semantic change, such as the Law of Conformity and the Law of Innovation.
+The \emph{Law of Conformity} indicates that more frequent words tend to undergo smaller semantic changes,
+suggesting that less frequent words may be more prone to shifts that could lead to data voids.
+Meanwhile, the \emph{Law of Innovation} reveals that words with higher polysemy are more likely to experience significant shifts in meaning.
+This insight is crucial for identifying data voids, as words with multiple meanings that rapidly evolve may create gaps in data coverage,
+highlighting areas where new data collection or analysis is needed to keep pace with linguistic changes.
 
-Through these discussions, we have gained a deeper understanding of how semantic change can be detected and analyzed,
-offering powerful tools for identifying data voids in evolving linguistic landscapes.
+Monitoring such shifts can help identify areas where existing data may no longer capture the full spectrum of meanings.

### 2024-08-23 src/sections/paper_kulkarni.tex
changes in sec 4.5, conclusion, front page

@@
 To assess the significance of the computed mean shifts, the algorithm employs a \emph{bootstrapping} method.
-The algorithm uses bootstrap sampling to ensure that the detected change point is not just due to random noise.
-It generates multiple synthetic time series by randomly resampling the original time series, denoted as $BS$.
-These resampled sets serve as a reference to evaluate whether the detected shifts in the real data are significant or not.
+The algorithm uses bootstrap samplingto estimate how likely an observed shift is, under the assumption that no real change has occurred (the null hypothesis).
+It generates multiple time series by randomly resampling the original time series, denoted as $BS$.
 This is done by repeatedly drawing samples from the normalized data $Z(w)$ until the number of samples equals $B$, which is predefined.
-For each sample, the algorithm calculates a p-value for each time point $i$.
-This p-value indicates how likely it is that the observed change at that time point is significant.
-The algorithm identifies potential change points by looking for times $j$ where the Z-Score $Z_j(w)$ is greater than or equal to the threshold $\gamma$.
+The p-value is then calculated by comparing the observed mean shift to the distribution of mean shifts from the bootstrapped samples.
+It reflects the proportion of times that the mean shift from the random (permuted) samples exceeds the observed mean shift.
+A low p-value indicates that the observed shift is unlikely to be due to random fluctuations alone.
+The algorithm identifies potential change points by looking for times $j$ where the Z-Score $Z_j(w)$ is greater than or equal to the threshold $\gamma$,
+which is set to the 96th percentile, to filter out less significant potential change points.
 It then checks the p-values for these change points and selects the one with the smallest p-value, indicating the most significant change.
